### 2019-10-11 tex/paper.tex
Add tex

%%
%% This is file `sample-sigconf.tex',
%% generated with the docstrip utility.
%%
%% The original source files were:
%%
%% samples.dtx  (with options: `sigconf')
%%
%% IMPORTANT NOTICE:
%%
%%% For the copyright see the source file.
%%
%% Any modified versions of this file must be renamed
%% with new filenames distinct from sample-sigconf.tex.
%%
%% For distribution of the original source see the terms
%% for copying and modification in the file samples.dtx.
%%
%% This generated file may be distributed as long as the
%% original source files, as listed above, are part of the
%% same distribution. (The sources need not necessarily be
%% in the same archive or directory.)
%%
%% The first command in your LaTeX source must be the \documentclass command.

% Modify this to remove all ACM and conference information (for arXiv for example)
\def\removeHeaders{no}

\def\tempYes{yes}
\documentclass[10pt,sigconf,letterpaper,dvipsnames\ifx\removeHeaders\tempYes ,nonacm\fi]{acmart}
\acmYear{2019}
\copyrightyear{2019}
\acmConference{Workshop on Buffer Sizing}{December 2-3, 2019}{Stanford, California, USA}

\usepackage[utf8]{inputenc}

\usepackage{multirow}
\usepackage{blindtext}
\usepackage{soul}
\usepackage[inline]{enumitem}
\usepackage{algorithmicx}
\usepackage{algpseudocode}
\usepackage{algorithm}

% Acronyms
\usepackage[nomain, toc, acronym]{glossaries}
\glsdisablehyper

\ifx\removeHeaders\tempYes
\settopmatter{printacmref=false} % Removes citation information below abstract
\renewcommand\footnotetextcopyrightpermission[1]{} % removes footnote with conference information in first column
\fi

\newcommand\note[2]{{\color{#1}#2}}
\newcommand\todo[1]{{\note{red}{TODO: #1}}}

\makeatletter
\newcommand{\StateIndent}[1][3]{%
  \setlength\@tempdima{\algorithmicindent}%
  \Statex\hskip\dimexpr#1\@tempdima\relax}
\algdef{S}[WHILE]{WhileNoDo}[1]{\algorithmicwhile\ #1}%
\makeatother

\algnewcommand\True{\textbf{true}}
\algnewcommand\False{\textbf{false}}
\algnewcommand\Break{\textbf{break}}

\newcommand{\algorithmautorefname}{Algorithm}

%%
%% \BibTeX command to typeset BibTeX logo in the docs
\AtBeginDocument{%
  \providecommand\BibTeX{{%
    \normalfont B\kern-0.5em{\scshape i\kern-0.25em b}\kern-0.8em\TeX}}}

%%
%% Submission ID.
%% Use this when submitting an article to a sponsored event. You'll
%% receive a unique submission ID from the organizers
%% of the event, and this ID should be used as the parameter to this command.
%%\acmSubmissionID{123-A56-BU3}

%%
%% The majority of ACM publications use numbered citations and
%% references.  The command \citestyle{authoryear} switches to the
%% "author year" style.
%%
%% If you are preparing content for an event
%% sponsored by ACM SIGGRAPH, you must use the "author year" style of
%% citations and references.
%% Uncommenting
%% the next command will enable that style.
%%\citestyle{acmauthoryear}

%%
%% end of the preamble, start of the body of the document source.
\begin{document}

%%
%% The "title" command has an optional parameter,
%% allowing the author to define a "short title" to be used in page headers.
\title{Congestion Control Aware Queuing}

%% GLOSSARY

\newacronym{aqm}{AQM}{Active Queue Management}
\newacronym{cc}{CC}{Congestion Control}
\newacronym{cca}{CCA}{Congestion Control Algorithm}
\newacronym{qdisc}{qdisc}{queuing discipline}
\newacronym{fq}{FQ}{Fair Queuing}
\newacronym{rtt}{RTT}{Round-Trip Time}

\begin{abstract}
Recent model-based congestion control algorithms such as BBR use repeated measurements at the end-point to build a model of the network connection and use it to achieve optimal throughput with low queuing delay. Conversely, applying this model-based approach to Active Queue Management (AQM) has so far been underinvestigated. We propose an AQM scheduler based on fair queuing, which adapts the buffer size depending on the needs of each flow without requiring active participation from the endpoint. We implement this scheduler for the Linux kernel and show that it interacts well with the most common congestion control algorithms and can significantly increase throughput compared to CoDeL while avoiding overbuffering.
\end{abstract}

%%
%% The "author" command and its associated commands are used to define
%% the authors and their affiliations.
%% Of note is the shared affiliation of the first two authors, and the
%% "authornote" and "authornotemark" commands
%% used to denote shared contribution to the research.
\author{Maximilian Bachl, Joachim Fabini, Tanja Zseby}
\affiliation{\institution{Technische Universität Wien}}
\email{firstname.lastname@tuwien.ac.at}

%%
%% By default, the full list of authors will be used in the page
%% headers. Often, this list is too long, and will overlap
%% other information printed in the page headers. This command allows
%% the author to define a more concise list
%% of authors' names for this purpose.
%\renewcommand{\shortauthors}{Hartl and Bachl}

%%
%% The abstract is a short summary of the work to be presented in the
%% article.
%\begin{abstract}
%  A clear and well-documented \LaTeX\ document is presented as an
%  article formatted for publication by ACM in a conference proceedings
%  or journal publication. Based on the ``acmart'' document class, this
%  article presents and explains many of the common variations, as well
%  as many of the formatting elements an author may use in the
%  preparation of the documentation of their work.
%\end{abstract}

\newcommand{\codel}{CoDeL}

\settopmatter{printfolios=true}
\maketitle

\section{Introduction}

In the last decades various \gls{aqm} mechanisms have been proposed to minimize excessive standing queues in the Internet. One of the most influential recent efforts is \codel{} \cite{nichols_controlling_2012} whose goal is that the queuing delay at the bottleneck link is at least once under 5\;ms in a moving window of 100\;ms.

While it is important to keep the queuing delay constrained, it is also necessary to ensure that overly ``aggressive'' flows cannot benefit by ``stealing'' less aggressive flows' bandwidth. Thus researchers and engineers have developed \textit{\gls{fq}} mechanisms \cite{shreedhar_efficient_1996,dumazet_pkt_sched:_2013} to isolate different flows' queues (\textit{flow queuing}) so that for example a delay-sensitive live video call cannot be impaired by a concurrent bulk transfer which takes all the available bandwidth.

Recent approaches have tried to combine \gls{aqm} with \gls{fq}. \cite{taht_flow_2018} developed \textit{fq\_codel}, a \gls{qdisc} that uses \gls{fq} and lets \codel{} manage each queue. \cite{hoiland-jorgensen_piece_2018} expand upon this and create the \textit{cake} \gls{qdisc} that also adds features such as not only per-flow queuing but also per-host queuing for even increased fairness. Furthermore they also include bandwidth shaping into their solution and aim to create one \gls{qdisc} that is easy to configure and can be easily deployed on home routers and offers all features in one solution.

While we do not want to make statements about the general performance of \codel{}, we show that fq\_codel and cake do not optimally use available bandwidth in common network configurations for common \gls{cca}s. This becomes especially prevalent for links with a high bandwidth or a large \gls{rtt} but is already noticeable for common scenarios, such as a link with 100\;Mbit/s and an \gls{rtt} of 50\;ms.

We show that the impaired performance is a result of keeping the queuing delay under 5\;ms, which hinders \glspl{cca}s such as Reno or Cubic from reaching maximum throughput. The problem is that these \glspl{qdisc} aim to keep the delay under the threshold no matter the effect on throughput and do not take the congestion control of the flow into account.

As a remedy, we conceive an \gls{aqm} mechanism that explicitly measures the behavior of the \gls{cc} of a flow and dynamically changes the buffer so that 
\begin{enumerate}[wide, labelwidth=!, labelindent=0pt]
\item link utilization is optimized and
\item queuing delay is kept at a minimum that is required to achieve optimum throughput considering the \gls{cc}
\end{enumerate}

Finally, we develop a prototype of our \gls{qdisc} and show that it can achieve these objectives for the most common loss-based \gls{cca}s, Reno \cite {jacobson_congestion_1988} and Cubic \cite{ha_cubic:_2008}. These \glspl{cca} work by continuously increasing the number of bytes that are allowed to be in the network if no packet loss is experienced and by sharply decreasing this number if a packet is lost. With Cubic being the default \gls{cc} in all major OS \todo{citation needed}, we especially emphasize our evaluation on improving its performance. Contrasting to the aforementioned \glspl{cca}, recently proposed BBR \cite{cardwell_bbr:_2016} does not continuously increase and then sharply decrease when packets are lost but instead uses periodic measurements to estimate the available bandwidth as well as the minimal \gls{rtt} and then tries to stay at this point of optimal bandwidth and minimal delay. BBR is thus considered to be \textit{model-based}. We demonstrate that our \gls{qdisc} also behaves well in interaction with \glspl{cca} BBR.

\section{Concept}

\begin{algorithm}
\caption{Procedure which determines when to synchronize the neural network weights from the global weights in case of congestion control. We assume that the window at time step $t$ is stored in a list $l_\text{windows}$ and that the number of bytes sent in each time step is stored in $l_\text{sent}$. Alternatively to this procedure one can simply synchronize the weights after each $t_\text{max}$ actions where $t_\text{max}$ is an arbitrarily chosen integer greater than 0.}
\label{alg:update}
\begin{algorithmic}[1]%
%\algrestore{palalg}
\Function{updateWeights?}{}
	\State $i \gets 0$
	\While{$i < \#(l_\text{actions})$}
		\State $w\gets \left\lfloor l_\text{windows}[i] + l_\text{actions}[i] \right\rfloor $
		\For{$j\gets i..\#(l_\text{actions})-1$}
			\State $w\gets \text{w} - l_\text{sent}[j]$
			\If{$w\leq 0 \land i+1 \geq 10$} \Comment{at least 10}
				\State $i \gets j$
				\State \Break
			\EndIf
		\EndFor
		\If{$i = \#(l_\text{windows})-1$}
			\State \Return \True
		\EndIf
	\EndWhile
	\State \Return \False
\EndFunction
\end{algorithmic}
\end{algorithm}

\section*{Acknowledgements}
We thank Gernot Vormayr for providing us with the \texttt{py-virtnet} toolkit for building virtual networks.

\bibliographystyle{ACM-Reference-Format}
\bibliography{bibliography}

\end{document}

### 2019-10-11 tex/traq_illustration.tex
Add tex

\documentclass{standalone}
\usepackage[dvipsnames]{xcolor}
\usepackage{tikz}
\usepackage{fix-cm}
\setlength{\parskip}{0pt}
\setlength{\baselineskip}{0pt}
%\usetikzlibrary{fit}
\usetikzlibrary{calc}
\usetikzlibrary{decorations.markings}
\usetikzlibrary{decorations.shapes}
\usetikzlibrary{decorations.pathreplacing}
\usetikzlibrary{intersections}
\tikzset{%
  from end of path/.style={
    insert path={
      \pgfextra{%
        \expandafter\pgfprocesspathextractpoints%
          \csname tikz@intersect@path@name@#1\endcsname%
        \pgfpointlastonpath%
        \pgfgetlastxy\lastx\lasty
      }
      (\lastx,\lasty)
}}}
\tikzset{-dot-/.style={decoration={
  markings,
  mark=at position #1 with {\fill[color=red] circle (0.5mm);}},postaction={decorate}}} %%% in this line added a ;

\begin{document}
\begin{tikzpicture}
\draw[name path=A] (0,0.5)-- ++(0.5, 0.5);
\draw[from end of path=A, name path=B, -dot-=0, -dot-=0.5, -dot-=1] -- ++(0.25,0);
\draw[from end of path=B, name path=C] ++(0,-0.5mm) -- ++(0,-0.95cm) -- ++(1,1);
\draw[from end of path=C, name path=D, -dot-=0, -dot-=0.5, -dot-=1] -- ++(0.25,0);
\draw[from end of path=D, name path=D] ++(0,-0.5mm) -- ++(0,-0.95cm) -- ++(0.5,0.5);
\draw[draw,decorate,decoration={brace,mirror}] (0.75,0) -- (1.75,0) node [inner sep=0, outer sep=0,midway,below,yshift=-1.25mm] {\fontsize{3pt}{0}\selectfont interval};
\draw[densely dotted] (1.75,0) -- (1.75,0.95);
\end{tikzpicture}
\end{document}

### 2019-10-14 average_rtt.py
Add plotting scripts

@@ -1,0 +1,11 @@
+#!/usr/bin/env python3
+import statistics
+import sys
+
+all_lines = sys.stdin.readlines()
+# print("all_lines", all_lines)
+all_lines = [float(line.split("\t")[-1]) for line in all_lines]
+
+# print("all_lines", all_lines)
+print("mean rtt")
+print(statistics.mean(all_lines))

### 2019-10-14 plot_rtt_and_bandwidth.py

@@ -1,0 +1,70 @@
+#!/usr/bin/env python3
+
+import matplotlib.pyplot as plt
+import sys
+import os
+import subprocess
+plt.rcParams["font.family"] = "serif"
+
+interval = 1
+pcap_file = sys.argv[1]
+assert pcap_file.startswith("sender_")
+
+rtt_command = f"tshark -Y tcp.srcport!=60000&&tcp.analysis.ack_rtt!=0 -r pcaps/{pcap_file} -Tfields -e frame.time_relative -e tcp.analysis.ack_rtt"
+
+receiver_pcap = 'receiver_'+('_'.join(pcap_file.split('_')[1:]))
+
+packets_command = f"tshark -Y tcp.srcport==60000 -r pcaps/{receiver_pcap} -q -z io,stat,{interval},tcp.srcport==60000"
+
+bytes_command = f"tshark -r pcaps/{receiver_pcap} -q -z io,stat,{interval},SUM(ip.len)ip.len&&tcp.srcport==60000"
+
+rtt_out = subprocess.run(rtt_command.split(), check=True, stdout=subprocess.PIPE, stderr=subprocess.STDOUT)
+rtt_out = [item.split("\t") for item in rtt_out.stdout.decode("utf-8").split("\n") if item!=""]
+rtt_results = [(float(item[0]), 1000*float(item[1])) for item in rtt_out]
+# print("rtt_results", rtt_results[:100])
+
+# packets_out = subprocess.run(packets_command.split(), check=True, stdout=subprocess.PIPE, stderr=subprocess.STDOUT)
+# packets_out = [item for item in packets_out.stdout.decode("utf-8").split("\n")[12:-3] if item!=""]
+# packets_out = [[subitem for subitem in item.split("|") if subitem!=''] for item in packets_out]
+# packets_out = [[item[0].split("<>"), *item[1:]] for item in packets_out]
+# packets_results = [(float(item[0][0]), float(item[0][1]), float(item[1])) for item in packets_out]
+# # print("packets_results", packets_results)
+
+bytes_out = subprocess.run(bytes_command.split(), check=True, stdout=subprocess.PIPE, stderr=subprocess.STDOUT)
+bytes_out = [item for item in bytes_out.stdout.decode("utf-8").split("\n")[12:-3] if item!=""]
+bytes_out = [[subitem for subitem in item.split("|") if subitem!=''] for item in bytes_out]
[…]

### 2019-10-14 test.py

@@ -11,6 +11,8 @@ import math
 import signal
 sys.path.insert(0, os.path.dirname(os.path.abspath(os.path.dirname(__file__))))
 
+start_time = int(time.time() * 1000)
+
 import subprocess
 # import matplotlib
 # matplotlib.use("Agg")
@@ -157,8 +159,8 @@ def run(vnet):
 		server_popen = hosts[1].Popen("iperf3 -V -4 -s".split(" "), stdout=subprocess.PIPE, stderr=subprocess.PIPE)
 
 		os.makedirs("pcaps", exist_ok=True)
-		tcpdump_sender_popen = hosts[0].Popen(f"/usr/sbin/tcpdump -s {opt.bytes_to_capture} -i eth0 -w pcaps/sender_{opt.qdisc}_{opt.cc}_{opt.delay_to_add}_{opt.rate}_{opt.time}_{opt.change}.pcap tcp port {opt.cport} && port 5201".split(" "), stdout=subprocess.PIPE, stderr=subprocess.PIPE)
-		tcpdump_receiver_popen = hosts[1].Popen(f"/usr/sbin/tcpdump -s {opt.bytes_to_capture} -i eth0 -w pcaps/receiver_{opt.qdisc}_{opt.cc}_{opt.delay_to_add}_{opt.rate}_{opt.time}_{opt.change}.pcap tcp port {opt.cport} && port 5201".split(" "), stdout=subprocess.PIPE, stderr=subprocess.PIPE)
+		tcpdump_sender_popen = hosts[0].Popen(f"/usr/sbin/tcpdump -s {opt.bytes_to_capture} -i eth0 -w pcaps/sender_{opt.qdisc}_{opt.cc}_{opt.delay_to_add}_{opt.rate}_{opt.time}_{opt.change}_{start_time}.pcap tcp port {opt.cport} && port 5201".split(" "), stdout=subprocess.PIPE, stderr=subprocess.PIPE)
+		tcpdump_receiver_popen = hosts[1].Popen(f"/usr/sbin/tcpdump -s {opt.bytes_to_capture} -i eth0 -w pcaps/receiver_{opt.qdisc}_{opt.cc}_{opt.delay_to_add}_{opt.rate}_{opt.time}_{opt.change}_{start_time}.pcap tcp port {opt.cport} && port 5201".split(" "), stdout=subprocess.PIPE, stderr=subprocess.PIPE)
 
 		# tcpdump_switch_popens = []
 		# for interface_name in switch.interfaces.keys():

### 2019-10-17 tex/paper.tex
Add plotting scripts
Finish draft of paper

@@
 \def\tempYes{yes}
-\documentclass[10pt,sigconf,letterpaper,dvipsnames\ifx\removeHeaders\tempYes ,nonacm\fi]{acmart}
+\def\tempNo{no}
+\documentclass[10pt,sigconf,letterpaper,dvipsnames\ifx\removeHeaders\tempYes ,nonacm\fi\ifx\removeHeaders\tempNo ,anonymous\fi]{acmart}
 \acmYear{2019}
@@
 
-\usepackage{multirow}
-\usepackage{blindtext}
-\usepackage{soul}
-\usepackage[inline]{enumitem}
+%\usepackage{multirow}
+%\usepackage{blindtext}
+%\usepackage{soul}
+\usepackage{enumitem}
 \usepackage{algorithmicx}
@@
 \usepackage{algorithm}
+\usepackage{graphicx}
+\usepackage{subfig}
 
@@
 \newcommand{\algorithmautorefname}{Algorithm}
-
-%%
-%% \BibTeX command to typeset BibTeX logo in the docs
-\AtBeginDocument{%
-  \providecommand\BibTeX{{%
-    \normalfont B\kern-0.5em{\scshape i\kern-0.25em b}\kern-0.8em\TeX}}}
 
@@
 \newacronym{rtt}{RTT}{Round-Trip Time}
+\newacronym{gi}{GI}{Guard Interval}
+\newacronym{li}{LI}{Longest Interval}
+\newacronym{bdp}{BDP}{Bandwidth Delay Product}
 
@@
 While it is important to keep the queuing delay constrained, it is also necessary to ensure that overly ``aggressive'' flows cannot benefit by ``stealing'' less aggressive flows' bandwidth.
-Thus researchers and engineers have developed \textit{\gls{fq}} mechanisms \cite{shreedhar_efficient_1996,dumazet_pkt_sched:_2013} to isolate different flows' queues (\textit{flow queuing}) so that for example a delay-sensitive live video call cannot be impaired by a concurrent bulk transfer which takes all the available bandwidth.
+Thus researchers and engineers have developed \textit{\gls{fq}} mechanisms \cite{shreedhar_efficient_1996,dumazet_pkt_sched:_2013} to isolate different flows' queues so that for example a delay-sensitive live video call cannot be impaired by a concurrent bulk transfer which takes all the available bandwidth.
 
 Recent approaches have tried to combine \gls{aqm} with \gls{fq}.
-\cite{taht_flow_2018} developed \textit{fq\_codel}, a \gls{qdisc} that uses \gls{fq} and lets \codel{} manage each queue.
+\cite{taht_flow_2018} demonstrate \textit{fq\_codel}, a \gls{qdisc} that uses \gls{fq} and lets \codel{} manage each queue.
 \cite{hoiland-jorgensen_piece_2018} expand upon this and create the \textit{cake} \gls{qdisc} that also adds features such as not only per-flow queuing but also per-host queuing for even increased fairness.
@@
 
-While we do not want to make statements about the general performance of \codel{}, we show that fq\_codel and cake do not optimally use available bandwidth in common network configurations for common \gls{cca}s.
+While we do not want to make statements about the general performance of \codel{}, we show that fq\_codel and cake do not optimally use available bandwidth in common network configurations for common \glspl{cca}.
 This becomes especially prevalent for links with a high bandwidth or a large \gls{rtt} but is already noticeable for common scenarios, such as a link with 100\;Mbit/s and an \gls{rtt} of 50\;ms.
 
-We show that the impaired performance is a result of keeping the queuing delay under 5\;ms, which hinders \glspl{cca}s such as Reno or Cubic from reaching maximum throughput.
-The problem is that these \glspl{qdisc} aim to keep the delay under the threshold no matter the effect on throughput and do not take the congestion control of the flow into account.
+We show that the impaired performance is a result of keeping the queuing delay under 5\;ms, which hinders \glspl{cca} such as Reno or Cubic from reaching maximum throughput.
+The problem is that these \glspl{qdisc} aim to keep the delay under the threshold no matter the effect on throughput and do not take the \gls{cc} of the flow into account.
 
 As a remedy, we conceive an \gls{aqm} mechanism that explicitly measures the behavior of the \gls{cc} of a flow and dynamically changes the buffer so that 
-\begin{enumerate}[wide, labelwidth=!, labelindent=0pt]
+\begin{enumerate}[topsep=0pt,wide,labelwidth=!,labelindent=0pt]
 \item link utilization is optimized and
-\item queuing delay is kept at a minimum that is required to achieve optimum throughput considering the \gls{cc}
+\item queuing delay is kept at the minimum that is required to achieve optimum throughput considering the \gls{cc}.
 \end{enumerate}
-
-Finally, we develop a prototype of our \gls{qdisc} and show that it can achieve these objectives for the most common loss-based \gls{cca}s, Reno \cite {jacobson_congestion_1988} and Cubic \cite{ha_cubic:_2008}.
-These \glspl{cca} work by continuously increasing the number of bytes that are allowed to be in the network if no packet loss is experienced and by sharply decreasing this number if a packet is lost.
-With Cubic being the default \gls{cc} in all major OS \todo{citation needed}, we especially emphasize our evaluation on improving its performance.
+Finally, we develop a prototype of our \gls{qdisc} and show that it can achieve these objectives for the most common loss-based \glspl{cca}, Reno \cite {jacobson_congestion_1988} and Cubic \cite{ha_cubic:_2008}.
+These \glspl{cca} work by continuously increasing the number of bytes that are allowed to be in the network (congestion window) if no packet loss is experienced and by sharply decreasing this number if a packet is lost (multiplicative decrease).
+With Cubic being the default \gls{cc} in all major OS \todo{citation maybe needed}, we especially emphasize our evaluation on improving its performance.
 Contrasting to the aforementioned \glspl{cca}, recently proposed BBR \cite{cardwell_bbr:_2016} does not continuously increase and then sharply decrease when packets are lost but instead uses periodic measurements to estimate the available bandwidth as well as the minimal \gls{rtt} and then tries to stay at this point of optimal bandwidth and minimal delay.
 BBR is thus considered to be \textit{model-based}.
-We demonstrate that our \gls{qdisc} also behaves well in interaction with \glspl{cca} BBR.
+We demonstrate that our \gls{qdisc} also behaves well in interaction with BBR~v1.
 
@@
 
-\begin{algorithm}
-\caption{Procedure which determines when to synchronize the neural network weights from the global weights in case of congestion control.
-We assume that the window at time step $t$ is stored in a list $l_\text{windows}$ and that the number of bytes sent in each time step is stored in $l_\text{sent}$.
-Alternatively to this procedure one can simply synchronize the weights after each $t_\text{max}$ actions where $t_\text{max}$ is an arbitrarily chosen integer greater than 0.}
-\label{alg:update}
+\begin{figure}[h]
+\includegraphics[width=\columnwidth]{figures/traq_illustration_too_little.pdf}
+\caption{If the buffer is too small, loss-based \glspl{cca} will not be able to fully utilize the link since they send too few data following the multiplicative decrease that occurs after packet loss.}
+\label{fig:tooLittle}
+\end{figure}
+
+\begin{figure}[h]
+\includegraphics[width=\columnwidth]{figures/traq_illustration_too_much.pdf}
+\caption{If the buffer is too large, loss-based \glspl{cca} will keep a unnecessary standing queue.
+This queue is unnecessary since it is not generally required to achieve full link utilization.
+The consequences are longer \glspl{rtt} which can impair application performance and fluidity as well as Quality of Experience.}
+\label{fig:tooMuch}
+\end{figure}
+
+Since the goal of our algorithm is to achieve optimal throughput irrespective of the \gls{cca}, we aim to avoid the scenario depicted in \autoref{fig:tooLittle}: Here the buffer is too small, meaning that the \glspl{cca} never manages to achieve full utilization and periodically underutilizes the link, leading to, for example, a user waiting longer for a software update to finish or a video game to download on their video game console.
+
+The other case we want to avoid is a persisting standing queue, as depicted in \autoref{fig:tooMuch}: In this case the buffer is too large meaning that optimal throughput is achieved but at the same time that an unnecessarily long standing queue is maintained.
+This oversized queue leads to \gls{rtt} being larger than necessary and can result in unresponsive applications.
+Furthermore, the buffer space has to be allocated in the bottleneck device and keeps the memory from serving a more useful purpose.
+
+These considerations lead us to designing an algorithm that dynamically measures the \gls{cca} and adapts the buffer size to reach our goal of maximum throughput and minimal delay.
+The basic functionality of this algorithm would be to
+\begin{enumerate}[topsep=0pt,wide,labelwidth=!,labelindent=0pt]
+\item observe the \gls{cca} for a certain measurement period and then
+\item increase or decrease the buffer size depending on whether the scenario of \autoref{fig:tooLittle} or \autoref{fig:tooMuch} is true. 
+\end{enumerate}
+The tricky thing about this is to define the measurement period: As \autoref{fig:tooLittle} and \autoref{fig:tooMuch} show, we would like to use the measurement period that is the longest period between two packet losses.
+If this period is observed, it is possible to compute the standing queue as simply the minimum queue observed in this interval.
+For example, if the minimum queue is 3 packets it means, that there is a standing queue of 3 packets and the algorithm will reduce buffer size by 3 to eliminate the unnecessary queuing.
+On the other side, if the buffer is too small, the algorithm computes how long the link was underutilized and how many packets could have been transmitted during the idle period.
+The buffer size is then increased by this number of packets.
+
+The problem is now the following: If the measurement period is mistakenly not the longest interval shown in the figure, but the interval between two adjacent packet losses, the algorithm would assume an enormous standing queue and drastically reduce the buffer to be close to zero.
+This would then lead to severe underutilization afterwards.
+
+The solution is thus to use a \textit{\gls{gi}}: The algorithm always waits for a specified minimum amount of time, picks the longest interval without packet loss in this guard interval and then applies its logic to reduce the standing queue on this \textit{\gls{li}}.
+We dynamically compute the \gls{gi} as a multiple (parameter of our algorithm) of the previous \gls{li} (\autoref{fig:intervals}):
+
+\begin{figure}[h]
+\includegraphics[width=\columnwidth]{{{figures/traq_intervals}}}
+\caption{Illustration of the interval mechanism used by our \gls{qdisc}.
+The multiplier used is $\frac{1}{2}$.
+Each \textit{guard interval} (GI) contains one or more intervals of which the longest one is considered the \textit{longest interval} (LI).
+The \gls{gi} is always at least the length of the previous \gls{li} times the multiplier.
+A \gls{gi} ends at the first packet loss after that is after the previous \gls{li} times the multiplier.}
+\label{fig:intervals}
+\end{figure}
+
+For example, we choose the multiplier to be $\frac{1}{2}$.
+Now, if the \gls{li} is in the previous \gls{gi} was 100\,ms, then the next \gls{gi} is at least $\frac{1}{2} 100\,ms$.
+During the \gls{gi} we monitor all intervals and if the current interval is the longest one, it becomes the new \gls{li}.
+The \gls{gi} ends when the first packet loss occurs after the end of the \gls{gi}.
+
+\begin{algorithm}[h]
+\caption{Procedure that is executed when a new packet is received to be enqueued.}
+\label{alg:enqueue}
 \begin{algorithmic}[1]%
 %\algrestore{palalg}
-\Function{updateWeights?}{}
-	\State $i \gets 0$
-	\While{$i < \#(l_\text{actions})$}
-		\State $w\gets \left\lfloor l_\text{windows}[i] + l_\text{actions}[i] \right\rfloor $
-		\For{$j\gets i..\#(l_\text{actions})-1$}
-			\State $w\gets \text{w} - l_\text{sent}[j]$
-			\If{$w\leq 0 \land i+1 \geq 10$} \Comment{at least 10}
-				\State $i \gets j$
-				\State \Break
+\Function{enqueue}{new\_packet}
+	\If{queue is full}
+		\If{link was idle during this interval $\land$ the buffer wasn't already enlarged in this interval $\land$ this is not the first interval of this flow}
+			\State buffer\_size $\gets$ buffer\_size + (packets\_transmitted\_in\_current\_interval / time\_active $\times$ time\_idle)
+			\State add new\_packet to queue
+		\ElsIf{the buffer was enlarged in this interval $\lor$ this is the first interval of this flow}
+			\State start a new interval
+			\State start a new \gls{gi}
+			\State drop new\_packet
+		\ElsIf{current\_time $\geq$ end\_the\_current\_\gls{gi}}
+			\If{there was a standing queue in the \gls{li} of the \gls{gi}}
+				\State buffer\_size $\gets$ buffer\_size - standing\_queue
 			\EndIf
-		\EndFor
-		\If{$i = \#(l_\text{windows})-1$}
-			\State \Return \True
+			\State start a new interval
+			\State start a new \gls{gi}
+			\State drop new\_packet
+		\Else
+			\State start a new interval
+			\State drop new\_packet
 		\EndIf
-	\EndWhile
-	\State \Return \False
+	\Else
+		\State add new\_packet to queue
+	\EndIf
 \EndFunction
@@
 \end{algorithm}
+Almost all of the logic of our algorithm is executed when a new packet is received to be enqueued. The full algorithm is depicted as \autoref{alg:enqueue}.
+
+\subsection{Choice of the multiplier}
+
+The purpose of the multiplier, which is used to multiply the \gls{li} of previous \gls{gi} to define the new \gls{gi}, is to prevent that the short gap between two subsequent packet losses is is used as the interval to determine whether the buffer size has to be adapted.
+This means that the multiplier must be large enough so that it allows to ``jump'' over the packet losses that occur when the buffer is full.
+For Reno and Cubic, the loss-free period is very large compared to the period during which packet loss occurs.
+Specifically, the period with packet loss has the length of one \gls{rtt} since Reno and Cubic reduce their sending rate as soon as packet loss occurs, thus ending the packet loss.
+This means that the multiplier can be very small in order to work for these \glspl{cca}.
+We conducted experiments with the multiplier being $0.5$ and have never encountered the problem that the interval between adjacent packet losses is considered the longest interval and that the buffer is cut to virtually nothing because a standing queue is assumed.
+However, BBR~v1 uses a completely different approach compared to Reno and Cubic to probe for bandwidth: It periodically increases its sending rate and then reduces it again to see if it can achieve more throughput.
+This happens with the following pattern: $[\frac{5}{4},\frac{3}{4},1,1,1,1,1,1]$.
+The sending rate is increased to probe for more bandwidth, then decreased to reduce the standing queue potentially formed and after that BBR is simply keeping the sending rate to match the estimated bandwidth.
+Each of these eight phases lasts one \gls{rtt}.
+BBR uses a random cyclic permutation of this pattern with the only condition being that it cannot start with $\frac{3}{4}$.
+The following pattern could thus occur: $[1,1,1,1,1,1,\frac{5}{4},\frac{3}{4},\frac{5}{4},\frac{3}{4},1,1,1,1,1,1]$.
+If the \gls{gi} ends after the first 6 phases and also the \gls{li} was these six phase, with a factor of $0.5$, the next \gls{gi} would thus end at the end of phase 8.
+Then, the \gls{li} would be 8th cycle.
+The next \gls{gi} would then be of the duration of $0.5$ phases and would then end in the middle of cycle 9.
+This would be erroneous because this is a probing phase during which packet loss occurs.
+The aforementioned problem of erroneously and significantly reducing the buffer could thus occur.
+This means that for BBR~v1 the multiplier must be larger than 1 since it can happen that non-probing phases are followed by probing phases of the same length.
+From these considerations it follows that for our experiments we generally use a multiplier of $1.25$.
+The behavior is different for BBR~v2 which drastically changed the probing phase and now consists of a random 2-3\,s phase of constant sending rate (cruising) followed by a quick probing phase followed by a decrease phase \cite{cardwell_bbr_2019-1,cardwell_bbr_2019}.
+We thus conjecture that a multiplier of $0.5$ is sufficient for BBR~v2.
+However, we could not finally confirm this since BBR~v2 is still not completely finished.
+
+\subsection{Maximum buffer increase}
+
+Another potential problem is the following: If the link speed suddenly increases drastically, our algorithm would sharply increase the buffer size.
+In this case it can happen that the buffer becomes much too large.
+We thus limit the maximum buffer increase to be 2 times the previous buffer size.
+This parameter is only relevant in the case of a sudden increase of available bandwidth or if a flow's throughput is application limited and not during stable state behavior.
+
+\subsection{Maximum \gls{gi}}
+
+One more challenge is posed by the following consideration: If a delay-based \gls{cca} is in its stable state, no packet loss should occur.
+The \gls{li} could thus become very large (like tens of seconds) and also the next \gls{gi} would be extremely large.
+This can be a problem if a delay-based \gls{cca} like Vegas \cite{brakmo_tcp_1995} is used for a long flow and then after several seconds, the bandwidth drastically lowers.
+Then suddenly packet loss could occur but the \gls{gi} would be so long that nothing would happen for a long time.
+To counter this, we add a parameter that specifies the maximum duration of the \gls{gi}.
+We set this parameter to 1\,s.
+This means that we also assume that no flows are handled by our \gls{qdisc} which have an \gls{rtt} larger than 1\,s.
+Another possibility for setting this parameter is to set it to a multiple of the \gls{rtt}, for example 2, 4 or 8 times the most recent \gls{rtt}.
+\gls{rtt} could be measured using TCP timestamps \cite{borman_tcp_2014} in case of TCP flows or the spin bit \cite{kuhlewind_quic_2018} for QUIC.
+
+\section{Implementation}
+
+We implement our \gls{qdisc} as an extension of the ``fq'' \gls{qdisc} \cite{dumazet_pkt_sched:_2013}.
+This means that when using our \gls{qdisc} also all features offered by fq are available.
+We add three configuration parameters to the \gls{qdisc}: the multiplier (default 1.25), the maximum buffer increase (default 2.0) and the maximum \gls{gi} duration (default 1.0\,s).
+We make the source code of our implementation freely available to enable reproducibility and encourage experimentation: \ifx\removeHeaders\tempYes\url{https://github.com/CN-TU/traq}\else{}[removed for anonymity]\fi .
+
+\section{Evaluation}
+
+We evaluate our \gls{qdisc} using the \ifx\removeHeaders\tempYes{}\texttt{py-virtnet}\else{}[removed for anonymity]\fi{} toolkit to build a virtual network using Linux's network namespaces.
+We initialize the buffer size for our \gls{qdisc} to 100 packets per flow like in fq.
+
+\begin{figure}[h]
+\subfloat[Throughput of the fq \gls{qdisc}\label{fig:fqThroughput}
+]{\includegraphics[width=\columnwidth]{{{../plots/throughput_1_fq_cubic_10_20_60_0.5_bw_1571240937211.pcap}}}}{}
+\subfloat[Delay of the fq \gls{qdisc}.\label{fig:fqDelay}
+]{\includegraphics[width=\columnwidth]{{{../plots/rtt_1_fq_cubic_10_20_60_0.5_bw_1571240937211.pcap}}}}{}
+\caption{Throughput (\ref{fig:fqThroughput}) and delay (\ref{fig:fqDelay}) of a Cubic flow that is managed by fq at the bottleneck.
+The initial bandwidth is 20\,Mbit/s but it is halved after 30\,s.
+The delay is 10\,ms.
+It is clear that there is a standing queue and after the bandwidth halves, the minimum delay is 100\,ms even though 10\,ms would be the possible.}
+\label{fig:fq}
+\end{figure}
+
+\begin{figure}[h]
+\subfloat[Throughput of our \gls{qdisc}\label{fig:cnThroughput}
+]{\includegraphics[width=\columnwidth]{{{../plots/throughput_1_cn_cubic_10_20_60_0.5_bw_1571241047054.pcap}}}}{}
+\subfloat[Delay of our \gls{qdisc}.\label{fig:cnDelay}
+]{\includegraphics[width=\columnwidth]{{{../plots/rtt_1_cn_cubic_10_20_60_0.5_bw_1571241047054.pcap}}}}{}
+\caption{Throughput (\ref{fig:cnThroughput}) and delay (\ref{fig:cnDelay}) of a Cubic flow that is managed by fq at the bottleneck.
+The initial bandwidth is 20\,Mbit/s but it is halved after 30\,s.
+The delay is 10\,ms.
+Our \gls{qdisc} manages to keep the queuing delay minimal while achieving the same throughput as with an oversized buffer.}
+\label{fig:cn}
+\end{figure}
+
+\begin{figure}[h]
+\subfloat[Throughput of our \gls{qdisc}\label{fig:cnThroughput2}
+]{\includegraphics[width=\columnwidth]{{{../plots/throughput_1_cn_cubic_10_20_60_2.0_bw_1571241151290.pcap}}}}{}
+\subfloat[Delay of our \gls{qdisc}.\label{fig:cnDelay2}
+]{\includegraphics[width=\columnwidth]{{{../plots/rtt_1_cn_cubic_10_20_60_2.0_bw_1571241151290.pcap}}}}{}
+\caption{Throughput (\ref{fig:cnThroughput2}) and delay (\ref{fig:cnDelay2}) of a Cubic flow that is managed by fq at the bottleneck.
+The initial bandwidth is 20\,Mbit/s but it is doubled after 30\,s.
+The delay is 10\,ms.
+Also a sudden large increase of bandwidth is handled well by our \gls{qdisc}.}
+\label{fig:cn2}
+\end{figure}
+
+First we evaluate if our \gls{qdisc} is able to maintain a small buffer like it is necessary in case of small \glspl{bdp}.
+As can be seen in \autoref{fig:fq} the regular fq qdisc maintains a standing queue and unnecessarily leads to a significant increase in \gls{rtt} (100\,ms minimum, when the real minimum is 10\,ms).
+In contrast, \autoref{fig:cn} maintains the full throughput while at the same time keeping the delay minimal and not keeping a standing queue.
+Also, when drastically changing the bandwidth by halving it (\autoref{fig:cn}) or doubling it (\autoref{fig:cn2}) our \gls{qdisc} rapidly adapts to the new conditions and returns to the optimum state.
+
+\begin{figure}[h]
+\subfloat[Throughput of our \gls{qdisc}\label{fig:fqCodelThroughputLong}
+]{\includegraphics[width=\columnwidth]{{{../plots/throughput_1_fq_codel_cubic_100_100_240_1.0_bw_1571244312838.pcap}}}}{}
+\subfloat[Delay of our \gls{qdisc}.\label{fig:fqCodelDelayLong}
+]{\includegraphics[width=\columnwidth]{{{../plots/rtt_1_fq_codel_cubic_100_100_240_1.0_bw_1571244312838.pcap}}}}{}
+\caption{Throughput (\ref{fig:fqCodelThroughputLong}) and delay (\ref{fig:fqCodelDelayLong}) of a Cubic flow that is managed by fq\_codel at the bottleneck.
+The bandwidth is 100\,Mbit/s.
+The delay is 100\,ms.
+The average \gls{rtt} is 101.36\,ms.
+Total throughput is 2615\,MB.}
+\label{fig:fqCodel}
+\end{figure}
+
+\begin{figure}[h]
+\subfloat[Throughput of our \gls{qdisc}\label{fig:cnThroughputLong}
+]{\includegraphics[width=\columnwidth]{{{../plots/throughput_1_cn_cubic_100_100_240_1.0_bw_1571244036720.pcap}}}}{}
+\subfloat[Delay of our \gls{qdisc}.\label{fig:cnDelayLong}
+]{\includegraphics[width=\columnwidth]{{{../plots/rtt_1_cn_cubic_100_100_240_1.0_bw_1571244036720.pcap}}}}{}
+\caption{Throughput (\ref{fig:cnThroughputLong}) and delay (\ref{fig:cnDelayLong}) of a Cubic flow that is managed by our \gls{qdisc} at the bottleneck.
+The bandwidth is 100\,Mbit/s.
+The delay is 100\,ms.
+The average \gls{rtt} is 129.45\,ms.
+Total throughput is 2893\,MB.}
+\label{fig:cnLong}
+\end{figure}
+
+Next we compare our \gls{qdisc} against fq\_codel.
+Figure\autoref{fig:fqCodelDelayLong} shows that fq\_codel keeps the queuing delay under 5\,ms.
+Figure\autoref{fig:fqCodelThroughputLong} shows that keeping the queuing delay that low is detrimental for achieving full throughput when using the Cubic \gls{cca}.
+With our \gls{qdisc} we achieve throughput that is more than 10\% higher overall, while only slightly increasing the average delay \todo{More systematic evaluation}.
+In addition to Cubic, we performed experiments with Reno.
+Here, over a 240\,s flow of 100\,Mbit/s with a delay of 50\,ms, our \gls{qdisc} achieves more than 20\% higher throughput than fq\_codel.
+The average \gls{rtt} for fq\_codel is 51.13\,ms while it is 96.59\,ms for our \gls{qdisc}.
+
+Besides fq\_codel possibly leading to lower throughput as we have shown, another problem is that for links with a small \gls{bdp} fq\_codel can keep a standing queue akin to fq since fq\_codel only wants queuing delay to fall below 5\,ms once every 100\,ms.
+For example, on a link with a base \gls{rtt} of 5\,ms, fq\_codel would accept a permanent standing queue of 4.9\,ms, leading to overall \gls{rtt} being doubled.
+
+We also ran experiments with the cake \gls{qdisc}. However, results were very similar to those of fq\_codel, which is no surprise since cake is based upon fq\_codel.
+
+\begin{figure}[h]
+\subfloat[Throughput of our \gls{qdisc}\label{fig:cnThroughputBBR}
+]{\includegraphics[width=\columnwidth]{{{../plots/throughput_1_cn_bbr_10_50_60_0.5_delay_1571288916582.pcap}}}}{}
+\subfloat[Delay of our \gls{qdisc}.\label{fig:cnDelayBBR}
+]{\includegraphics[width=\columnwidth]{{{../plots/rtt_1_cn_bbr_10_50_60_0.5_delay_1571288916582.pcap}}}}{}
+\caption{Throughput (\ref{fig:cnThroughputBBR}) and delay (\ref{fig:cnDelayBBR}) of a BBR~v1 flow that is managed by our \gls{qdisc} at the bottleneck.
+The bandwidth is 50\,Mbit/s.
+The delay is 10\,ms and it is cut by half after 30\,s.}
+\label{fig:cnBBR}
+\end{figure}
+
+Also for BBR~v1, throughput quickly reaches its maximum while no standing queue is forming \autoref{fig:cnBBR}.
+
+\section{Discussion}
+
+Our goal was to design a \gls{qdisc} which achieves maximum throughput while keeping the queuing delay as small as possible.
+The results of the experiments, which we perform with a prototype of our \gls{qdisc}, show that there are certain common scenarios in which current state-of-the-art \glspl{qdisc} like fq, fq\_codel and cake fall short of the optimum throughput by a significant margin while our approach succeeds in leading to full utilization of the link at the bottleneck.
+Furthermore, the aforementioned \glspl{qdisc} also suffer from standing delays is certain scenarios which we can also mitigate with our approach.
+
+We see this work as a first step towards adaptive queuing at bottlenecks and think that a promising direction for future work might be to explore the use of reinforcement learning.
+Such an approach would fingerprint a flow and choose the buffer size accordingly, maximizing a chosen objective such as high throughput with minimal delay.
 
 \section*{Acknowledgements}
-We thank Gernot Vormayr for providing us with the \texttt{py-virtnet} toolkit for building virtual networks.
+We thank \ifx\removeHeaders\tempYes{}Gernot Vormayr\else{}[removed for anonymity]\fi{} for providing us with the \ifx\removeHeaders\tempYes{}\texttt{py-virtnet}\else{}[removed for anonymity]\fi{} toolkit for building virtual networks.
 

### 2019-10-17 plot_rtt_and_bandwidth.py

@@ -4,7 +4,9 @@ import matplotlib.pyplot as plt
 import sys
 import os
 import subprocess
+import statistics
 plt.rcParams["font.family"] = "serif"
+plt.rcParams['pdf.fonttype'] = 42
 
 interval = 1
 pcap_file = sys.argv[1]
@@ -14,14 +16,28 @@ rtt_command = f"tshark -Y tcp.srcport!=60000&&tcp.analysis.ack_rtt!=0 -r pcaps/{
 
 receiver_pcap = 'receiver_'+('_'.join(pcap_file.split('_')[1:]))
 
+retransmissions_command = f"tshark -Y tcp.srcport==60000&&tcp.analysis.retransmission -r pcaps/{pcap_file} -Tfields -e frame.time_relative"
+
 packets_command = f"tshark -Y tcp.srcport==60000 -r pcaps/{receiver_pcap} -q -z io,stat,{interval},tcp.srcport==60000"
 
 bytes_command = f"tshark -r pcaps/{receiver_pcap} -q -z io,stat,{interval},SUM(ip.len)ip.len&&tcp.srcport==60000"
 
+bytes_command_total = f"tshark -r pcaps/{receiver_pcap} -q -z io,stat,0,SUM(ip.len)ip.len&&tcp.srcport==60000"
+
 rtt_out = subprocess.run(rtt_command.split(), check=True, stdout=subprocess.PIPE, stderr=subprocess.STDOUT)
 rtt_out = [item.split("\t") for item in rtt_out.stdout.decode("utf-8").split("\n") if item!=""]
 rtt_results = [(float(item[0]), 1000*float(item[1])) for item in rtt_out]
+print("average rtt", statistics.mean(list(zip(*rtt_results))[1]))
 # print("rtt_results", rtt_results[:100])
+
+retransmissions_out = subprocess.run(retransmissions_command.split(), check=True, stdout=subprocess.PIPE, stderr=subprocess.STDOUT)
+retransmissions_out = [item for item in retransmissions_out.stdout.decode("utf-8").split("\n") if item!=""]
+retransmissions_results = [float(item) for item in retransmissions_out]
+print("total retransmissions", len(retransmissions_results))
+
+bytes_total_out = subprocess.run(bytes_command_total.split(), check=True, stdout=subprocess.PIPE, stderr=subprocess.STDOUT)
+bytes_total_out = bytes_total_out.stdout.decode("utf-8")
[…]

### 2019-10-17 sch_cn.c

@@ -1145,7 +1145,7 @@ static int cn_init(struct Qdisc *sch, struct nlattr *opt,
 	q->orphan_mask		= 1024 - 1;
 	q->low_rate_threshold	= 550000 / 8;
 	kernel_fpu_begin();
-	q->guard_interval = 0.5;
+	q->guard_interval = 1.25;
 	q->max_increase = 2.0;
 	// 1 second
 	q->max_monitoring_interval = 1*NANOSECONDS_IN_ONE_SECOND;

### 2019-10-17 tex/figures/traq_common_beginning.tex
Add tikz figures

--- 2019-10-11 tex/traq_illustration.tex
@@
 \usetikzlibrary{intersections}
+\usetikzlibrary{patterns}
 \tikzset{%
@@
   markings,
-  mark=at position #1 with {\fill[color=red] circle (0.5mm);}},postaction={decorate}}} %%% in this line added a ;
-
+  mark=at position #1 with {\fill[color=red] circle (0.5mm);}},postaction={decorate}}}  
+    \tikzset{
+        hatch distance/.store in=\hatchdistance,
+        hatch distance=10pt,
+        hatch thickness/.store in=\hatchthickness,
+        hatch thickness=2pt
+    }
 \begin{document}
 \begin{tikzpicture}
+
+    \makeatletter
+    \pgfdeclarepatternformonly[\hatchdistance,\hatchthickness]{flexible hatch}
+    {\pgfqpoint{0pt}{0pt}}
+    {\pgfqpoint{\hatchdistance}{\hatchdistance}}
+    {\pgfpoint{\hatchdistance-1pt}{\hatchdistance-1pt}}%
+    {
+        \pgfsetcolor{\tikz@pattern@color}
+        \pgfsetlinewidth{\hatchthickness}
+        \pgfpathmoveto{\pgfqpoint{0pt}{0pt}}
+        \pgfpathlineto{\pgfqpoint{\hatchdistance}{\hatchdistance}}
+        \pgfusepath{stroke}
+    }
+    \makeatother
+
 \draw[name path=A] (0,0.5)-- ++(0.5, 0.5);
 \draw[from end of path=A, name path=B, -dot-=0, -dot-=0.5, -dot-=1] -- ++(0.25,0);
+\node[align=center,font=\fontsize{3pt}{0}\selectfont,inner sep=0mm] at (0.5, 1.25) (lossLabel) {packet loss};
+\draw[->, line width=0.05mm] (lossLabel) -- (0.5,1.05);
 \draw[from end of path=B, name path=C] ++(0,-0.5mm) -- ++(0,-0.95cm) -- ++(1,1);
 \draw[from end of path=C, name path=D, -dot-=0, -dot-=0.5, -dot-=1] -- ++(0.25,0);
-\draw[from end of path=D, name path=D] ++(0,-0.5mm) -- ++(0,-0.95cm) -- ++(0.5,0.5);
-\draw[draw,decorate,decoration={brace,mirror}] (0.75,0) -- (1.75,0) node [inner sep=0, outer sep=0,midway,below,yshift=-1.25mm] {\fontsize{3pt}{0}\selectfont interval};
-\draw[densely dotted] (1.75,0) -- (1.75,0.95);
-\end{tikzpicture}
-\end{document}
+\draw[from end of path=D, name path=E] ++(0,-0.5mm) -- ++(0,-0.95cm) -- ++(0.5,0.5);

### 2019-10-17 tex/figures/traq_illustration_too_little.tex
Add tikz figures

\input{traq_common_beginning.tex}
%\draw[draw opacity=0,pattern=flexible hatch, pattern color=red, hatch distance=1mm, hatch thickness=0.01mm] (0.75,0) -- ++(0.5,0.5) -- (0.75,0.5) -- cycle;
\draw[draw opacity=0,fill=red, fill opacity=0.3] (0.75,0) -- ++(0.5,0.5) -- (0.75,0.5) -- cycle;
%\draw[name path=A] (0,-0.5)-- ++(2.5, 0);

\draw[draw,decorate,decoration={brace}] (0.75,0.5) -- (1.25,0.5) node [midway,above,align=center,font=\fontsize{3pt}{0}\selectfont] {under-\\ utilized\\ link};
%\draw[densely dotted] (0.75,0) -- (0.75,-0.5);

\draw[dashed] (0,0.5) -- ++(2.5,0);
\node[align=center,font=\fontsize{3pt}{0}\selectfont,inner sep=0mm] at (0.375, 0.25) (optimal) {optimal\\ window};
\draw[->, line width=0.05mm] (optimal) -- (0.375,0.5);

\input{traq_common_end.tex}

### 2019-10-17 tex/figures/traq_illustration_too_much.tex
Add tikz figures

\input{traq_common_beginning.tex}
%\fill[pattern=flexible hatch, pattern color=red, hatch distance=1mm, hatch thickness=0.01mm] (0,0) rectangle ++(2.5,-0.5);
\draw[draw opacity=0,fill=red,fill opacity=0.3] (0,0) rectangle ++(2.5,-0.5);
%\draw[name path=A] (0,-0.5)-- ++(2.5, 0);

\draw[draw,decorate,decoration={brace,mirror}] (0.75,0) -- (0.75,-0.5) node [midway,left,align=right,font=\fontsize{3pt}{0}\selectfont] {standing\\ queue};
%\draw[densely dotted] (0.75,0) -- (0.75,-0.5);

\draw[dashed] (0,-0.5) -- ++(2.5,0);
\node[align=center,font=\fontsize{3pt}{0}\selectfont,inner sep=0mm] at (1.95,-0.25) (optimal) {optimal\\ window};
\draw[->, line width=0.05mm] (optimal) -- ++(0,-0.25);

\input{traq_common_end.tex}

### 2019-10-17 tex/figures/traq_intervals.tex
Add tikz figures

--- 2019-10-17 tex/figures/traq_common_beginning.tex
@@
 \usetikzlibrary{patterns}
+\newcommand{\liy}{-0.3}
+\newcommand{\giy}{-0.6}
 \tikzset{%
@@
   mark=at position #1 with {\fill[color=red] circle (0.5mm);}},postaction={decorate}}}  
+\tikzset{-dot2-/.style={decoration={
+  markings,
+  mark=at position #1 with {\draw[color=black] (0,-0.25mm) -- (0,0.25mm); (0.25mm);}},postaction={decorate}}}
     \tikzset{
@@
 
-\draw[name path=A] (0,0.5)-- ++(0.5, 0.5);
+\draw[name path=A] (0,0)-- ++(1, 1);
 \draw[from end of path=A, name path=B, -dot-=0, -dot-=0.5, -dot-=1] -- ++(0.25,0);
-\node[align=center,font=\fontsize{3pt}{0}\selectfont,inner sep=0mm] at (0.5, 1.25) (lossLabel) {packet loss};
-\draw[->, line width=0.05mm] (lossLabel) -- (0.5,1.05);
+\node[align=center,font=\fontsize{3pt}{0}\selectfont,inner sep=0mm] at (1, 1.25) (lossLabel) {packet loss};
+\draw[->, line width=0.05mm] (lossLabel) -- (1,1.05);
 \draw[from end of path=B, name path=C] ++(0,-0.5mm) -- ++(0,-0.95cm) -- ++(1,1);
 \draw[from end of path=C, name path=D, -dot-=0, -dot-=0.5, -dot-=1] -- ++(0.25,0);
-\draw[from end of path=D, name path=E] ++(0,-0.5mm) -- ++(0,-0.95cm) -- ++(0.5,0.5);
+\draw[from end of path=D, name path=E, -dot-=1] ++(0,-0.5mm) -- ++(0,-0.95cm) -- ++(1,1);
+
+%\node[anchor=north west,align=left,font=\fontsize{3pt}{0}\selectfont,inner sep=0mm] at (0.5mm, -1mm) (intervalsLabel) {intervals};
+\draw[name path=F] (0,-0.25)-- ++(1, 0);
+\draw[from end of path=F, name path=G, -dot2-=0, -dot2-=0.5, -dot2-=1] -- ++(0.25,0);
+\draw[from end of path=G, name path=H] -- ++(1,0);
+\draw[from end of path=H, name path=I, -dot2-=0, -dot2-=0.5, -dot2-=1] -- ++(0.25,0);
+\draw[from end of path=I, name path=J] -- ++(1,0);
+
+\draw[draw,decorate,decoration={brace}] (0,-0.2) -- ++(3.5,0) node [,midway,above,font=\fontsize{3pt}{0}\selectfont] {intervals};
+
+\draw[draw,decorate,decoration={brace,mirror}] (0,\liy) -- ++(1,0) node [inner sep=0, outer sep=0,midway,below,yshift=-1.25mm,font=\fontsize{3pt}{0}\selectfont] {longest interval};
+\draw[draw,decorate,decoration={brace,mirror}] (1.25,\liy) -- ++(1,0) node [inner sep=0, outer sep=0,midway,below,yshift=-1.25mm,font=\fontsize{3pt}{0}\selectfont] {longest interval};
+\draw[draw,decorate,decoration={brace,mirror}] (2.5,\liy) -- ++(1,0) node [inner sep=0, outer sep=0,midway,below,yshift=-1.25mm,font=\fontsize{3pt}{0}\selectfont] {longest interval};
+
+\draw[draw,decorate,decoration={brace,mirror}] (0,\giy) -- ++(1,0) node [inner sep=0, outer sep=0,midway,below,yshift=-1.25mm,font=\fontsize{3pt}{0}\selectfont] {guard interval};
+\draw[draw,decorate,decoration={brace,mirror}] (1,\giy) -- ++(1.25,0) node [inner sep=0, outer sep=0,midway,below,yshift=-1.25mm,font=\fontsize{3pt}{0}\selectfont] {guard interval};
+\draw[draw,decorate,decoration={brace,mirror}] (2.25,\giy) -- ++(1.25,0) node [inner sep=0, outer sep=0,midway,below,yshift=-1.25mm,font=\fontsize{3pt}{0}\selectfont] {guard interval};
+
+\end{tikzpicture}
+\end{document}

### 2019-10-17 tex/paper.tex
Change font in figures to Linux Libertine

@@
 These \glspl{cca} work by continuously increasing the number of bytes that are allowed to be in the network (congestion window) if no packet loss is experienced and by sharply decreasing this number if a packet is lost (multiplicative decrease).
-With Cubic being the default \gls{cc} in all major OS \todo{citation maybe needed}, we especially emphasize our evaluation on improving its performance.
+With Cubic being the default \gls{cc} in all major OS, we especially emphasize our evaluation on improving its performance.
 Contrasting to the aforementioned \glspl{cca}, recently proposed BBR \cite{cardwell_bbr:_2016} does not continuously increase and then sharply decrease when packets are lost but instead uses periodic measurements to estimate the available bandwidth as well as the minimal \gls{rtt} and then tries to stay at this point of optimal bandwidth and minimal delay.
@@
 ]{\includegraphics[width=\columnwidth]{{{../plots/rtt_1_cn_cubic_10_20_60_0.5_bw_1571241047054.pcap}}}}{}
-\caption{Throughput (\ref{fig:cnThroughput}) and delay (\ref{fig:cnDelay}) of a Cubic flow that is managed by fq at the bottleneck.
+\caption{Throughput (\ref{fig:cnThroughput}) and delay (\ref{fig:cnDelay}) of a Cubic flow that is managed by our \gls{qdisc} at the bottleneck.
 The initial bandwidth is 20\,Mbit/s but it is halved after 30\,s.
@@
 ]{\includegraphics[width=\columnwidth]{{{../plots/rtt_1_cn_cubic_10_20_60_2.0_bw_1571241151290.pcap}}}}{}
-\caption{Throughput (\ref{fig:cnThroughput2}) and delay (\ref{fig:cnDelay2}) of a Cubic flow that is managed by fq at the bottleneck.
+\caption{Throughput (\ref{fig:cnThroughput2}) and delay (\ref{fig:cnDelay2}) of a Cubic flow that is managed by our \gls{qdisc} at the bottleneck.
 The initial bandwidth is 20\,Mbit/s but it is doubled after 30\,s.
@@
 \begin{figure}[h]
-\subfloat[Throughput of our \gls{qdisc}\label{fig:fqCodelThroughputLong}
+\subfloat[Throughput of the fq\_codel \gls{qdisc}\label{fig:fqCodelThroughputLong}
 ]{\includegraphics[width=\columnwidth]{{{../plots/throughput_1_fq_codel_cubic_100_100_240_1.0_bw_1571244312838.pcap}}}}{}
-\subfloat[Delay of our \gls{qdisc}.\label{fig:fqCodelDelayLong}
+\subfloat[Delay of the fq\_codel \gls{qdisc}.\label{fig:fqCodelDelayLong}
 ]{\includegraphics[width=\columnwidth]{{{../plots/rtt_1_fq_codel_cubic_100_100_240_1.0_bw_1571244312838.pcap}}}}{}

### 2019-10-17 tex/paper.tex
Improve paper a little bit

@@
 \def\tempNo{no}
-\documentclass[10pt,sigconf,letterpaper,dvipsnames\ifx\removeHeaders\tempYes ,nonacm\fi\ifx\removeHeaders\tempNo ,anonymous\fi]{acmart}
+\documentclass[10pt,sigconf,letterpaper,dvipsnames\ifx\removeHeaders\tempYes ,nonacm\fi]{acmart}
 \acmYear{2019}
@@
 \includegraphics[width=\columnwidth]{figures/traq_illustration_too_much.pdf}
-\caption{If the buffer is too large, loss-based \glspl{cca} will keep a unnecessary standing queue.
-This queue is unnecessary since it is not generally required to achieve full link utilization.
-The consequences are longer \glspl{rtt} which can impair application performance and fluidity as well as Quality of Experience.}
+\caption{If the buffer is too large, loss-based \glspl{cca} will keep a unnecessary standing queue. This queue is unnecessary since it is not generally required to achieve full link utilization.}
 \label{fig:tooMuch}
@@
 The other case we want to avoid is a persisting standing queue, as depicted in \autoref{fig:tooMuch}: In this case the buffer is too large meaning that optimal throughput is achieved but at the same time that an unnecessarily long standing queue is maintained.
-This oversized queue leads to \gls{rtt} being larger than necessary and can result in unresponsive applications.
+This oversized queue leads to \gls{rtt} being larger than necessary and can result in unresponsive applications and reduced Quality of Experience.
 Furthermore, the buffer space has to be allocated in the bottleneck device and keeps the memory from serving a more useful purpose.
@@
 \includegraphics[width=\columnwidth]{{{figures/traq_intervals}}}
-\caption{Illustration of the interval mechanism used by our \gls{qdisc}.
-The multiplier used is $\frac{1}{2}$.
+\caption{Illustration of the interval mechanism used by our \gls{qdisc} with a multiplier of $\frac{1}{2}$.
 Each \textit{guard interval} (GI) contains one or more intervals of which the longest one is considered the \textit{longest interval} (LI).
@@
 We add three configuration parameters to the \gls{qdisc}: the multiplier (default 1.25), the maximum buffer increase (default 2.0) and the maximum \gls{gi} duration (default 1.0\,s).
-We make the source code of our implementation freely available to enable reproducibility and encourage experimentation: \ifx\removeHeaders\tempYes\url{https://github.com/CN-TU/traq}\else{}[removed for anonymity]\fi .
+We make the source code of our implementation freely available to enable reproducibility and encourage experimentation: \url{https://github.com/CN-TU/traq}.
 
@@
 Figure\autoref{fig:fqCodelThroughputLong} shows that keeping the queuing delay that low is detrimental for achieving full throughput when using the Cubic \gls{cca}.
-With our \gls{qdisc} we achieve throughput that is more than 10\% higher overall, while only slightly increasing the average delay \todo{More systematic evaluation}.
+With our \gls{qdisc} we achieve throughput that is more than 10\% higher overall, while only slightly increasing the average delay.
 In addition to Cubic, we performed experiments with Reno.
@@
 
-We see this work as a first step towards adaptive queuing at bottlenecks and think that a promising direction for future work might be to explore the use of reinforcement learning.
+We see this work as a first step towards adaptive queuing at bottlenecks.
+An interesting further improvement could be to not statically initialize the buffer with a constant 100 packets but instead use experience from previous flows.
+Moreover, we think that a promising direction for future work might be to explore the use of reinforcement learning.
 Such an approach would fingerprint a flow and choose the buffer size accordingly, maximizing a chosen objective such as high throughput with minimal delay.
@@
 \section*{Acknowledgements}
-We thank \ifx\removeHeaders\tempYes{}Gernot Vormayr\else{}[removed for anonymity]\fi{} for providing us with the \ifx\removeHeaders\tempYes{}\texttt{py-virtnet}\else{}[removed for anonymity]\fi{} toolkit for building virtual networks.
+We thank Gernot Vormayr for providing us with a pre-release version of the \texttt{py-virtnet} toolkit for building virtual networks.
 

### 2019-10-17 tex/paper.tex
Removed anonymity

@@
 
-We evaluate our \gls{qdisc} using the \ifx\removeHeaders\tempYes{}\texttt{py-virtnet}\else{}[removed for anonymity]\fi{} toolkit to build a virtual network using Linux's network namespaces.
+We evaluate our \gls{qdisc} using the \texttt{py-virtnet} (\url{https://github.com/CN-TU/py-virtnet}) toolkit to build a virtual network using Linux's network namespaces.
 We initialize the buffer size for our \gls{qdisc} to 100 packets per flow like in fq.

### 2019-10-17 tex/figures/traq_common_beginning.tex
Change font in figures to Linux Libertine
Improve paper a little bit
Change figures

@@
 \documentclass{standalone}
+\usepackage{libertine}
 \usepackage[dvipsnames]{xcolor}
@@
 
+\draw[dashed] (0,1) -- ++(2.5,0);
+\node[align=center,font=\fontsize{3pt}{0}\selectfont,inner sep=0mm] at (2.25, 0.75) (bufferSizeLabel) {BDP+\\ buffer size};
+\draw[->, line width=0.05mm] (bufferSizeLabel) -- (2.25,1);
 \draw[name path=A] (0,0.5)-- ++(0.5, 0.5);
 \draw[from end of path=A, name path=B, -dot-=0, -dot-=0.5, -dot-=1] -- ++(0.25,0);
-\node[align=center,font=\fontsize{3pt}{0}\selectfont,inner sep=0mm] at (0.5, 1.25) (lossLabel) {packet loss};
-\draw[->, line width=0.05mm] (lossLabel) -- (0.5,1.05);
+\node[align=center,font=\fontsize{3pt}{0}\selectfont,inner sep=0mm] at (0.5, 0.7) (lossLabel) {packet loss};
+\draw[->, line width=0.05mm] (lossLabel) -- (0.5,0.95);
 \draw[from end of path=B, name path=C] ++(0,-0.5mm) -- ++(0,-0.95cm) -- ++(1,1);

### 2019-10-21 tex/paper.tex
Reduce fig size

@@
 \begin{figure}[h]
-\subfloat[Throughput of the fq \gls{qdisc}\label{fig:fqThroughput}
-]{\includegraphics[width=\columnwidth]{{{../plots/throughput_1_fq_cubic_10_20_60_0.5_bw_1571240937211.pcap}}}}{}
-\subfloat[Delay of the fq \gls{qdisc}.\label{fig:fqDelay}
-]{\includegraphics[width=\columnwidth]{{{../plots/rtt_1_fq_cubic_10_20_60_0.5_bw_1571240937211.pcap}}}}{}
-\caption{Throughput (\ref{fig:fqThroughput}) and delay (\ref{fig:fqDelay}) of a Cubic flow that is managed by fq at the bottleneck.
+%\subfloat[Throughput of the fq \gls{qdisc}\label{fig:fqThroughput}
+%]{
+\includegraphics[width=\columnwidth]{{{../plots/throughput_1_fq_cubic_10_20_60_0.5_bw_1571240937211.pcap}}}
+%}{}
+%\subfloat[Delay of the fq \gls{qdisc}.\label{fig:fqDelay}
+%]{
+\includegraphics[width=\columnwidth]{{{../plots/rtt_1_fq_cubic_10_20_60_0.5_bw_1571240937211.pcap}}}
+%}{}
+\caption{Throughput and delay of a Cubic flow that is managed by fq at the bottleneck.
 The initial bandwidth is 20\,Mbit/s but it is halved after 30\,s.
@@
 The delay is 100\,ms.
-The average \gls{rtt} is 101.36\,ms.
+The average \gls{rtt} is 101.61\,ms.
 Total throughput is 2615\,MB.}
@@
 The delay is 100\,ms.
-The average \gls{rtt} is 129.45\,ms.
+The average \gls{rtt} is 130.36\,ms.
 Total throughput is 2893\,MB.}
@@
 Here, over a 240\,s flow of 100\,Mbit/s with a delay of 50\,ms, our \gls{qdisc} achieves more than 20\% higher throughput than fq\_codel.
-The average \gls{rtt} for fq\_codel is 51.13\,ms while it is 96.59\,ms for our \gls{qdisc}.
+The average \gls{rtt} for fq\_codel is 51.34\,ms while it is 96.78\,ms for our \gls{qdisc}.
 
 Besides fq\_codel possibly leading to lower throughput as we have shown, another problem is that for links with a small \gls{bdp} fq\_codel can keep a standing queue akin to fq since fq\_codel only wants queuing delay to fall below 5\,ms once every 100\,ms.
-For example, on a link with a base \gls{rtt} of 5\,ms, fq\_codel would accept a permanent standing queue of 4.9\,ms, leading to overall \gls{rtt} being doubled.
+For example, on a link with a base \gls{rtt} of 5\,ms, fq\_codel would accept a permanent standing queue of 4.9\,ms, leading to overall \gls{rtt} being doubled, which is not required for achieving optimal throughput.
 
@@
 Our goal was to design a \gls{qdisc} which achieves maximum throughput while keeping the queuing delay as small as possible.
-The results of the experiments, which we perform with a prototype of our \gls{qdisc}, show that there are certain common scenarios in which current state-of-the-art \glspl{qdisc} like fq, fq\_codel and cake fall short of the optimum throughput by a significant margin while our approach succeeds in leading to full utilization of the link at the bottleneck.
-Furthermore, the aforementioned \glspl{qdisc} also suffer from standing delays is certain scenarios which we can also mitigate with our approach.
+The results of the experiments, which we perform with a prototype of our \gls{qdisc}, show that there are certain common scenarios in which current state-of-the-art \glspl{qdisc} like fq, fq\_codel and cake fall short of the optimum throughput by a significant margin while our approach succeeds in leading to full utilization of the link at the bottleneck (with a potential increase in delay).
+Furthermore, the aforementioned \glspl{qdisc} also suffer from standing queues in scenarios with small \gls{bdp} which we can also mitigate with our approach.
 

### 2019-10-21 plot_rtt_and_bandwidth.py

@@ -5,16 +5,18 @@ import sys
 import os
 import subprocess
 import statistics
+import pandas as pd
 plt.rcParams["font.family"] = "serif"
 plt.rcParams['pdf.fonttype'] = 42
 
 interval = 1
 pcap_file = sys.argv[1]
-assert pcap_file.startswith("sender_")
-
-rtt_command = f"tshark -Y tcp.srcport!=60000&&tcp.analysis.ack_rtt!=0 -r pcaps/{pcap_file} -Tfields -e frame.time_relative -e tcp.analysis.ack_rtt"
+assert pcap_file.startswith("sender_"), pcap_file
 
 receiver_pcap = 'receiver_'+('_'.join(pcap_file.split('_')[1:]))
+
+# rtt_command = f"tshark -r pcaps/{pcap_file} -Tfields -e frame.time_relative -e tcp.analysis.ack_rtt"
+new_rtt_command = f"../pantheon/tools/wintracker pcaps/{pcap_file}"
 
 retransmissions_command = f"tshark -Y tcp.srcport==60000&&tcp.analysis.retransmission -r pcaps/{pcap_file} -Tfields -e frame.time_relative"
 
@@ -24,10 +26,21 @@ bytes_command = f"tshark -r pcaps/{receiver_pcap} -q -z io,stat,{interval},SUM(i
 
 bytes_command_total = f"tshark -r pcaps/{receiver_pcap} -q -z io,stat,0,SUM(ip.len)ip.len&&tcp.srcport==60000"
 
-rtt_out = subprocess.run(rtt_command.split(), check=True, stdout=subprocess.PIPE, stderr=subprocess.STDOUT)
-rtt_out = [item.split("\t") for item in rtt_out.stdout.decode("utf-8").split("\n") if item!=""]
-rtt_results = [(float(item[0]), 1000*float(item[1])) for item in rtt_out]
-print("average rtt", statistics.mean(list(zip(*rtt_results))[1]))
+# rtt_out = subprocess.run(rtt_command.split(), check=True, stdout=subprocess.PIPE, stderr=subprocess.STDOUT)
+rtt_out = subprocess.run(new_rtt_command, check=True, stdout=subprocess.PIPE, stderr=subprocess.STDOUT, shell=True)
+
+df = pd.read_csv(f"pcaps/{'.'.join(pcap_file.split('.')[:-1])}_full_1.csv")
+
+ack_timestamps = df['ackTimestamp'].tolist()
[…]

### 2019-10-21 tex/paper.tex
Rename

@@
 We add three configuration parameters to the \gls{qdisc}: the multiplier (default 1.25), the maximum buffer increase (default 2.0) and the maximum \gls{gi} duration (default 1.0\,s).
-We make the source code of our implementation freely available to enable reproducibility and encourage experimentation: \url{https://github.com/CN-TU/traq}.
+We make the source code of our implementation freely available to enable reproducibility and encourage experimentation: \url{https://github.com/CN-TU/cocoa-qdisc}.
 
@@
 \begin{figure}[h]
-\subfloat[Throughput of our \gls{qdisc}\label{fig:cnThroughput}
-]{\includegraphics[width=\columnwidth]{{{../plots/throughput_1_cn_cubic_10_20_60_0.5_bw_1571241047054.pcap}}}}{}
-\subfloat[Delay of our \gls{qdisc}.\label{fig:cnDelay}
-]{\includegraphics[width=\columnwidth]{{{../plots/rtt_1_cn_cubic_10_20_60_0.5_bw_1571241047054.pcap}}}}{}
-\caption{Throughput (\ref{fig:cnThroughput}) and delay (\ref{fig:cnDelay}) of a Cubic flow that is managed by our \gls{qdisc} at the bottleneck.
+%\subfloat[Throughput of our \gls{qdisc}\label{fig:cnThroughput}
+%]{
+\includegraphics[width=\columnwidth]{{{../plots/throughput_1_cn_cubic_10_20_60_0.5_bw_1571241047054.pcap}}}
+%}{}
+%\subfloat[Delay of our \gls{qdisc}.\label{fig:cnDelay}
+%]{
+\includegraphics[width=\columnwidth]{{{../plots/rtt_1_cn_cubic_10_20_60_0.5_bw_1571241047054.pcap}}}
+%}{}
+\caption{Throughput and delay of a Cubic flow that is managed by our \gls{qdisc} at the bottleneck.
 The initial bandwidth is 20\,Mbit/s but it is halved after 30\,s.
@@
 \begin{figure}[h]
-\subfloat[Throughput of our \gls{qdisc}\label{fig:cnThroughput2}
-]{\includegraphics[width=\columnwidth]{{{../plots/throughput_1_cn_cubic_10_20_60_2.0_bw_1571241151290.pcap}}}}{}
-\subfloat[Delay of our \gls{qdisc}.\label{fig:cnDelay2}
-]{\includegraphics[width=\columnwidth]{{{../plots/rtt_1_cn_cubic_10_20_60_2.0_bw_1571241151290.pcap}}}}{}
-\caption{Throughput (\ref{fig:cnThroughput2}) and delay (\ref{fig:cnDelay2}) of a Cubic flow that is managed by our \gls{qdisc} at the bottleneck.
+%\subfloat[Throughput of our \gls{qdisc}\label{fig:cnThroughput2}
+%]{
+\includegraphics[width=\columnwidth]{{{../plots/throughput_1_cn_cubic_10_20_60_2.0_bw_1571241151290.pcap}}}
+%}{}
+%\subfloat[Delay of our \gls{qdisc}.\label{fig:cnDelay2}
+%]{
+\includegraphics[width=\columnwidth]{{{../plots/rtt_1_cn_cubic_10_20_60_2.0_bw_1571241151290.pcap}}}
+%}{}
+\caption{Throughput and delay of a Cubic flow that is managed by our \gls{qdisc} at the bottleneck.
 The initial bandwidth is 20\,Mbit/s but it is doubled after 30\,s.
@@
 \begin{figure}[h]
-\subfloat[Throughput of the fq\_codel \gls{qdisc}\label{fig:fqCodelThroughputLong}
-]{\includegraphics[width=\columnwidth]{{{../plots/throughput_1_fq_codel_cubic_100_100_240_1.0_bw_1571244312838.pcap}}}}{}
-\subfloat[Delay of the fq\_codel \gls{qdisc}.\label{fig:fqCodelDelayLong}
-]{\includegraphics[width=\columnwidth]{{{../plots/rtt_1_fq_codel_cubic_100_100_240_1.0_bw_1571244312838.pcap}}}}{}
-\caption{Throughput (\ref{fig:fqCodelThroughputLong}) and delay (\ref{fig:fqCodelDelayLong}) of a Cubic flow that is managed by fq\_codel at the bottleneck.
+%\subfloat[Throughput of the fq\_codel \gls{qdisc}\label{fig:fqCodelThroughputLong}
+%]{
+\includegraphics[width=\columnwidth]{{{../plots/throughput_1_fq_codel_cubic_100_100_240_1.0_bw_1571244312838.pcap}}}
+%}{}
+%\subfloat[Delay of the fq\_codel \gls{qdisc}.\label{fig:fqCodelDelayLong}
+%]{
+\includegraphics[width=\columnwidth]{{{../plots/rtt_1_fq_codel_cubic_100_100_240_1.0_bw_1571244312838.pcap}}}
+%}{}
+\caption{Throughput and delay of a Cubic flow that is managed by fq\_codel at the bottleneck.
 The bandwidth is 100\,Mbit/s.
@@
 \begin{figure}[h]
-\subfloat[Throughput of our \gls{qdisc}\label{fig:cnThroughputLong}
-]{\includegraphics[width=\columnwidth]{{{../plots/throughput_1_cn_cubic_100_100_240_1.0_bw_1571244036720.pcap}}}}{}
-\subfloat[Delay of our \gls{qdisc}.\label{fig:cnDelayLong}
-]{\includegraphics[width=\columnwidth]{{{../plots/rtt_1_cn_cubic_100_100_240_1.0_bw_1571244036720.pcap}}}}{}
-\caption{Throughput (\ref{fig:cnThroughputLong}) and delay (\ref{fig:cnDelayLong}) of a Cubic flow that is managed by our \gls{qdisc} at the bottleneck.
+%\subfloat[Throughput of our \gls{qdisc}\label{fig:cnThroughputLong}
+%]{
+\includegraphics[width=\columnwidth]{{{../plots/throughput_1_cn_cubic_100_100_240_1.0_bw_1571244036720.pcap}}}
+%}{}
+%\subfloat[Delay of our \gls{qdisc}.\label{fig:cnDelayLong}
+%]{
+\includegraphics[width=\columnwidth]{{{../plots/rtt_1_cn_cubic_100_100_240_1.0_bw_1571244036720.pcap}}}
+%}{}
+\caption{Throughput and delay of a Cubic flow that is managed by our \gls{qdisc} at the bottleneck.
 The bandwidth is 100\,Mbit/s.
@@
 Next we compare our \gls{qdisc} against fq\_codel.
-Figure\autoref{fig:fqCodelDelayLong} shows that fq\_codel keeps the queuing delay under 5\,ms.
-Figure\autoref{fig:fqCodelThroughputLong} shows that keeping the queuing delay that low is detrimental for achieving full throughput when using the Cubic \gls{cca}.
+\autoref{fig:fqCodel} shows that fq\_codel keeps the queuing delay under 5\,ms.
+Also it shows that keeping the queuing delay that low is detrimental for achieving full throughput when using the Cubic \gls{cca}.
 With our \gls{qdisc} we achieve throughput that is more than 10\% higher overall, while only slightly increasing the average delay.
@@
 \begin{figure}[h]
-\subfloat[Throughput of our \gls{qdisc}\label{fig:cnThroughputBBR}
-]{\includegraphics[width=\columnwidth]{{{../plots/throughput_1_cn_bbr_10_50_60_0.5_delay_1571288916582.pcap}}}}{}
-\subfloat[Delay of our \gls{qdisc}.\label{fig:cnDelayBBR}
-]{\includegraphics[width=\columnwidth]{{{../plots/rtt_1_cn_bbr_10_50_60_0.5_delay_1571288916582.pcap}}}}{}
-\caption{Throughput (\ref{fig:cnThroughputBBR}) and delay (\ref{fig:cnDelayBBR}) of a BBR~v1 flow that is managed by our \gls{qdisc} at the bottleneck.
-The bandwidth is 50\,Mbit/s.
-The delay is 10\,ms and it is cut by half after 30\,s.}
+%\subfloat[Throughput of our \gls{qdisc}\label{fig:cnThroughputBBR}
+%]{
+\includegraphics[width=\columnwidth]{{{../plots/throughput_1_cn_bbr_10_50_60_0.5_delay_1571288916582.pcap}}}
+%}{}
+%\subfloat[Delay of our \gls{qdisc}.\label{fig:cnDelayBBR}
+%]{
+\includegraphics[width=\columnwidth]{{{../plots/rtt_1_cn_bbr_10_50_60_0.5_delay_1571288916582.pcap}}}
+%}{}
+\caption{Throughput and delay of a BBR~v1 flow that is managed by our \gls{qdisc} at the bottleneck. The bandwidth is 50\,Mbit/s. The delay is 10\,ms and it is cut by half after 30\,s.}
 \label{fig:cnBBR}

### 2019-10-21 tex/paper.tex
Incorporate changes

@@
 \newcommand{\algorithmautorefname}{Algorithm}
+\newcommand{\name}{cocoa}
 
@@
 \newacronym{bdp}{BDP}{Bandwidth Delay Product}
+\newacronym{cocoa}{cocoa}{Congestion Control Aware qdisc}
 
@@
 Recent model-based congestion control algorithms such as BBR use repeated measurements at the end-point to build a model of the network connection and use it to achieve optimal throughput with low queuing delay.
-Conversely, applying this model-based approach to Active Queue Management (AQM) has so far been underinvestigated.
+Conversely, applying this model-based approach to Active Queue Management (AQM) has so far received less attention.
 We propose an AQM scheduler based on fair queuing, which adapts the buffer size depending on the needs of each flow without requiring active participation from the endpoint.
-We implement this scheduler for the Linux kernel and show that it interacts well with the most common congestion control algorithms and can significantly increase throughput compared to CoDeL while avoiding overbuffering.
+We implement this scheduler for the Linux kernel and show that it interacts well with the most common congestion control algorithms and can significantly increase throughput compared to CoDel while avoiding overbuffering.
 \end{abstract}
@@
 
-\newcommand{\codel}{CoDeL}
+\newcommand{\codel}{CoDel}
 
@@
 
-In the last decades various \gls{aqm} mechanisms have been proposed to minimize excessive standing queues in the Internet.
+In the last decades various \gls{aqm} mechanisms have been proposed to minimize excessive \textit{standing queues} in the Internet.
 One of the most influential recent efforts is \codel{} \cite{nichols_controlling_2012} whose goal is that the queuing delay at the bottleneck link is at least once under 5\;ms in a moving window of 100\;ms.
 
-While it is important to keep the queuing delay constrained, it is also necessary to ensure that overly ``aggressive'' flows cannot benefit by ``stealing'' less aggressive flows' bandwidth.
+While it is important to keep the queuing delay constrained, it is also necessary to ensure that overly aggressive flows cannot benefit by ``stealing'' less aggressive flows' bandwidth.
 Thus researchers and engineers have developed \textit{\gls{fq}} mechanisms \cite{shreedhar_efficient_1996,dumazet_pkt_sched:_2013} to isolate different flows' queues so that for example a delay-sensitive live video call cannot be impaired by a concurrent bulk transfer which takes all the available bandwidth.
@@
 \cite{hoiland-jorgensen_piece_2018} expand upon this and create the \textit{cake} \gls{qdisc} that also adds features such as not only per-flow queuing but also per-host queuing for even increased fairness.
-Furthermore they also include bandwidth shaping into their solution and aim to create one \gls{qdisc} that is easy to configure and can be easily deployed on home routers and offers all features in one solution.
+Furthermore they also include bandwidth shaping into their solution and aim to create one \gls{qdisc} that is easy to configure, can be easily deployed on home routers and offers all features in one solution.
 
@@
 We show that the impaired performance is a result of keeping the queuing delay under 5\;ms, which hinders \glspl{cca} such as Reno or Cubic from reaching maximum throughput.
-The problem is that these \glspl{qdisc} aim to keep the delay under the threshold no matter the effect on throughput and do not take the \gls{cc} of the flow into account.
+The main drawback is that these \glspl{qdisc} aim to keep the delay under the threshold no matter the effect on throughput and do not take the \gls{cc} of the flow into account.
 
@@
 \end{enumerate}
-Finally, we develop a prototype of our \gls{qdisc} and show that it can achieve these objectives for the most common loss-based \glspl{cca}, Reno \cite {jacobson_congestion_1988} and Cubic \cite{ha_cubic:_2008}.
-These \glspl{cca} work by continuously increasing the number of bytes that are allowed to be in the network (congestion window) if no packet loss is experienced and by sharply decreasing this number if a packet is lost (multiplicative decrease).
+Finally, we develop a prototype of \gls{cocoa} and show that it can achieve these objectives for the most common loss-based \glspl{cca}, Reno \cite {jacobson_congestion_1988} and Cubic \cite{ha_cubic:_2008}.
+These \glspl{cca} operate by continuously increasing the number of bytes that are allowed to be in the network (congestion window) if no packet loss is experienced and by sharply decreasing this number if a packet is lost (multiplicative decrease).
 With Cubic being the default \gls{cc} in all major OS, we especially emphasize our evaluation on improving its performance.
@@
 BBR is thus considered to be \textit{model-based}.
-We demonstrate that our \gls{qdisc} also behaves well in interaction with BBR~v1.
+We demonstrate that \gls{cocoa} also behaves well in interaction with BBR~v1.
 
@@
 \begin{figure}[h]
-\includegraphics[width=\columnwidth]{figures/traq_illustration_too_little.pdf}
+\includegraphics[width=\columnwidth]{figures/cocoa_illustration_too_little.pdf}
 \caption{If the buffer is too small, loss-based \glspl{cca} will not be able to fully utilize the link since they send too few data following the multiplicative decrease that occurs after packet loss.}
@@
 \begin{figure}[h]
-\includegraphics[width=\columnwidth]{figures/traq_illustration_too_much.pdf}
+\includegraphics[width=\columnwidth]{figures/cocoa_illustration_too_much.pdf}
 \caption{If the buffer is too large, loss-based \glspl{cca} will keep a unnecessary standing queue. This queue is unnecessary since it is not generally required to achieve full link utilization.}
@@
 
-Since the goal of our algorithm is to achieve optimal throughput irrespective of the \gls{cca}, we aim to avoid the scenario depicted in \autoref{fig:tooLittle}: Here the buffer is too small, meaning that the \glspl{cca} never manages to achieve full utilization and periodically underutilizes the link, leading to, for example, a user waiting longer for a software update to finish or a video game to download on their video game console.
+Since the goal of our algorithm is to achieve optimal throughput irrespective of the \gls{cca}, we aim to avoid the scenario depicted in \autoref{fig:tooLittle}: Here the buffer is too small, meaning that the \glspl{cca} never manages to achieve full utilization and periodically underutilizes the link, leading to, for example, a user waiting longer for a software update to finish or a game to download on their video game console.
 
@@
 \end{enumerate}
-The tricky thing about this is to define the measurement period: As \autoref{fig:tooLittle} and \autoref{fig:tooMuch} show, we would like to use the measurement period that is the longest period between two packet losses.
+The challenge about this is to define the measurement period: As \autoref{fig:tooLittle} and \autoref{fig:tooMuch} show, we would like to use the measurement period that is the longest period between two packet losses.
 If this period is observed, it is possible to compute the standing queue as simply the minimum queue observed in this interval.
 For example, if the minimum queue is 3 packets it means, that there is a standing queue of 3 packets and the algorithm will reduce buffer size by 3 to eliminate the unnecessary queuing.
-On the other side, if the buffer is too small, the algorithm computes how long the link was underutilized and how many packets could have been transmitted during the idle period.
+Conversely, if the buffer is too small, the algorithm computes how long the link was underutilized and how many packets could have been transmitted during the idle period.
 The buffer size is then increased by this number of packets.
 
-The problem is now the following: If the measurement period is mistakenly not the longest interval shown in the figure, but the interval between two adjacent packet losses, the algorithm would assume an enormous standing queue and drastically reduce the buffer to be close to zero.
+This is not so trivial because of the following consideration: If the measurement period is mistakenly not the longest interval shown in the figure, but the interval between two adjacent packet losses, the algorithm would assume an enormous standing queue and drastically reduce the buffer to be close to zero.
 This would then lead to severe underutilization afterwards.
@@
 The solution is thus to use a \textit{\gls{gi}}: The algorithm always waits for a specified minimum amount of time, picks the longest interval without packet loss in this guard interval and then applies its logic to reduce the standing queue on this \textit{\gls{li}}.
-We dynamically compute the \gls{gi} as a multiple (parameter of our algorithm) of the previous \gls{li} (\autoref{fig:intervals}):
+We dynamically compute the \gls{gi} as a multiple or a fraction (parameter of our algorithm) of the previous \gls{li} (\autoref{fig:intervals}):
 
 \begin{figure}[h]
-\includegraphics[width=\columnwidth]{{{figures/traq_intervals}}}
-\caption{Illustration of the interval mechanism used by our \gls{qdisc} with a multiplier of $\frac{1}{2}$.
+\includegraphics[width=\columnwidth]{{{figures/cocoa_intervals}}}
+\caption{Illustration of the interval mechanism used by \gls{cocoa} with a multiplier of $\frac{1}{2}$.
 Each \textit{guard interval} (GI) contains one or more intervals of which the longest one is considered the \textit{longest interval} (LI).
 The \gls{gi} is always at least the length of the previous \gls{li} times the multiplier.
-A \gls{gi} ends at the first packet loss after that is after the previous \gls{li} times the multiplier.}
+A \gls{gi} ends at the first packet loss after the previous \gls{li} times the multiplier.}
 \label{fig:intervals}
@@
 
-The purpose of the multiplier, which is used to multiply the \gls{li} of previous \gls{gi} to define the new \gls{gi}, is to prevent that the short gap between two subsequent packet losses is is used as the interval to determine whether the buffer size has to be adapted.
-This means that the multiplier must be large enough so that it allows to ``jump'' over the packet losses that occur when the buffer is full.
+The purpose of the multiplier, which is used to multiply the \gls{li} of previous \gls{gi} to define the new \gls{gi}, is to prevent that the short gap between two subsequent packet losses is used as the interval to determine whether the buffer size has to be adapted.
+This means that the multiplier must be large enough so that it allows to skip over the packet losses that occur when the buffer is full.
 For Reno and Cubic, the loss-free period is very large compared to the period during which packet loss occurs.
@@
 If the \gls{gi} ends after the first 6 phases and also the \gls{li} was these six phase, with a factor of $0.5$, the next \gls{gi} would thus end at the end of phase 8.
-Then, the \gls{li} would be 8th cycle.
+Then, the \gls{li} would be the 8th cycle.
 The next \gls{gi} would then be of the duration of $0.5$ phases and would then end in the middle of cycle 9.
@@
 
-Another potential problem is the following: If the link speed suddenly increases drastically, our algorithm would sharply increase the buffer size.
+Another potential problem is the following: If the available share of bandwidth suddenly increases drastically, our algorithm would sharply increase the buffer size.
 In this case it can happen that the buffer becomes much too large.
@@
 We set this parameter to 1\,s.
-This means that we also assume that no flows are handled by our \gls{qdisc} which have an \gls{rtt} larger than 1\,s.
+This means that we also assume that no flows are handled by \gls{cocoa} which have an \gls{rtt} larger than 1\,s.
 Another possibility for setting this parameter is to set it to a multiple of the \gls{rtt}, for example 2, 4 or 8 times the most recent \gls{rtt}.
@@
 
-We implement our \gls{qdisc} as an extension of the ``fq'' \gls{qdisc} \cite{dumazet_pkt_sched:_2013}.
-This means that when using our \gls{qdisc} also all features offered by fq are available.
+We implement \gls{cocoa} as an extension of the ``fq'' \gls{qdisc} \cite{dumazet_pkt_sched:_2013}.
+This means that when using \gls{cocoa} also all features offered by fq are available.
 We add three configuration parameters to the \gls{qdisc}: the multiplier (default 1.25), the maximum buffer increase (default 2.0) and the maximum \gls{gi} duration (default 1.0\,s).
@@
 
-We evaluate our \gls{qdisc} using the \texttt{py-virtnet} (\url{https://github.com/CN-TU/py-virtnet}) toolkit to build a virtual network using Linux's network namespaces.
-We initialize the buffer size for our \gls{qdisc} to 100 packets per flow like in fq.
+We evaluate \gls{cocoa} using the \texttt{py-virtnet} (\url{https://github.com/CN-TU/py-virtnet}) toolkit to build a virtual network using Linux's network namespaces.
+We initialize the buffer size for \gls{cocoa} to 100 packets per flow like in fq.
 
@@
 The delay is 10\,ms.
-It is clear that there is a standing queue and after the bandwidth halves, the minimum delay is 100\,ms even though 10\,ms would be the possible.}
+It is clear that there is a standing queue and after the bandwidth halves, the minimum delay is 100\,ms even though 10\,ms would be possible.}
 \label{fig:fq}
@@
 \begin{figure}[h]
-%\subfloat[Throughput of our \gls{qdisc}\label{fig:cnThroughput}
+%\subfloat[Throughput of \gls{cocoa}\label{fig:cnThroughput}
 %]{
@@
 %}{}
-%\subfloat[Delay of our \gls{qdisc}.\label{fig:cnDelay}
+%\subfloat[Delay of \gls{cocoa}.\label{fig:cnDelay}
 %]{
@@
 %}{}
-\caption{Throughput and delay of a Cubic flow that is managed by our \gls{qdisc} at the bottleneck.
-The initial bandwidth is 20\,Mbit/s but it is halved after 30\,s.
+\caption{Throughput and delay of a Cubic flow that is managed by \gls{cocoa} at the bottleneck.
+The initial bandwidth is 20\,Mbit/s but we halve it after 30\,s.
 The delay is 10\,ms.
-Our \gls{qdisc} manages to keep the queuing delay minimal while achieving the same throughput as with an oversized buffer.}
+\gls{cocoa} manages to keep the queuing delay minimal while achieving the same throughput as with an oversized buffer.}
 \label{fig:cn}
@@
 \begin{figure}[h]
-%\subfloat[Throughput of our \gls{qdisc}\label{fig:cnThroughput2}
+%\subfloat[Throughput of \gls{cocoa}\label{fig:cnThroughput2}
 %]{
@@
 %}{}
-%\subfloat[Delay of our \gls{qdisc}.\label{fig:cnDelay2}
+%\subfloat[Delay of \gls{cocoa}.\label{fig:cnDelay2}
 %]{
@@
 %}{}
-\caption{Throughput and delay of a Cubic flow that is managed by our \gls{qdisc} at the bottleneck.
-The initial bandwidth is 20\,Mbit/s but it is doubled after 30\,s.
+\caption{Throughput and delay of a Cubic flow that is managed by \gls{cocoa} at the bottleneck.
+The initial bandwidth is 20\,Mbit/s but we double after 30\,s.
 The delay is 10\,ms.
-Also a sudden large increase of bandwidth is handled well by our \gls{qdisc}.}
+Also a sudden large increase of bandwidth is handled well by \gls{cocoa}.}
 \label{fig:cn2}
@@
 
-First we evaluate if our \gls{qdisc} is able to maintain a small buffer like it is necessary in case of small \glspl{bdp}.
+First we evaluate if \gls{cocoa} is able to maintain a small buffer like it is necessary in case of small \glspl{bdp}.
 As can be seen in \autoref{fig:fq} the regular fq qdisc maintains a standing queue and unnecessarily leads to a significant increase in \gls{rtt} (100\,ms minimum, when the real minimum is 10\,ms).
 In contrast, \autoref{fig:cn} maintains the full throughput while at the same time keeping the delay minimal and not keeping a standing queue.
-Also, when drastically changing the bandwidth by halving it (\autoref{fig:cn}) or doubling it (\autoref{fig:cn2}) our \gls{qdisc} rapidly adapts to the new conditions and returns to the optimum state.
+Also, when drastically changing the bandwidth by halving it (\autoref{fig:cn}) or doubling it (\autoref{fig:cn2}) \gls{cocoa} rapidly adapts to the new conditions and returns to the optimum state.
 
@@
 \begin{figure}[h]
-%\subfloat[Throughput of our \gls{qdisc}\label{fig:cnThroughputLong}
+%\subfloat[Throughput of \gls{cocoa}\label{fig:cnThroughputLong}
 %]{
@@
 %}{}
-%\subfloat[Delay of our \gls{qdisc}.\label{fig:cnDelayLong}
+%\subfloat[Delay of \gls{cocoa}.\label{fig:cnDelayLong}
 %]{
@@
 %}{}
-\caption{Throughput and delay of a Cubic flow that is managed by our \gls{qdisc} at the bottleneck.
+\caption{Throughput and delay of a Cubic flow that is managed by \gls{cocoa} at the bottleneck.
 The bandwidth is 100\,Mbit/s.
@@
 
-Next we compare our \gls{qdisc} against fq\_codel.
+\begin{table}[t]
+\caption{Comparison of \glspl{qdisc} for loss-based \glspl{cca} on a link of 100\;Mbit/s with an \gls{rtt} of 50\;ms.
+Utilization in \% of the maximum capacity; \gls{rtt} in milliseconds.
+For Cubic fq achieves good performance.
+This is because its default buffer size of 100 packets coincidentially works well for this specific link speed and \gls{rtt}.} \label{tab:performance_results}
+\begin{tabular}{l r r r r r r} \toprule
+& \multicolumn{2}{l}{fq} & \multicolumn{2}{l}{fq\_codel} & \multicolumn{2}{l}{\gls{cocoa}} \\
+\cmidrule(r){2-3} \cmidrule(lr){4-5} \cmidrule(l){6-7}
+& Util. & \gls{rtt} & Util. & \gls{rtt} & Util. & \gls{rtt} \\ \midrule
+Cubic	& 96.2 & 54 & 92.6 & 52 & 98.5 & 66.5 \\
+Reno	& 84.1 & 52.7 & 81.1 & 51.5 & 97.9 & 98.2 \\
+\bottomrule
+\end{tabular}
+\end{table}
+
+Next we compare \gls{cocoa} against fq\_codel.
 \autoref{fig:fqCodel} shows that fq\_codel keeps the queuing delay under 5\,ms.
 Also it shows that keeping the queuing delay that low is detrimental for achieving full throughput when using the Cubic \gls{cca}.
-With our \gls{qdisc} we achieve throughput that is more than 10\% higher overall, while only slightly increasing the average delay.
+With \gls{cocoa} we achieve throughput that is more than 10\% higher overall, while only slightly increasing the average delay.
 In addition to Cubic, we performed experiments with Reno.
-Here, over a 240\,s flow of 100\,Mbit/s with a delay of 50\,ms, our \gls{qdisc} achieves more than 20\% higher throughput than fq\_codel.
-The average \gls{rtt} for fq\_codel is 51.34\,ms while it is 96.78\,ms for our \gls{qdisc}.
+Here, over a 240\,s flow of 100\,Mbit/s with a delay of 50\,ms, \gls{cocoa} achieves more than 20\% higher throughput than fq\_codel.
+The average \gls{rtt} for fq\_codel is 51.34\,ms while it is 96.78\,ms for \gls{cocoa}.
+\autoref{tab:performance_results} shows that \gls{cocoa} reaches link utilizations close to 100\% for both Cubic and Reno in large \gls{bdp} scenarios.
+\gls{rtt} can be higher than for other \glspl{qdisc} if this is required to achieve optimum throughput for the \gls{cca} for the given link speed and base \gls{rtt}.
 
@@
 
-We also ran experiments with the cake \gls{qdisc}. However, results were very similar to those of fq\_codel, which is no surprise since cake is based upon fq\_codel.
+We also ran experiments with the cake \gls{qdisc}. However, results were very similar to those of fq\_codel, which is no surprise since cake is based on fq\_codel.
 
 \begin{figure}[h]
-%\subfloat[Throughput of our \gls{qdisc}\label{fig:cnThroughputBBR}
+%\subfloat[Throughput of \gls{cocoa}\label{fig:cnThroughputBBR}
 %]{
@@
 %}{}
-%\subfloat[Delay of our \gls{qdisc}.\label{fig:cnDelayBBR}
+%\subfloat[Delay of \gls{cocoa}.\label{fig:cnDelayBBR}
 %]{
@@
 %}{}
-\caption{Throughput and delay of a BBR~v1 flow that is managed by our \gls{qdisc} at the bottleneck. The bandwidth is 50\,Mbit/s. The delay is 10\,ms and it is cut by half after 30\,s.}
+\caption{Throughput and delay of a BBR~v1 flow that is managed by \gls{cocoa} at the bottleneck. The bandwidth is 50\,Mbit/s. The delay is 10\,ms and we halve it after 30\,s.}
 \label{fig:cnBBR}
@@
 
-Also for BBR~v1, throughput quickly reaches its maximum while no standing queue is forming \autoref{fig:cnBBR}.
+Also for BBR~v1, throughput quickly reaches its maximum while no standing queue is forming (\autoref{fig:cnBBR}).
 
@@
 Our goal was to design a \gls{qdisc} which achieves maximum throughput while keeping the queuing delay as small as possible.
-The results of the experiments, which we perform with a prototype of our \gls{qdisc}, show that there are certain common scenarios in which current state-of-the-art \glspl{qdisc} like fq, fq\_codel and cake fall short of the optimum throughput by a significant margin while our approach succeeds in leading to full utilization of the link at the bottleneck (with a potential increase in delay).
+The results of the experiments, which we perform with a prototype of \gls{cocoa}, show that there are certain common scenarios in which current state-of-the-art \glspl{qdisc} like fq, fq\_codel and cake fall short of the optimum throughput by a significant margin while our approach succeeds in leading to full utilization of the link at the bottleneck (with a potential increase in delay).
 Furthermore, the aforementioned \glspl{qdisc} also suffer from standing queues in scenarios with small \gls{bdp} which we can also mitigate with our approach.
 
-We see this work as a first step towards adaptive queuing at bottlenecks.
+We see this work as a initial step towards flow-adaptive queuing at bottlenecks.
 An interesting further improvement could be to not statically initialize the buffer with a constant 100 packets but instead use experience from previous flows.
@@
 \section*{Acknowledgements}
-We thank Gernot Vormayr for providing us with a pre-release version of the \texttt{py-virtnet} toolkit for building virtual networks.
+We thank Gernot Vormayr for providing us with \texttt{py-virtnet}, a toolkit for easily building virtual networks.
 

### 2019-10-22 tex/paper.tex
Remove private script

@@
 \includegraphics[width=\columnwidth]{figures/cocoa_illustration_too_much.pdf}
-\caption{If the buffer is too large, loss-based \glspl{cca} will keep a unnecessary standing queue. This queue is unnecessary since it is not generally required to achieve full link utilization.}
+\caption{If the buffer is too large, loss-based \glspl{cca} will keep a unnecessary standing queue which is not necessary for achieving full link utilization.}
 \label{fig:tooMuch}
@@
 We initialize the buffer size for \gls{cocoa} to 100 packets per flow like in fq.
+The testbed consists of a sender and a receiver connected via a switch.
+We runs the \gls{qdisc} being tested on the interface that connects the switch to the receiver.
+We introduce delay with \texttt{netem} and limit bandwidth with \texttt{htb}.
 
@@
 %}{}
-\caption{Throughput and delay of a Cubic flow that is managed by fq at the bottleneck.
+\caption{Throughput and delay of a Cubic flow with fq at the bottleneck.
 The initial bandwidth is 20\,Mbit/s but it is halved after 30\,s.
@@
 %}{}
-\caption{Throughput and delay of a Cubic flow that is managed by \gls{cocoa} at the bottleneck.
+\caption{Throughput and delay of a Cubic flow with \gls{cocoa} at the bottleneck.
 The initial bandwidth is 20\,Mbit/s but we halve it after 30\,s.
 The delay is 10\,ms.
-\gls{cocoa} manages to keep the queuing delay minimal while achieving the same throughput as with an oversized buffer.}
+\gls{cocoa} keeps the queuing delay minimal while achieving the same throughput as with an oversized buffer.}
 \label{fig:cn}
@@
 %}{}
-\caption{Throughput and delay of a Cubic flow that is managed by \gls{cocoa} at the bottleneck.
+\caption{Throughput and delay of a Cubic flow with \gls{cocoa} at the bottleneck.
 The initial bandwidth is 20\,Mbit/s but we double after 30\,s.
@@
 %}{}
-\caption{Throughput and delay of a Cubic flow that is managed by fq\_codel at the bottleneck.
-The bandwidth is 100\,Mbit/s.
-The delay is 100\,ms.
-The average \gls{rtt} is 101.61\,ms.
-Total throughput is 2615\,MB.}
+\caption{Throughput and delay of a Cubic flow with fq\_codel at the bottleneck. The bandwidth is 100\,Mbit/s. The delay is 100\,ms. The average \gls{rtt} is 101.61\,ms. Total throughput is 2615\,MB.}
 \label{fig:fqCodel}
@@
 %}{}
-\caption{Throughput and delay of a Cubic flow that is managed by \gls{cocoa} at the bottleneck.
+\caption{Throughput and delay of a Cubic flow with \gls{cocoa} at the bottleneck.
 The bandwidth is 100\,Mbit/s.
@@
 \caption{Comparison of \glspl{qdisc} for loss-based \glspl{cca} on a link of 100\;Mbit/s with an \gls{rtt} of 50\;ms.
-Utilization in \% of the maximum capacity; \gls{rtt} in milliseconds.
+Utilization in \% of maximum capacity; \gls{rtt} in milliseconds.
 For Cubic fq achieves good performance.
@@
 Besides fq\_codel possibly leading to lower throughput as we have shown, another problem is that for links with a small \gls{bdp} fq\_codel can keep a standing queue akin to fq since fq\_codel only wants queuing delay to fall below 5\,ms once every 100\,ms.
-For example, on a link with a base \gls{rtt} of 5\,ms, fq\_codel would accept a permanent standing queue of 4.9\,ms, leading to overall \gls{rtt} being doubled, which is not required for achieving optimal throughput.
+%For example, on a link with a base \gls{rtt} of 5\,ms, fq\_codel would accept a permanent standing queue of 4.9\,ms, leading to overall \gls{rtt} being doubled, which is not required for achieving optimal throughput.
+We also ran experiments for these low \gls{bdp} scenarios: For a link speed of 20\;Mbit/s and a base \gls{rtt} of 1\;ms for a Cubic flow both \gls{cocoa} and fq\_codel achieve 98\% utilization.
+The average \gls{rtt} is 6.86\,ms for \gls{cocoa} and 7.31\;ms for fq\_codel.
 
@@
 %}{}
-\caption{Throughput and delay of a BBR~v1 flow that is managed by \gls{cocoa} at the bottleneck. The bandwidth is 50\,Mbit/s. The delay is 10\,ms and we halve it after 30\,s.}
+\caption{Throughput and delay of a BBR~v1 flow with \gls{cocoa} at the bottleneck. The bandwidth is 50\,Mbit/s. The delay is 10\,ms and we halve it after 30\,s.}
 \label{fig:cnBBR}

### 2019-10-22 average_rtt.py

@@ -1,11 +1,0 @@
-#!/usr/bin/env python3
-import statistics
-import sys
-
-all_lines = sys.stdin.readlines()
-# print("all_lines", all_lines)
-all_lines = [float(line.split("\t")[-1]) for line in all_lines]
-
-# print("all_lines", all_lines)
-print("mean rtt")
-print(statistics.mean(all_lines))

### 2019-10-22 iproute2/tc/cn_opts.h
Rename module to cocoa

@@ -1,46 +1,0 @@
-/* SPDX-License-Identifier: GPL-2.0 WITH Linux-syscall-note */
-#ifndef __LINUX_CN_OPTS_H
-#define __LINUX_CN_OPTS_H
-
-// FIXME: This should include a file which defines "TCA_CN_MAX", but I couldn't find it.
-// So this only works if you include that mysterious file before this one.
-
-enum {
-	TCA_CN_UNSPEC,
-
-	TCA_CN_PLIMIT,		/* limit of total number of packets in queue */
-
-	TCA_CN_FLOW_PLIMIT,	/* limit of packets per flow */
-
-	TCA_CN_QUANTUM,		/* RR quantum */
-
-	TCA_CN_INITIAL_QUANTUM,		/* RR quantum for new flow */
-
-	TCA_CN_RATE_ENABLE,	/* enable/disable rate limiting */
-
-	TCA_CN_FLOW_DEFAULT_RATE,/* obsolete, do not use */
-
-	TCA_CN_FLOW_MAX_RATE,	/* per flow max rate */
-
-	TCA_CN_BUCKETS_LOG,	/* log2(number of buckets) */
-
-	TCA_CN_FLOW_REFILL_DELAY,	/* flow credit refill delay in usec */
-
-	TCA_CN_ORPHAN_MASK,	/* mask applied to orphaned skb hashes */
-
-	TCA_CN_LOW_RATE_THRESHOLD, /* per packet delay under this rate */
-
-	TCA_CN_CE_THRESHOLD,	/* DCTCP-like CE-marking threshold */
-
-	TCA_CN_GUARD_INTERVAL,
-
-	TCA_CN_MAX_INCREASE,
-
-	TCA_CN_MAX_MONITORING_INTERVAL,
-
-	__TCA_CN_MAX,
-};
-
-#define TCA_CN_MAX	(__TCA_CN_MAX - 1)
-
-#endif

### 2019-10-22 iproute2/tc/cocoa_opts.h

@@ -1,0 +1,46 @@
+/* SPDX-License-Identifier: GPL-2.0 WITH Linux-syscall-note */
+#ifndef __LINUX_COCOA_OPTS_H
+#define __LINUX_COCOA_OPTS_H
+
+// FIXME: This should include a file which defines "TCA_COCOA_MAX", but I couldn't find it.
+// So this only works if you include that mysterious file before this one.
+
+enum {
+	TCA_COCOA_UNSPEC,
+
+	TCA_COCOA_PLIMIT,		/* limit of total number of packets in queue */
+
+	TCA_COCOA_FLOW_PLIMIT,	/* limit of packets per flow */
+
+	TCA_COCOA_QUANTUM,		/* RR quantum */
+
+	TCA_COCOA_INITIAL_QUANTUM,		/* RR quantum for new flow */
+
+	TCA_COCOA_RATE_ENABLE,	/* enable/disable rate limiting */
+
+	TCA_COCOA_FLOW_DEFAULT_RATE,/* obsolete, do not use */
+
+	TCA_COCOA_FLOW_MAX_RATE,	/* per flow max rate */
+
+	TCA_COCOA_BUCKETS_LOG,	/* log2(number of buckets) */
+
+	TCA_COCOA_FLOW_REFILL_DELAY,	/* flow credit refill delay in usec */
+
+	TCA_COCOA_ORPHAN_MASK,	/* mask applied to orphaned skb hashes */
+
+	TCA_COCOA_LOW_RATE_THRESHOLD, /* per packet delay under this rate */
+
+	TCA_COCOA_CE_THRESHOLD,	/* DCTCP-like CE-marking threshold */
+
+	TCA_COCOA_GUARD_INTERVAL,
+
+	TCA_COCOA_MAX_INCREASE,
+
+	TCA_COCOA_MAX_MONITORING_INTERVAL,
+
+	__TCA_COCOA_MAX,
+};
+
+#define TCA_COCOA_MAX	(__TCA_COCOA_MAX - 1)
+
+#endif

### 2019-10-22 iproute2/tc/q_cocoa.c

@@ -47,11 +47,11 @@
 
 #include "utils.h"
 #include "tc_util.h"
-#include "cn_opts.h"
+#include "cocoa_opts.h"
 
 static void explain(void)
 {
-	fprintf(stderr, "Usage: ... cn [ limit PACKETS ] [ flow_limit PACKETS ]\n");
+	fprintf(stderr, "Usage: ... cocoa [ limit PACKETS ] [ flow_limit PACKETS ]\n");
 	fprintf(stderr, "              [ quantum BYTES ] [ initial_quantum BYTES ]\n");
 	fprintf(stderr, "              [ maxrate RATE  ] [ buckets NUMBER ]\n");
 	fprintf(stderr, "              [ [no]pacing ] [ refill_delay TIME ]\n");
@@ -75,7 +75,7 @@ static unsigned int ilog2(unsigned int val)
 	return res;
 }
 
-static int cn_parse_opt(struct qdisc_util *qu, int argc, char **argv,
+static int cocoa_parse_opt(struct qdisc_util *qu, int argc, char **argv,
 			struct nlmsghdr *n, const char *dev)
 {
 	unsigned int plimit;
@@ -242,66 +242,66 @@ static int cn_parse_opt(struct qdisc_util *qu, int argc, char **argv,
 	if (buckets) {
 		unsigned int log = ilog2(buckets);
 
-		addattr_l(n, 1024, TCA_CN_BUCKETS_LOG,
+		addattr_l(n, 1024, TCA_COCOA_BUCKETS_LOG,
 			  &log, sizeof(log));
 	}
 	if (set_plimit)
-		addattr_l(n, 1024, TCA_CN_PLIMIT,
+		addattr_l(n, 1024, TCA_COCOA_PLIMIT,
 			  &plimit, sizeof(plimit));
 	if (set_flow_plimit)
-		addattr_l(n, 1024, TCA_CN_FLOW_PLIMIT,
+		addattr_l(n, 1024, TCA_COCOA_FLOW_PLIMIT,
 			  &flow_plimit, sizeof(flow_plimit));
 	if (set_quantum)
[…]

### 2019-10-22 run_experiments.py

@@ -1,0 +1,24 @@
+#!/usr/bin/env python3
+
+import subprocess
+import argparse
+
+parser = argparse.ArgumentParser()
+parser.add_argument('--number', type=int, default=10)
+parser.add_argument('--time', type=float, default=240)
+parser.add_argument('--cc', type=str, default="cubic")
+parser.add_argument('--qdisc', type=str, default="cn")
+
+opt = parser.parse_args()
+print(opt)
+
+cmd = f'''sudo bash -c 'echo > /sys/kernel/debug/tracing/trace' && sudo bash -c 'echo 1 > /sys/kernel/debug/tracing/tracing_on' && sudo python3 test.py --rate 100 --delay_to_add 50 --time {opt.time} --qdisc {opt.qdisc} --change 1 --cc {opt.cc}'''
+
+for i in range(opt.number):
+	try:
+		output = subprocess.run(cmd, stdout=subprocess.PIPE, stderr=subprocess.STDOUT, check=True, shell=True, timeout=opt.time+100)
+	except subprocess.CalledProcessError as e:
+		print(e.cmd, e.returncode, e.output)
+		raise e
+	except subprocess.TimeoutExpired as e:
+		pass

### 2019-10-22 sch_cocoa.c

@@ -1,5 +1,5 @@
 /*
- * net/sched/sch_cn.c Fair Queue Packet Scheduler (per flow pacing)
+ * net/sched/sch_cocoa.c Fair Queue Packet Scheduler (per flow pacing)
  *
  *  Copyright (C) 2013-2015 Eric Dumazet <edumazet@google.com>
  *
@@ -58,7 +58,7 @@
 #include <linux/timekeeping.h>
 
 // Added by max
-#include "iproute2/tc/cn_opts.h"
+#include "iproute2/tc/cocoa_opts.h"
 
 #define NANOSECONDS_IN_ONE_SECOND 1000000000
 
@@ -87,13 +87,13 @@ struct five_tuple {
 /*
  * Per flow structure, dynamically allocated
  */
-struct cn_flow {
+struct cocoa_flow {
 	struct sk_buff	*head;		/* list of skbs for this flow : first skb */
 	union {
 		struct sk_buff *tail;	/* last skb in the list */
 		unsigned long  age;	/* jiffies when flow was emptied, for gc */
 	};
-	struct rb_node	cn_node;	/* anchor in cn_root[] trees */
+	struct rb_node	cocoa_node;	/* anchor in cocoa_root[] trees */
 	struct sock	*sk;
 	int		qlen;		/* number of packets in flow queue */
 
@@ -101,7 +101,7 @@ struct cn_flow {
 
 	int		credit;
 	u32		socket_hash;	/* sk_hash */
-	struct cn_flow *next;		/* next pointer in RR lists, or &detached */
+	struct cocoa_flow *next;		/* next pointer in RR lists, or &detached */
 	struct rb_node  rate_node;	/* anchor in q->delayed tree */
 	u64		time_next_packet;
 
@@ -117,21 +117,21 @@ struct cn_flow {
 	u64 enlarge;
 };
 
-struct cn_flow_head {
-	struct cn_flow *first;
-	struct cn_flow *last;
+struct cocoa_flow_head {
+	struct cocoa_flow *first;
+	struct cocoa_flow *last;
 };
 
-struct cn_sched_data {
-	struct cn_flow_head new_flows;
+struct cocoa_sched_data {
+	struct cocoa_flow_head new_flows;
 
-	struct cn_flow_head old_flows;
+	struct cocoa_flow_head old_flows;
 
 	struct rb_root	delayed;	/* for rate limited flows */
 	u64		time_next_delayed_flow;
[…]

### 2019-10-22 tex/paper.tex
Rename module to cocoa
Finish renaming

@@
 \newcommand{\algorithmautorefname}{Algorithm}
-\newcommand{\name}{cocoa}
+\newcommand{\codel}{CoDel}
 
@@
 We propose an AQM scheduler based on fair queuing, which adapts the buffer size depending on the needs of each flow without requiring active participation from the endpoint.
-We implement this scheduler for the Linux kernel and show that it interacts well with the most common congestion control algorithms and can significantly increase throughput compared to CoDel while avoiding overbuffering.
+We implement this scheduler for the Linux kernel and show that it interacts well with the most common congestion control algorithms and can significantly increase throughput compared to \codel while avoiding overbuffering.
 \end{abstract}
@@
 
-\newcommand{\codel}{CoDel}
-
 \settopmatter{printfolios=true}
@@
 
-While it is important to keep the queuing delay constrained, it is also necessary to ensure that overly aggressive flows cannot benefit by ``stealing'' less aggressive flows' bandwidth.
-Thus researchers and engineers have developed \textit{\gls{fq}} mechanisms \cite{shreedhar_efficient_1996,dumazet_pkt_sched:_2013} to isolate different flows' queues so that for example a delay-sensitive live video call cannot be impaired by a concurrent bulk transfer which takes all the available bandwidth.
+While it is important to keep the queuing delay constrained, it is also necessary to ensure that overly aggressive flows cannot benefit from ``stealing'' less aggressive flows' bandwidth.
+Thus researchers and engineers have developed \textit{\gls{fq}} mechanisms \cite{shreedhar_efficient_1996,dumazet_pkt_sched:_2013} to isolate different flows' queues so that for example a delay-sensitive live video call cannot be impaired by a concurrent bulk transfer which consumes all the available bandwidth.
 
@@
 We show that the impaired performance is a result of keeping the queuing delay under 5\;ms, which hinders \glspl{cca} such as Reno or Cubic from reaching maximum throughput.
-The main drawback is that these \glspl{qdisc} aim to keep the delay under the threshold no matter the effect on throughput and do not take the \gls{cc} of the flow into account.
+Moreover, this behavior can result in unnecessary standing queues on links with very low latencies under 10\;ms, which occur in data centers and between users and close by servers of content delivery networks.
 
-As a remedy, we conceive an \gls{aqm} mechanism that explicitly measures the behavior of the \gls{cc} of a flow and dynamically changes the buffer so that 
+As a remedy, we conceive a fair \gls{aqm} mechanism that explicitly measures the behavior of the \gls{cc} of a flow and dynamically changes the buffer so that 
 \begin{enumerate}[topsep=0pt,wide,labelwidth=!,labelindent=0pt]
-\item link utilization is optimized and
+\item link utilization is maximized and
 \item queuing delay is kept at the minimum that is required to achieve optimum throughput considering the \gls{cc}.
 \end{enumerate}
-Finally, we develop a prototype of \gls{cocoa} and show that it can achieve these objectives for the most common loss-based \glspl{cca}, Reno \cite {jacobson_congestion_1988} and Cubic \cite{ha_cubic:_2008}.
+Finally, we develop a prototype of the \gls{cocoa} and show that it can achieve these objectives for the most common loss-based \glspl{cca}, Reno \cite {jacobson_congestion_1988} and Cubic \cite{ha_cubic:_2008}.
 These \glspl{cca} operate by continuously increasing the number of bytes that are allowed to be in the network (congestion window) if no packet loss is experienced and by sharply decreasing this number if a packet is lost (multiplicative decrease).
@@
 BBR is thus considered to be \textit{model-based}.
-We demonstrate that \gls{cocoa} also behaves well in interaction with BBR~v1.
+We show that \gls{cocoa} also behaves well in interaction with BBR~v1.
 
@@
 \includegraphics[width=\columnwidth]{figures/cocoa_illustration_too_much.pdf}
-\caption{If the buffer is too large, loss-based \glspl{cca} will keep a unnecessary standing queue which is not necessary for achieving full link utilization.}
+\caption{If the buffer is too large, loss-based \glspl{cca} will keep an unnecessary standing queue which is not necessary for achieving full link utilization.}
 \label{fig:tooMuch}
@@
 \begin{enumerate}[topsep=0pt,wide,labelwidth=!,labelindent=0pt]
-\item observe the \gls{cca} for a certain measurement period and then
-\item increase or decrease the buffer size depending on whether the scenario of \autoref{fig:tooLittle} or \autoref{fig:tooMuch} is true. 
+\item observe the \gls{cca} of each flow for a certain measurement period and
+\item increase or decrease the buffer size for this flow depending on whether the scenario of \autoref{fig:tooLittle} or \autoref{fig:tooMuch} is true. 
 \end{enumerate}
@@
 For example, we choose the multiplier to be $\frac{1}{2}$.
-Now, if the \gls{li} is in the previous \gls{gi} was 100\,ms, then the next \gls{gi} is at least $\frac{1}{2} 100\,ms$.
+Now, if the \gls{li} in the previous \gls{gi} was 100\,ms, then the next \gls{gi} is at least $\frac{1}{2} 100\,ms$.
 During the \gls{gi} we monitor all intervals and if the current interval is the longest one, it becomes the new \gls{li}.
-The \gls{gi} ends when the first packet loss occurs after the end of the \gls{gi}.
+The \gls{gi} ends when the first packet loss occurs after the minimum duration of the \gls{gi} elapsed.
 
@@
 				\State buffer\_size $\gets$ buffer\_size - standing\_queue
+				\State drop superfluous packets
 			\EndIf
@@
 We thus conjecture that a multiplier of $0.5$ is sufficient for BBR~v2.
-However, we could not finally confirm this since BBR~v2 is still not completely finished.
+However, we could not finally confirm this since BBR~v2 is still under development.
 
@@
 This means that we also assume that no flows are handled by \gls{cocoa} which have an \gls{rtt} larger than 1\,s.
-Another possibility for setting this parameter is to set it to a multiple of the \gls{rtt}, for example 2, 4 or 8 times the most recent \gls{rtt}.
+Another possibility for setting this parameter is to use a multiple of the \gls{rtt}, for example 2, 4 or 8 times the most recent \gls{rtt}.
 \gls{rtt} could be measured using TCP timestamps \cite{borman_tcp_2014} in case of TCP flows or the spin bit \cite{kuhlewind_quic_2018} for QUIC.
@@
 The testbed consists of a sender and a receiver connected via a switch.
-We runs the \gls{qdisc} being tested on the interface that connects the switch to the receiver.
+We run the \gls{qdisc} being tested on the interface that connects the switch to the receiver.
 We introduce delay with \texttt{netem} and limit bandwidth with \texttt{htb}.
@@
 \caption{Throughput and delay of a Cubic flow with \gls{cocoa} at the bottleneck.
-The initial bandwidth is 20\,Mbit/s but we double after 30\,s.
+The initial bandwidth is 20\,Mbit/s but we double it after 30\,s.
 The delay is 10\,ms.
@@
 As can be seen in \autoref{fig:fq} the regular fq qdisc maintains a standing queue and unnecessarily leads to a significant increase in \gls{rtt} (100\,ms minimum, when the real minimum is 10\,ms).
-In contrast, \autoref{fig:cn} maintains the full throughput while at the same time keeping the delay minimal and not keeping a standing queue.
+In contrast, \autoref{fig:cn} maintains the full throughput while at the same time keeping the delay minimal and not having a standing queue.
 Also, when drastically changing the bandwidth by halving it (\autoref{fig:cn}) or doubling it (\autoref{fig:cn2}) \gls{cocoa} rapidly adapts to the new conditions and returns to the optimum state.
@@
 \autoref{fig:fqCodel} shows that fq\_codel keeps the queuing delay under 5\,ms.
-Also it shows that keeping the queuing delay that low is detrimental for achieving full throughput when using the Cubic \gls{cca}.
+It also shows that keeping the queuing delay that low is detrimental for achieving full throughput when using the Cubic \gls{cca}.
 With \gls{cocoa} we achieve throughput that is more than 10\% higher overall, while only slightly increasing the average delay.

### 2019-10-22 test.py

@@ -39,7 +39,7 @@ parser.add_argument('--rate', type=int, default=20)
 parser.add_argument('--mtu', type=int, default=1514)
 parser.add_argument('--time', type=int, default=60)
 parser.add_argument('--change', type=float, default=0.5)
-parser.add_argument('--qdisc', type=str, default="cn")
+parser.add_argument('--qdisc', type=str, default="cocoa")
 parser.add_argument('--cc', type=str, default="cubic")
 parser.add_argument('--thing_to_change', type=str, default="bw")
 
@@ -55,7 +55,7 @@ print("BDP_packets", BDP_packets)
 # BDP_packets = 501
 # BDP_packets = 1
 
-# if opt.qdisc=="cn":
+# if opt.qdisc=="cocoa":
 # 	print("actual packets", BDP_packets)
 
 def run_commands(cmds, Popen=False):
@@ -84,7 +84,7 @@ def run_commands(cmds, Popen=False):
 env_with_tc = os.environ.copy()
 # Idiotic
 # env_with_tc["TC_LIB_DIR"] = os.path.expanduser('~/repos/iproute2/tc')
-if opt.qdisc=="cn":
+if opt.qdisc=="cocoa":
 	env_with_tc["TC_LIB_DIR"] = "/home/max/repos/cocoa-qdisc/iproute2/tc"
 
 # print("os.environ", os.environ)
@@ -134,7 +134,7 @@ def run(vnet):
 			execute_popen_and_show_result(f"ethtool -K {interface} gso off")
 			execute_popen_and_show_result(f"ethtool -K {interface} tso off")
 
[…]

### 2019-10-23 tex/paper.tex
Typos and add missing reference

@@
 %% allowing the author to define a "short title" to be used in page headers.
-\title{Congestion Control Aware Queuing}
+\title{Cocoa: Congestion Control Aware Queuing}
 
@@
 Conversely, applying this model-based approach to Active Queue Management (AQM) has so far received less attention.
-We propose an AQM scheduler based on fair queuing, which adapts the buffer size depending on the needs of each flow without requiring active participation from the endpoint.
+We propose the new AQM scheduler \textit{cocoa} based on fair queuing, which adapts the buffer size depending on the needs of each flow without requiring active participation from the endpoint.
 We implement this scheduler for the Linux kernel and show that it interacts well with the most common congestion control algorithms and can significantly increase throughput compared to \codel while avoiding overbuffering.
@@
 This becomes especially prevalent for links with a high bandwidth or a large \gls{rtt} but is already noticeable for common scenarios, such as a link with 100\;Mbit/s and an \gls{rtt} of 50\;ms.
-
 We show that the impaired performance is a result of keeping the queuing delay under 5\;ms, which hinders \glspl{cca} such as Reno or Cubic from reaching maximum throughput.
@@
 \end{enumerate}
-Finally, we develop a prototype of the \gls{cocoa} and show that it can achieve these objectives for the most common loss-based \glspl{cca}, Reno \cite {jacobson_congestion_1988} and Cubic \cite{ha_cubic:_2008}.
+The concept of dynamically adjusting the buffer size depending on flows' needs was recently proposed by \cite{bless_policy-oriented_2018}.
+However, their solution is not tailored for fair queuing but for flows sharing a queue.
+
+We show that a prototype of our solution \gls{cocoa} can achieve these objectives for the most common loss-based \glspl{cca}, Reno \cite {jacobson_congestion_1988} and Cubic \cite{ha_cubic:_2008}.
 These \glspl{cca} operate by continuously increasing the number of bytes that are allowed to be in the network (congestion window) if no packet loss is experienced and by sharply decreasing this number if a packet is lost (multiplicative decrease).
@@
 \includegraphics[width=\columnwidth]{figures/cocoa_illustration_too_little.pdf}
-\caption{If the buffer is too small, loss-based \glspl{cca} will not be able to fully utilize the link since they send too few data following the multiplicative decrease that occurs after packet loss.}
+\caption{If the buffer is too small, loss-based \glspl{cca} cannot fully utilize the link since they send too few data following the multiplicative decrease that occurs after packet loss.}
 \label{fig:tooLittle}
@@
 \includegraphics[width=\columnwidth]{figures/cocoa_illustration_too_much.pdf}
-\caption{If the buffer is too large, loss-based \glspl{cca} will keep an unnecessary standing queue which is not necessary for achieving full link utilization.}
+\caption{If the buffer is too large, loss-based \glspl{cca} keep an unnecessary standing queue not required for achieving full link utilization.}
 \label{fig:tooMuch}
@@
 
-This is not so trivial because of the following consideration: If the measurement period is mistakenly not the longest interval shown in the figure, but the interval between two adjacent packet losses, the algorithm would assume an enormous standing queue and drastically reduce the buffer to be close to zero.
-This would then lead to severe underutilization afterwards.
+This is not trivial because of the following consideration: If the measurement period is mistakenly not the longest interval shown in the figure, but the interval between two adjacent packet losses, the algorithm would assume an enormous standing queue and drastically reduce the buffer to be close to zero.
+This would lead to severe underutilization.
 
@@
 From these considerations it follows that for our experiments we generally use a multiplier of $1.25$.
-The behavior is different for BBR~v2 which drastically changed the probing phase and now consists of a random 2-3\,s phase of constant sending rate (cruising) followed by a quick probing phase followed by a decrease phase \cite{cardwell_bbr_2019-1,cardwell_bbr_2019}.
+The behavior is different for BBR~v2 which drastically changed the probing phase and now consists of a random 2-3\,s phase of constant sending rate (cruising) followed by a quick probing phase followed by a decrease phase \cite{cardwell_bbr_2019}.
 We thus conjecture that a multiplier of $0.5$ is sufficient for BBR~v2.
@@
 %}{}
-\caption{Throughput and delay of a Cubic flow with fq at the bottleneck.
+\caption{A Cubic flow with fq at the bottleneck.
 The initial bandwidth is 20\,Mbit/s but it is halved after 30\,s.
@@
 %}{}
-\caption{Throughput and delay of a Cubic flow with \gls{cocoa} at the bottleneck.
+\caption{A Cubic flow with \gls{cocoa} at the bottleneck.
 The initial bandwidth is 20\,Mbit/s but we halve it after 30\,s.
 The delay is 10\,ms.
-\gls{cocoa} keeps the queuing delay minimal while achieving the same throughput as with an oversized buffer.}
+\gls{cocoa} keeps the queue smaller than fq while achieving the same throughput.}
 \label{fig:cn}
@@
 %}{}
-\caption{Throughput and delay of a Cubic flow with \gls{cocoa} at the bottleneck.
+\caption{A Cubic flow with \gls{cocoa} at the bottleneck.
 The initial bandwidth is 20\,Mbit/s but we double it after 30\,s.
@@
 %}{}
-\caption{Throughput and delay of a Cubic flow with fq\_codel at the bottleneck. The bandwidth is 100\,Mbit/s. The delay is 100\,ms. The average \gls{rtt} is 101.61\,ms. Total throughput is 2615\,MB.}
+\caption{A Cubic flow with fq\_codel at the bottleneck. The bandwidth is 100\,Mbit/s. The delay is 100\,ms. The average \gls{rtt} is 101.61\,ms. Total throughput is 2615\,MB.}
 \label{fig:fqCodel}
@@
 %}{}
-\caption{Throughput and delay of a Cubic flow with \gls{cocoa} at the bottleneck.
-The bandwidth is 100\,Mbit/s.
-The delay is 100\,ms.
-The average \gls{rtt} is 130.36\,ms.
-Total throughput is 2893\,MB.}
+\caption{A Cubic flow with \gls{cocoa} at the bottleneck. The bandwidth is 100\,Mbit/s. The delay is 100\,ms. The average \gls{rtt} is 130.36\,ms. Total throughput is 2893\,MB.}
 \label{fig:cnLong}
@@
 For Cubic fq achieves good performance.
-This is because its default buffer size of 100 packets coincidentially works well for this specific link speed and \gls{rtt}.} \label{tab:performance_results}
+This is because its default buffer size of 100 packets coincidentially works well for this specific link speed and \gls{rtt}.
+We take the mean of 10 runs for each scenario.} \label{tab:performance_results}
 \begin{tabular}{l r r r r r r} \toprule
@@
 %}{}
-\caption{Throughput and delay of a BBR~v1 flow with \gls{cocoa} at the bottleneck. The bandwidth is 50\,Mbit/s. The delay is 10\,ms and we halve it after 30\,s.}
+\caption{A BBR~v1 flow with \gls{cocoa} at the bottleneck. The bandwidth is 50\,Mbit/s. The delay is 10\,ms and we halve it after 30\,s.}
 \label{fig:cnBBR}
@@
 Our goal was to design a \gls{qdisc} which achieves maximum throughput while keeping the queuing delay as small as possible.
-The results of the experiments, which we perform with a prototype of \gls{cocoa}, show that there are certain common scenarios in which current state-of-the-art \glspl{qdisc} like fq, fq\_codel and cake fall short of the optimum throughput by a significant margin while our approach succeeds in leading to full utilization of the link at the bottleneck (with a potential increase in delay).
+The results of the experiments, which we perform with a prototype of \gls{cocoa}, show that there are certain common scenarios in which current state-of-the-art \glspl{qdisc} like fq, fq\_codel and cake fall short of the optimum throughput by a significant margin while our approach succeeds in fully utilizing the bottleneck link (with a potential increase in delay).
 Furthermore, the aforementioned \glspl{qdisc} also suffer from standing queues in scenarios with small \gls{bdp} which we can also mitigate with our approach.
 
-We see this work as a initial step towards flow-adaptive queuing at bottlenecks.
+We see this work as an initial step towards flow-adaptive queuing at bottlenecks.
 An interesting further improvement could be to not statically initialize the buffer with a constant 100 packets but instead use experience from previous flows.

### 2019-10-25 tex/paper.tex
Add wintracker

@@
 
-%\usepackage{multirow}
-%\usepackage{blindtext}
-%\usepackage{soul}
 \usepackage{enumitem}
@@
 %]{
-\includegraphics[width=\columnwidth]{{{../plots/throughput_1_fq_cubic_10_20_60_0.5_bw_1571240937211.pcap}}}
+\includegraphics[width=\columnwidth]{{{plots/throughput_1_fq_cubic_10_20_60_0.5_bw_1571240937211.pcap}}}
 %}{}
@@
 %]{
-\includegraphics[width=\columnwidth]{{{../plots/rtt_1_fq_cubic_10_20_60_0.5_bw_1571240937211.pcap}}}
+\includegraphics[width=\columnwidth]{{{plots/rtt_1_fq_cubic_10_20_60_0.5_bw_1571240937211.pcap}}}
 %}{}
@@
 %]{
-\includegraphics[width=\columnwidth]{{{../plots/throughput_1_cn_cubic_10_20_60_0.5_bw_1571241047054.pcap}}}
+\includegraphics[width=\columnwidth]{{{plots/throughput_1_cn_cubic_10_20_60_0.5_bw_1571241047054.pcap}}}
 %}{}
@@
 %]{
-\includegraphics[width=\columnwidth]{{{../plots/rtt_1_cn_cubic_10_20_60_0.5_bw_1571241047054.pcap}}}
+\includegraphics[width=\columnwidth]{{{plots/rtt_1_cn_cubic_10_20_60_0.5_bw_1571241047054.pcap}}}
 %}{}
@@
 %]{
-\includegraphics[width=\columnwidth]{{{../plots/throughput_1_cn_cubic_10_20_60_2.0_bw_1571241151290.pcap}}}
+\includegraphics[width=\columnwidth]{{{plots/throughput_1_cn_cubic_10_20_60_2.0_bw_1571241151290.pcap}}}
 %}{}
@@
 %]{
-\includegraphics[width=\columnwidth]{{{../plots/rtt_1_cn_cubic_10_20_60_2.0_bw_1571241151290.pcap}}}
+\includegraphics[width=\columnwidth]{{{plots/rtt_1_cn_cubic_10_20_60_2.0_bw_1571241151290.pcap}}}
 %}{}
@@
 %]{
-\includegraphics[width=\columnwidth]{{{../plots/throughput_1_fq_codel_cubic_100_100_240_1.0_bw_1571244312838.pcap}}}
+\includegraphics[width=\columnwidth]{{{plots/throughput_1_fq_codel_cubic_100_100_240_1.0_bw_1571244312838.pcap}}}
 %}{}
@@
 %]{
-\includegraphics[width=\columnwidth]{{{../plots/rtt_1_fq_codel_cubic_100_100_240_1.0_bw_1571244312838.pcap}}}
+\includegraphics[width=\columnwidth]{{{plots/rtt_1_fq_codel_cubic_100_100_240_1.0_bw_1571244312838.pcap}}}
 %}{}
@@
 %]{
-\includegraphics[width=\columnwidth]{{{../plots/throughput_1_cn_cubic_100_100_240_1.0_bw_1571244036720.pcap}}}
+\includegraphics[width=\columnwidth]{{{plots/throughput_1_cn_cubic_100_100_240_1.0_bw_1571244036720.pcap}}}
 %}{}
@@
 %]{
-\includegraphics[width=\columnwidth]{{{../plots/rtt_1_cn_cubic_100_100_240_1.0_bw_1571244036720.pcap}}}
+\includegraphics[width=\columnwidth]{{{plots/rtt_1_cn_cubic_100_100_240_1.0_bw_1571244036720.pcap}}}
 %}{}
@@
 %]{
-\includegraphics[width=\columnwidth]{{{../plots/throughput_1_cn_bbr_10_50_60_0.5_delay_1571288916582.pcap}}}
+\includegraphics[width=\columnwidth]{{{plots/throughput_1_cn_bbr_10_50_60_0.5_delay_1571288916582.pcap}}}
 %}{}
@@
 %]{
-\includegraphics[width=\columnwidth]{{{../plots/rtt_1_cn_bbr_10_50_60_0.5_delay_1571288916582.pcap}}}
+\includegraphics[width=\columnwidth]{{{plots/rtt_1_cn_bbr_10_50_60_0.5_delay_1571288916582.pcap}}}
 %}{}

### 2019-10-25 plot_rtt_and_bandwidth.py

@@ -16,7 +16,7 @@ assert pcap_file.startswith("sender_"), pcap_file
 receiver_pcap = 'receiver_'+('_'.join(pcap_file.split('_')[1:]))
 
 # rtt_command = f"tshark -r pcaps/{pcap_file} -Tfields -e frame.time_relative -e tcp.analysis.ack_rtt"
-new_rtt_command = f"../pantheon/tools/wintracker pcaps/{pcap_file}"
+new_rtt_command = f"./wintracker pcaps/{pcap_file}"
 
 retransmissions_command = f"tshark -Y tcp.srcport==60000&&tcp.analysis.retransmission -r pcaps/{pcap_file} -Tfields -e frame.time_relative"
 
@@ -72,7 +72,7 @@ bytes_results = [(float(item[0][0]), float(item[0][1]), float(item[1])) for item
 # print("divided", divided)
 
 with_correct_time = [((item[0] + item[1])/2, item[2]/1000000*8) for item in bytes_results]
-os.makedirs("plots", exist_ok=True)
+os.makedirs("tex/plots", exist_ok=True)
 
 plt.figure(figsize=(5,2))
 plt.xlabel("Time [s]")

### 2019-10-25 wintracker.go

@@ -1,0 +1,727 @@
+// Long term TODO: Calculation of RTT is not correct. Should use previous ACK...
+
+package main
+
+import (
+	"bytes"
+	"encoding/binary"
+	"encoding/csv"
+	"fmt"
+	"io"
+	"log"
+	"math"
+	"net"
+	"os"
+	"strconv"
+	"time"
+
+	"github.com/google/gopacket"
+	"github.com/google/gopacket/layers"
+	"github.com/google/gopacket/pcap"
+
+	"gonum.org/v1/gonum/floats"
+	"gonum.org/v1/gonum/stat"
+)
+
+func floatToString(inputNum float64) string {
+	// to convert a float number to a string
+	return strconv.FormatFloat(inputNum, 'f', -1, 64)
+}
+
+func writeFullArray(resultsArray []map[string]float64, key flowKey, fileName string) {
+	records := make([][]string, 0, len(resultsArray)+1)
+	records = append(records, []string{"window", "rtt", "ackTimestamp", "dataTimestamp", "loss"})
+	for _, v := range resultsArray {
+		records = append(records, []string{floatToString(v["window"]), floatToString(v["rtt"]), floatToString(v["ackTimestamp"]), floatToString(v["dataTimestamp"]), floatToString(v["loss"])})
+	}
+
+	f, err := os.Create(fileName)
+	if err != nil {
+		log.Fatalln("Couldn't open file for writing", err)
+	}
+	defer f.Close()
+	w := csv.NewWriter(f)
+
+	for _, record := range records {
+		if err := w.Write(record); err != nil {
+			log.Fatalln("error writing record to csv:", err)
+		}
+	}
+
+	// Write any buffered data to the underlying writer (standard output).
+	w.Flush()
+
+	if err := w.Error(); err != nil {
+		log.Fatal(err)
+	}
+}
+
+func writeMinimalArray(resultsArray []map[string]float64, key flowKey, fileName string) {
+
+	windows := make([]float64, 0, len(resultsArray))
+	rtts := make([]float64, 0, len(resultsArray))
+	ackTimestamps := make([]float64, 0, len(resultsArray))
+	// dataTimestamps := make([]float64, 0, len(resultsArray))
[…]

### 2019-10-25 tex/paper.tex
Fix hardcoded path in test.py

@@
 
-We show that a prototype of our solution \gls{cocoa} can achieve these objectives for the most common loss-based \glspl{cca}, Reno \cite {jacobson_congestion_1988} and Cubic \cite{ha_cubic:_2008}.
+We show that a prototype of our solution \textit{\gls{cocoa}} can achieve these objectives for the most common loss-based \glspl{cca}, Reno \cite {jacobson_congestion_1988} and Cubic \cite{ha_cubic:_2008}.
 These \glspl{cca} operate by continuously increasing the number of bytes that are allowed to be in the network (congestion window) if no packet loss is experienced and by sharply decreasing this number if a packet is lost (multiplicative decrease).

### 2019-10-25 test.py

@@ -42,6 +42,7 @@ parser.add_argument('--change', type=float, default=0.5)
 parser.add_argument('--qdisc', type=str, default="cocoa")
 parser.add_argument('--cc', type=str, default="cubic")
 parser.add_argument('--thing_to_change', type=str, default="bw")
+parser.add_argument('--path_to_tc_module', type=str, required=True)
 
 opt = parser.parse_args()
 print(opt)
@@ -82,10 +83,8 @@ def run_commands(cmds, Popen=False):
 	return return_stuff
 
 env_with_tc = os.environ.copy()
-# Idiotic
-# env_with_tc["TC_LIB_DIR"] = os.path.expanduser('~/repos/iproute2/tc')
 if opt.qdisc=="cocoa":
-	env_with_tc["TC_LIB_DIR"] = "/home/max/repos/cocoa-qdisc/iproute2/tc"
+	env_with_tc["TC_LIB_DIR"] = opt.path_to_tc_module
 
 # print("os.environ", os.environ)
 

### 2019-11-15 tex/paper.tex
Clean iproute2 folder

@@
 \copyrightyear{2019}
-\acmConference{Workshop on Buffer Sizing}{December 2-3, 2019}{Stanford, California, USA}
+\acmConference[BS'19]{Buffer Sizing Workshop}{December 2-3}{Stanford, CA}
 
@@
 \begin{abstract}
-Recent model-based congestion control algorithms such as BBR use repeated measurements at the end-point to build a model of the network connection and use it to achieve optimal throughput with low queuing delay.
+Recent model-based congestion control algorithms such as BBR use repeated measurements at the endpoint to build a model of the network connection and use it to achieve optimal throughput with low queuing delay.
 Conversely, applying this model-based approach to Active Queue Management (AQM) has so far received less attention.
 We propose the new AQM scheduler \textit{cocoa} based on fair queuing, which adapts the buffer size depending on the needs of each flow without requiring active participation from the endpoint.
-We implement this scheduler for the Linux kernel and show that it interacts well with the most common congestion control algorithms and can significantly increase throughput compared to \codel while avoiding overbuffering.
+We implement this scheduler for the Linux kernel and show that it interacts well with the most common congestion control algorithms and can significantly increase throughput compared to \codel{} while avoiding overbuffering.
 \end{abstract}
@@
 \caption{A Cubic flow with fq at the bottleneck.
-The initial bandwidth is 20\,Mbit/s but it is halved after 30\,s.
+The initial bandwidth is 20\,Mbit/s but we halve it after 30\,s.
 The delay is 10\,ms.
@@
 
-Our goal was to design a \gls{qdisc} which achieves maximum throughput while keeping the queuing delay as small as possible.
-The results of the experiments, which we perform with a prototype of \gls{cocoa}, show that there are certain common scenarios in which current state-of-the-art \glspl{qdisc} like fq, fq\_codel and cake fall short of the optimum throughput by a significant margin while our approach succeeds in fully utilizing the bottleneck link (with a potential increase in delay).
-Furthermore, the aforementioned \glspl{qdisc} also suffer from standing queues in scenarios with small \gls{bdp} which we can also mitigate with our approach.
+Our goal was to design a fair \gls{qdisc} which achieves maximum throughput while keeping the queuing delay as small as possible.
+The results of the experiments, which we perform with a prototype of \gls{cocoa}, show that there are certain common scenarios in which current state-of-the-art fair \glspl{qdisc} like fq, fq\_codel and cake fall short of the optimum throughput by a significant margin while our approach succeeds in fully utilizing the bottleneck link (with a potential increase in delay) (\autoref{tab:performance_results}; \autoref{fig:fqCodel} vs.~\autoref{fig:cnLong}).
+Furthermore, the aforementioned \glspl{qdisc} also suffer from standing queues in scenarios with small \gls{bdp} which we can also mitigate with our approach (\autoref{fig:fq} vs.~\autoref{fig:cn}).
 
-We see this work as an initial step towards flow-adaptive queuing at bottlenecks.
+We see this work as an initial step towards flow-adaptive fair queuing at bottlenecks.
 An interesting further improvement could be to not statically initialize the buffer with a constant 100 packets but instead use experience from previous flows.

### 2019-11-15 iproute2/bridge/fdb.c

@@ -1,545 +1,0 @@
-/* SPDX-License-Identifier: GPL-2.0 */
-/*
- * Get/set/delete fdb table with netlink
- *
- * TODO: merge/replace this with ip neighbour
- *
- * Authors:	Stephen Hemminger <shemminger@vyatta.com>
- */
-
-#include <stdio.h>
-#include <stdlib.h>
-#include <unistd.h>
-#include <netdb.h>
-#include <time.h>
-#include <fcntl.h>
-#include <sys/socket.h>
-#include <sys/time.h>
-#include <net/if.h>
-#include <netinet/in.h>
-#include <linux/if_bridge.h>
-#include <linux/if_ether.h>
-#include <linux/neighbour.h>
-#include <string.h>
-#include <limits.h>
-#include <stdbool.h>
-
-#include "json_print.h"
-#include "libnetlink.h"
-#include "br_common.h"
-#include "rt_names.h"
-#include "utils.h"
-
-static unsigned int filter_index, filter_vlan, filter_state, filter_master;
-
-static void usage(void)
-{
-	fprintf(stderr,
-		"Usage: bridge fdb { add | append | del | replace } ADDR dev DEV\n"
-		"              [ self ] [ master ] [ use ] [ router ] [ extern_learn ]\n"
-		"              [ sticky ] [ local | static | dynamic ] [ dst IPADDR ]\n"
-		"              [ vlan VID ] [ port PORT] [ vni VNI ] [ via DEV ]\n"
-		"              [ src_vni VNI ]\n"
-		"       bridge fdb [ show [ br BRDEV ] [ brport DEV ] [ vlan VID ] [ state STATE ] ]\n");
-	exit(-1);
-}
-
-static const char *state_n2a(unsigned int s)
-{
-	static char buf[32];
-
-	if (s & NUD_PERMANENT)
-		return "permanent";
-
-	if (s & NUD_NOARP)
-		return "static";
-
-	if (s & NUD_STALE)
-		return "stale";
-
-	if (s & NUD_REACHABLE)
-		return "";
-
-	sprintf(buf, "state=%#x", s);
-	return buf;
-}
-
[…]

### 2019-11-15 iproute2/bridge/link.c

@@ -1,585 +1,0 @@
-/* SPDX-License-Identifier: GPL-2.0 */
-
-#include <stdio.h>
-#include <stdlib.h>
-#include <unistd.h>
-#include <time.h>
-#include <sys/socket.h>
-#include <sys/time.h>
-#include <netinet/in.h>
-#include <linux/if.h>
-#include <linux/if_bridge.h>
-#include <string.h>
-#include <stdbool.h>
-
-#include "json_print.h"
-#include "libnetlink.h"
-#include "utils.h"
-#include "br_common.h"
-
-static unsigned int filter_index;
-
-static const char *port_states[] = {
-	[BR_STATE_DISABLED] = "disabled",
-	[BR_STATE_LISTENING] = "listening",
-	[BR_STATE_LEARNING] = "learning",
-	[BR_STATE_FORWARDING] = "forwarding",
-	[BR_STATE_BLOCKING] = "blocking",
-};
-
-static const char *hw_mode[] = {
-	"VEB", "VEPA"
-};
-
-static void print_link_flags(FILE *fp, unsigned int flags, unsigned int mdown)
-{
-	open_json_array(PRINT_ANY, is_json_context() ? "flags" : "<");
-	if (flags & IFF_UP && !(flags & IFF_RUNNING))
-		print_string(PRINT_ANY, NULL,
-			     flags ? "%s," : "%s", "NO-CARRIER");
-	flags &= ~IFF_RUNNING;
-
-#define _PF(f) if (flags&IFF_##f) {					\
-		flags &= ~IFF_##f ;					\
-		print_string(PRINT_ANY, NULL, flags ? "%s," : "%s", #f); }
-	_PF(LOOPBACK);
-	_PF(BROADCAST);
-	_PF(POINTOPOINT);
-	_PF(MULTICAST);
-	_PF(NOARP);
-	_PF(ALLMULTI);
-	_PF(PROMISC);
-	_PF(MASTER);
-	_PF(SLAVE);
-	_PF(DEBUG);
-	_PF(DYNAMIC);
-	_PF(AUTOMEDIA);
-	_PF(PORTSEL);
-	_PF(NOTRAILERS);
-	_PF(UP);
-	_PF(LOWER_UP);
-	_PF(DORMANT);
-	_PF(ECHO);
-#undef _PF
-	if (flags)
-		print_hex(PRINT_ANY, NULL, "%x", flags);
-	if (mdown)
[…]

### 2019-11-15 iproute2/bridge/mdb.c

@@ -1,461 +1,0 @@
-/* SPDX-License-Identifier: GPL-2.0 */
-/*
- * Get mdb table with netlink
- */
-
-#include <stdio.h>
-#include <stdlib.h>
-#include <unistd.h>
-#include <fcntl.h>
-#include <sys/socket.h>
-#include <net/if.h>
-#include <netinet/in.h>
-#include <linux/if_bridge.h>
-#include <linux/if_ether.h>
-#include <string.h>
-#include <arpa/inet.h>
-
-#include "libnetlink.h"
-#include "br_common.h"
-#include "rt_names.h"
-#include "utils.h"
-#include "json_print.h"
-
-#ifndef MDBA_RTA
-#define MDBA_RTA(r) \
-	((struct rtattr *)(((char *)(r)) + NLMSG_ALIGN(sizeof(struct br_port_msg))))
-#endif
-
-static unsigned int filter_index, filter_vlan;
-
-static void usage(void)
-{
-	fprintf(stderr, "Usage: bridge mdb { add | del } dev DEV port PORT grp GROUP [permanent | temp] [vid VID]\n");
-	fprintf(stderr, "       bridge mdb {show} [ dev DEV ] [ vid VID ]\n");
-	exit(-1);
-}
-
-static bool is_temp_mcast_rtr(__u8 type)
-{
-	return type == MDB_RTR_TYPE_TEMP_QUERY || type == MDB_RTR_TYPE_TEMP;
-}
-
-static const char *format_timer(__u32 ticks)
-{
-	struct timeval tv;
-	static char tbuf[32];
-
-	__jiffies_to_tv(&tv, ticks);
-	snprintf(tbuf, sizeof(tbuf), "%4lu.%.2lu",
-		 (unsigned long)tv.tv_sec,
-		 (unsigned long)tv.tv_usec / 10000);
-
-	return tbuf;
-}
-
-static void __print_router_port_stats(FILE *f, struct rtattr *pattr)
-{
-	struct rtattr *tb[MDBA_ROUTER_PATTR_MAX + 1];
-
-	parse_rtattr(tb, MDBA_ROUTER_PATTR_MAX, MDB_RTR_RTA(RTA_DATA(pattr)),
-		     RTA_PAYLOAD(pattr) - RTA_ALIGN(sizeof(uint32_t)));
-
-	if (tb[MDBA_ROUTER_PATTR_TIMER]) {
[…]

### 2019-11-15 iproute2/bridge/vlan.c

@@ -1,693 +1,0 @@
-/* SPDX-License-Identifier: GPL-2.0 */
-#include <stdio.h>
-#include <stdlib.h>
-#include <unistd.h>
-#include <fcntl.h>
-#include <sys/socket.h>
-#include <net/if.h>
-#include <netinet/in.h>
-#include <linux/if_bridge.h>
-#include <linux/if_ether.h>
-#include <string.h>
-
-#include "json_print.h"
-#include "libnetlink.h"
-#include "br_common.h"
-#include "utils.h"
-
-static unsigned int filter_index, filter_vlan;
-static int show_vlan_tunnel_info = 0;
-
-static void usage(void)
-{
-	fprintf(stderr,
-		"Usage: bridge vlan { add | del } vid VLAN_ID dev DEV [ tunnel_info id TUNNEL_ID ]\n"
-		"                                                     [ pvid ] [ untagged ]\n"
-		"                                                     [ self ] [ master ]\n"
-		"       bridge vlan { show } [ dev DEV ] [ vid VLAN_ID ]\n"
-		"       bridge vlan { tunnelshow } [ dev DEV ] [ vid VLAN_ID ]\n");
-	exit(-1);
-}
-
-static int parse_tunnel_info(int *argcp, char ***argvp, __u32 *tun_id_start,
-			     __u32 *tun_id_end)
-{
-	char **argv = *argvp;
-	int argc = *argcp;
-	char *t;
-
-	NEXT_ARG();
-	if (!matches(*argv, "id")) {
-		NEXT_ARG();
-		t = strchr(*argv, '-');
-		if (t) {
-			*t = '\0';
-			if (get_u32(tun_id_start, *argv, 0) ||
-				    *tun_id_start >= 1u << 24)
-				invarg("invalid tun id", *argv);
-			if (get_u32(tun_id_end, t + 1, 0) ||
-				    *tun_id_end >= 1u << 24)
-				invarg("invalid tun id", *argv);
-
-		} else {
-			if (get_u32(tun_id_start, *argv, 0) ||
-				    *tun_id_start >= 1u << 24)
-				invarg("invalid tun id", *argv);
-		}
-	} else {
[…]

### 2019-11-15 iproute2/ip/ip.c

@@ -1,321 +1,0 @@
-/*
- * ip.c		"ip" utility frontend.
- *
- *		This program is free software; you can redistribute it and/or
- *		modify it under the terms of the GNU General Public License
- *		as published by the Free Software Foundation; either version
- *		2 of the License, or (at your option) any later version.
- *
- * Authors:	Alexey Kuznetsov, <kuznet@ms2.inr.ac.ru>
- */
-
-#include <stdio.h>
-#include <stdlib.h>
-#include <unistd.h>
-#include <fcntl.h>
-#include <sys/socket.h>
-#include <netinet/in.h>
-#include <string.h>
-#include <errno.h>
-
-#include "SNAPSHOT.h"
-#include "utils.h"
-#include "ip_common.h"
-#include "namespace.h"
-#include "color.h"
-
-int preferred_family = AF_UNSPEC;
-int human_readable;
-int use_iec;
-int show_stats;
-int show_details;
-int oneline;
-int brief;
-int json;
-int timestamp;
-int force;
-int max_flush_loops = 10;
-int batch_mode;
-bool do_all;
-
-struct rtnl_handle rth = { .fd = -1 };
-
-static void usage(void) __attribute__((noreturn));
-
-static void usage(void)
-{
-	fprintf(stderr,
-"Usage: ip [ OPTIONS ] OBJECT { COMMAND | help }\n"
-"       ip [ -force ] -batch filename\n"
-"where  OBJECT := { link | address | addrlabel | route | rule | neigh | ntable |\n"
-"                   tunnel | tuntap | maddress | mroute | mrule | monitor | xfrm |\n"
-"                   netns | l2tp | fou | macsec | tcp_metrics | token | netconf | ila |\n"
-"                   vrf | sr }\n"
-"       OPTIONS := { -V[ersion] | -s[tatistics] | -d[etails] | -r[esolve] |\n"
-"                    -h[uman-readable] | -iec | -j[son] | -p[retty] |\n"
-"                    -f[amily] { inet | inet6 | mpls | bridge | link } |\n"
-"                    -4 | -6 | -I | -D | -M | -B | -0 |\n"
[…]

### 2019-11-15 iproute2/ip/ip_common.h

@@ -1,159 +1,0 @@
-/* SPDX-License-Identifier: GPL-2.0 */
-#ifndef _IP_COMMON_H_
-#define _IP_COMMON_H_
-
-#include <stdbool.h>
-
-#include "json_print.h"
-
-struct link_filter {
-	int ifindex;
-	int family;
-	int oneline;
-	int showqueue;
-	inet_prefix pfx;
-	int scope, scopemask;
-	int flags, flagmask;
-	int up;
-	char *label;
-	int flushed;
-	char *flushb;
-	int flushp;
-	int flushe;
-	int group;
-	int master;
-	char *kind;
-	char *slave_kind;
-};
-
-int get_operstate(const char *name);
-int print_linkinfo(struct nlmsghdr *n, void *arg);
-int print_addrinfo(struct nlmsghdr *n, void *arg);
-int print_addrlabel(struct nlmsghdr *n, void *arg);
-int print_neigh(struct nlmsghdr *n, void *arg);
-int ipaddr_list_link(int argc, char **argv);
-void ipaddr_get_vf_rate(int, int *, int *, const char *);
-void iplink_usage(void) __attribute__((noreturn));
-
-void iproute_reset_filter(int ifindex);
-void ipmroute_reset_filter(int ifindex);
-void ipaddr_reset_filter(int oneline, int ifindex);
-void ipneigh_reset_filter(int ifindex);
-void ipnetconf_reset_filter(int ifindex);
-
-int print_route(struct nlmsghdr *n, void *arg);
-int print_mroute(struct nlmsghdr *n, void *arg);
-int print_prefix(struct nlmsghdr *n, void *arg);
-int print_rule(struct nlmsghdr *n, void *arg);
-int print_netconf(struct rtnl_ctrl_data *ctrl,
-		  struct nlmsghdr *n, void *arg);
-void netns_map_init(void);
-void netns_nsid_socket_init(void);
-int print_nsid(struct nlmsghdr *n, void *arg);
-char *get_name_from_nsid(int nsid);
-int get_netnsid_from_name(const char *name);
-int set_netnsid_from_name(const char *name, int nsid);
-int do_ipaddr(int argc, char **argv);
-int do_ipaddrlabel(int argc, char **argv);
-int do_iproute(int argc, char **argv);
-int do_iprule(int argc, char **argv);
-int do_ipneigh(int argc, char **argv);
-int do_ipntable(int argc, char **argv);
[…]

### 2019-11-15 iproute2/ip/ipaddress.c

@@ -1,2390 +1,0 @@
-/*
- * ipaddress.c		"ip address".
- *
- *		This program is free software; you can redistribute it and/or
- *		modify it under the terms of the GNU General Public License
- *		as published by the Free Software Foundation; either version
- *		2 of the License, or (at your option) any later version.
- *
- * Authors:	Alexey Kuznetsov, <kuznet@ms2.inr.ac.ru>
- *
- */
-
-#include <stdio.h>
-#include <stdlib.h>
-#include <unistd.h>
-#include <inttypes.h>
-#include <fcntl.h>
-#include <sys/ioctl.h>
-#include <sys/socket.h>
-#include <sys/param.h>
-#include <errno.h>
-#include <netinet/in.h>
-#include <arpa/inet.h>
-#include <string.h>
-#include <fnmatch.h>
-
-#include <linux/netdevice.h>
-#include <linux/if_arp.h>
-#include <linux/sockios.h>
-#include <linux/net_namespace.h>
-
-#include "utils.h"
-#include "rt_names.h"
-#include "utils.h"
-#include "ll_map.h"
-#include "ip_common.h"
-#include "color.h"
-
-enum {
-	IPADD_LIST,
-	IPADD_FLUSH,
-	IPADD_SAVE,
-};
-
-static struct link_filter filter;
-static int do_link;
-
-static void usage(void) __attribute__((noreturn));
-
-static void usage(void)
-{
-	if (do_link)
-		iplink_usage();
-
-	fprintf(stderr, "Usage: ip address {add|change|replace} IFADDR dev IFNAME [ LIFETIME ]\n");
-	fprintf(stderr, "                                                      [ CONFFLAG-LIST ]\n");
-	fprintf(stderr, "       ip address del IFADDR dev IFNAME [mngtmpaddr]\n");
-	fprintf(stderr, "       ip address {save|flush} [ dev IFNAME ] [ scope SCOPE-ID ]\n");
-	fprintf(stderr, "                            [ to PREFIX ] [ FLAG-LIST ] [ label LABEL ] [up]\n");
[…]

### 2019-11-15 iproute2/ip/ipila.c

@@ -1,307 +1,0 @@
-/*
- * ipila.c	ILA (Identifier Locator Addressing) support
- *
- *              This program is free software; you can redistribute it and/or
- *              modify it under the terms of the GNU General Public License
- *              as published by the Free Software Foundation; either version
- *              2 of the License, or (at your option) any later version.
- *
- * Authors:	Tom Herbert <tom@herbertland.com>
- */
-
-#include <netdb.h>
-#include <stdio.h>
-#include <stdlib.h>
-#include <string.h>
-#include <net/if.h>
-#include <linux/ila.h>
-#include <linux/genetlink.h>
-#include <linux/ip.h>
-#include <arpa/inet.h>
-
-#include "libgenl.h"
-#include "utils.h"
-#include "ip_common.h"
-#include "ila_common.h"
-#include "json_print.h"
-
-static void usage(void)
-{
-	fprintf(stderr,
-"Usage: ip ila add loc_match LOCATOR_MATCH loc LOCATOR [ dev DEV ] OPTIONS\n"
-"       ip ila del loc_match LOCATOR_MATCH [ loc LOCATOR ] [ dev DEV ]\n"
-"       ip ila list\n"
-"OPTIONS := [ csum-mode { adj-transport | neutral-map | neutral-map-auto | no-action } ]\n"
-"           [ ident-type { luid | use-format } ]\n");
-
-	exit(-1);
-}
-
-/* netlink socket */
-static struct rtnl_handle genl_rth = { .fd = -1 };
-static int genl_family = -1;
-
-#define ILA_REQUEST(_req, _bufsiz, _cmd, _flags)	\
-	GENL_REQUEST(_req, _bufsiz, genl_family, 0,	\
-		     ILA_GENL_VERSION, _cmd, _flags)
-
-#define ILA_RTA(g) ((struct rtattr *)(((char *)(g)) +	\
-	NLMSG_ALIGN(sizeof(struct genlmsghdr))))
-
-static void print_addr64(__u64 addr, char *buff, size_t len)
-{
-	__u16 *words = (__u16 *)&addr;
-	__u16 v;
-	int i, ret;
-	size_t written = 0;
[…]

### 2019-11-15 iproute2/ip/iplink_bond.c

@@ -1,819 +1,0 @@
-/*
- * iplink_bond.c	Bonding device support
- *
- *              This program is free software; you can redistribute it and/or
- *              modify it under the terms of the GNU General Public License
- *              as published by the Free Software Foundation; either version
- *              2 of the License, or (at your option) any later version.
- *
- * Authors:     Jiri Pirko <jiri@resnulli.us>
- *              Scott Feldman <sfeldma@cumulusnetworks.com>
- */
-
-#include <stdio.h>
-#include <stdlib.h>
-#include <string.h>
-#include <linux/if_bonding.h>
-
-#include "rt_names.h"
-#include "utils.h"
-#include "ip_common.h"
-#include "json_print.h"
-
-#define BOND_MAX_ARP_TARGETS    16
-
-static unsigned int xstats_print_attr;
-static int filter_index;
-
-static const char *mode_tbl[] = {
-	"balance-rr",
-	"active-backup",
-	"balance-xor",
-	"broadcast",
-	"802.3ad",
-	"balance-tlb",
-	"balance-alb",
-	NULL,
-};
-
-static const char *arp_validate_tbl[] = {
-	"none",
-	"active",
-	"backup",
-	"all",
-	NULL,
-};
-
-static const char *arp_all_targets_tbl[] = {
-	"any",
-	"all",
-	NULL,
-};
-
-static const char *primary_reselect_tbl[] = {
-	"always",
-	"better",
-	"failure",
-	NULL,
-};
-
-static const char *fail_over_mac_tbl[] = {
-	"none",
-	"active",
-	"follow",
-	NULL,
-};
-
-static const char *xmit_hash_policy_tbl[] = {
-	"layer2",
-	"layer3+4",
-	"layer2+3",
-	"encap2+3",
-	"encap3+4",
-	NULL,
-};
-
-static const char *lacp_rate_tbl[] = {
-	"slow",
-	"fast",
-	NULL,
-};
-
-static const char *ad_select_tbl[] = {
-	"stable",
-	"bandwidth",
-	"count",
-	NULL,
-};
-
-static const char *get_name(const char **tbl, int index)
-{
-	int i;
-
-	for (i = 0; tbl[i]; i++)
[…]

### 2019-11-15 iproute2/ip/iplink_bond_slave.c

@@ -1,161 +1,0 @@
-/*
- * iplink_bond_slave.c	Bonding slave device support
- *
- *              This program is free software; you can redistribute it and/or
- *              modify it under the terms of the GNU General Public License
- *              as published by the Free Software Foundation; either version
- *              2 of the License, or (at your option) any later version.
- *
- * Authors:     Jiri Pirko <jiri@resnulli.us>
- */
-
-#include <stdio.h>
-#include <sys/socket.h>
-#include <linux/if_bonding.h>
-
-#include "rt_names.h"
-#include "utils.h"
-#include "ip_common.h"
-
-static void print_explain(FILE *f)
-{
-	fprintf(f, "Usage: ... bond_slave [ queue_id ID ]\n");
-}
-
-static void explain(void)
-{
-	print_explain(stderr);
-}
-
-static const char *slave_states[] = {
-	[BOND_STATE_ACTIVE] = "ACTIVE",
-	[BOND_STATE_BACKUP] = "BACKUP",
-};
-
-static void print_slave_state(FILE *f, struct rtattr *tb)
-{
-	unsigned int state = rta_getattr_u8(tb);
-
-	if (state >= ARRAY_SIZE(slave_states))
-		print_int(PRINT_ANY, "state_index", "state %d ", state);
-	else
-		print_string(PRINT_ANY,
-			     "state",
-			     "state %s ",
-			     slave_states[state]);
-}
-
-static const char *slave_mii_status[] = {
-	[BOND_LINK_UP] = "UP",
-	[BOND_LINK_FAIL] = "GOING_DOWN",
-	[BOND_LINK_DOWN] = "DOWN",
-	[BOND_LINK_BACK] = "GOING_BACK",
-};
-
-static void print_slave_mii_status(FILE *f, struct rtattr *tb)
-{
-	unsigned int status = rta_getattr_u8(tb);
-
-	if (status >= ARRAY_SIZE(slave_mii_status))
-		print_int(PRINT_ANY,
-			  "mii_status_index",
-			  "mii_status %d ",
-			  status);
-	else
-		print_string(PRINT_ANY,
[…]

### 2019-11-15 iproute2/ip/iplink_bridge.c

@@ -1,871 +1,0 @@
-/*
- * iplink_bridge.c	Bridge device support
- *
- *              This program is free software; you can redistribute it and/or
- *              modify it under the terms of the GNU General Public License
- *              as published by the Free Software Foundation; either version
- *              2 of the License, or (at your option) any later version.
- *
- * Authors:     Jiri Pirko <jiri@resnulli.us>
- */
-
-#include <stdio.h>
-#include <stdlib.h>
-#include <string.h>
-#include <netinet/in.h>
-#include <netinet/ether.h>
-#include <linux/if_link.h>
-#include <linux/if_bridge.h>
-#include <net/if.h>
-
-#include "rt_names.h"
-#include "utils.h"
-#include "ip_common.h"
-
-static unsigned int xstats_print_attr;
-static int filter_index;
-
-static void print_explain(FILE *f)
-{
-	fprintf(f,
-		"Usage: ... bridge [ fdb_flush ]\n"
-		"                  [ forward_delay FORWARD_DELAY ]\n"
-		"                  [ hello_time HELLO_TIME ]\n"
-		"                  [ max_age MAX_AGE ]\n"
-		"                  [ ageing_time AGEING_TIME ]\n"
-		"                  [ stp_state STP_STATE ]\n"
-		"                  [ priority PRIORITY ]\n"
-		"                  [ group_fwd_mask MASK ]\n"
-		"                  [ group_address ADDRESS ]\n"
-		"                  [ vlan_filtering VLAN_FILTERING ]\n"
-		"                  [ vlan_protocol VLAN_PROTOCOL ]\n"
-		"                  [ vlan_default_pvid VLAN_DEFAULT_PVID ]\n"
-		"                  [ vlan_stats_enabled VLAN_STATS_ENABLED ]\n"
-		"                  [ vlan_stats_per_port VLAN_STATS_PER_PORT ]\n"
-		"                  [ mcast_snooping MULTICAST_SNOOPING ]\n"
-		"                  [ mcast_router MULTICAST_ROUTER ]\n"
[…]

### 2019-11-15 iproute2/ip/iplink_bridge_slave.c

@@ -1,447 +1,0 @@
-/*
- * iplink_bridge_slave.c	Bridge slave device support
- *
- *              This program is free software; you can redistribute it and/or
- *              modify it under the terms of the GNU General Public License
- *              as published by the Free Software Foundation; either version
- *              2 of the License, or (at your option) any later version.
- *
- * Authors:     Jiri Pirko <jiri@resnulli.us>
- */
-
-#include <stdio.h>
-#include <sys/socket.h>
-#include <netinet/in.h>
-#include <linux/if_link.h>
-#include <linux/if_bridge.h>
-
-#include "rt_names.h"
-#include "utils.h"
-#include "ip_common.h"
-
-static void print_explain(FILE *f)
-{
-	fprintf(f,
-		"Usage: ... bridge_slave [ fdb_flush ]\n"
-		"                        [ state STATE ]\n"
-		"                        [ priority PRIO ]\n"
-		"                        [ cost COST ]\n"
-		"                        [ guard {on | off} ]\n"
-		"                        [ hairpin {on | off} ]\n"
-		"                        [ fastleave {on | off} ]\n"
-		"                        [ root_block {on | off} ]\n"
-		"                        [ learning {on | off} ]\n"
-		"                        [ flood {on | off} ]\n"
-		"                        [ proxy_arp {on | off} ]\n"
-		"                        [ proxy_arp_wifi {on | off} ]\n"
-		"                        [ mcast_router MULTICAST_ROUTER ]\n"
-		"                        [ mcast_fast_leave {on | off} ]\n"
-		"                        [ mcast_flood {on | off} ]\n"
-		"                        [ mcast_to_unicast {on | off} ]\n"
-		"                        [ group_fwd_mask MASK ]\n"
[…]

### 2019-11-15 iproute2/ip/iplink_geneve.c

@@ -1,386 +1,0 @@
-/*
- * iplink_geneve.c	GENEVE device support
- *
- *              This program is free software; you can redistribute it and/or
- *              modify it under the terms of the GNU General Public License
- *              as published by the Free Software Foundation; either version
- *              2 of the License, or (at your option) any later version.
- *
- * Authors:     John W. Linville <linville@tuxdriver.com>
- */
-
-#include <stdio.h>
-
-#include "rt_names.h"
-#include "utils.h"
-#include "ip_common.h"
-
-#define GENEVE_ATTRSET(attrs, type) (((attrs) & (1L << (type))) != 0)
-
-static void print_explain(FILE *f)
-{
-	fprintf(f,
-		"Usage: ... geneve id VNI\n"
-		"                  remote ADDR\n"
-		"                  [ ttl TTL ]\n"
-		"                  [ tos TOS ]\n"
-		"                  [ df DF ]\n"
-		"                  [ flowlabel LABEL ]\n"
-		"                  [ dstport PORT ]\n"
-		"                  [ [no]external ]\n"
-		"                  [ [no]udpcsum ]\n"
-		"                  [ [no]udp6zerocsumtx ]\n"
-		"                  [ [no]udp6zerocsumrx ]\n"
-		"\n"
-		"Where: VNI   := 0-16777215\n"
-		"       ADDR  := IP_ADDRESS\n"
-		"       TOS   := { NUMBER | inherit }\n"
-		"       TTL   := { 1..255 | auto | inherit }\n"
-		"       DF    := { unset | set | inherit }\n"
-		"       LABEL := 0-1048575\n"
-	);
-}
-
-static void explain(void)
-{
-	print_explain(stderr);
-}
-
-static void check_duparg(__u64 *attrs, int type, const char *key,
-			 const char *argv)
-{
-	if (!GENEVE_ATTRSET(*attrs, type)) {
-		*attrs |= (1L << type);
-		return;
-	}
[…]

### 2019-11-15 iproute2/ip/iplink_vlan.c

@@ -1,285 +1,0 @@
-/*
- * iplink_vlan.c	VLAN device support
- *
- *              This program is free software; you can redistribute it and/or
- *              modify it under the terms of the GNU General Public License
- *              as published by the Free Software Foundation; either version
- *              2 of the License, or (at your option) any later version.
- *
- * Authors:     Patrick McHardy <kaber@trash.net>
- */
-
-#include <stdio.h>
-#include <stdlib.h>
-#include <string.h>
-#include <linux/if_vlan.h>
-
-#include "rt_names.h"
-#include "utils.h"
-#include "ip_common.h"
-
-static void print_explain(FILE *f)
-{
-	fprintf(f,
-		"Usage: ... vlan id VLANID\n"
-		"                [ protocol VLANPROTO ]\n"
-		"                [ reorder_hdr { on | off } ]\n"
-		"                [ gvrp { on | off } ]\n"
-		"                [ mvrp { on | off } ]\n"
-		"                [ loose_binding { on | off } ]\n"
-		"                [ bridge_binding { on | off } ]\n"
-		"                [ ingress-qos-map QOS-MAP ]\n"
-		"                [ egress-qos-map QOS-MAP ]\n"
-		"\n"
-		"VLANID := 0-4095\n"
-		"VLANPROTO: [ 802.1Q | 802.1ad ]\n"
-		"QOS-MAP := [ QOS-MAP ] QOS-MAPPING\n"
-		"QOS-MAPPING := FROM:TO\n"
-	);
-}
-
-static void explain(void)
-{
-	print_explain(stderr);
-}
-
-static int on_off(const char *msg, const char *arg)
-{
-	fprintf(stderr, "Error: argument of \"%s\" must be \"on\" or \"off\", not \"%s\"\n", msg, arg);
-	return -1;
-}
-
-static int vlan_parse_qos_map(int *argcp, char ***argvp, struct nlmsghdr *n,
-			      int attrtype)
-{
-	int argc = *argcp;
-	char **argv = *argvp;
-	struct ifla_vlan_qos_mapping m;
[…]

### 2019-11-15 iproute2/ip/ipmonitor.c

@@ -1,326 +1,0 @@
-/*
- * ipmonitor.c		"ip monitor".
- *
- *		This program is free software; you can redistribute it and/or
- *		modify it under the terms of the GNU General Public License
- *		as published by the Free Software Foundation; either version
- *		2 of the License, or (at your option) any later version.
- *
- * Authors:	Alexey Kuznetsov, <kuznet@ms2.inr.ac.ru>
- *
- */
-
-#include <stdio.h>
-#include <stdlib.h>
-#include <unistd.h>
-#include <fcntl.h>
-#include <sys/socket.h>
-#include <netinet/in.h>
-#include <arpa/inet.h>
-#include <string.h>
-#include <time.h>
-
-#include "utils.h"
-#include "ip_common.h"
-
-static void usage(void) __attribute__((noreturn));
-static int prefix_banner;
-int listen_all_nsid;
-
-static void usage(void)
-{
-	fprintf(stderr, "Usage: ip monitor [ all | LISTofOBJECTS ] [ FILE ] [ label ] [all-nsid] [dev DEVICE]\n");
-	fprintf(stderr, "LISTofOBJECTS := link | address | route | mroute | prefix |\n");
-	fprintf(stderr, "                 neigh | netconf | rule | nsid\n");
-	fprintf(stderr, "FILE := file FILENAME\n");
-	exit(-1);
-}
-
-static void print_headers(FILE *fp, char *label, struct rtnl_ctrl_data *ctrl)
-{
-	if (timestamp)
-		print_timestamp(fp);
-
-	if (listen_all_nsid) {
-		if (ctrl == NULL || ctrl->nsid < 0)
-			fprintf(fp, "[nsid current]");
-		else
-			fprintf(fp, "[nsid %d]", ctrl->nsid);
-	}
-
-	if (prefix_banner)
-		fprintf(fp, "%s", label);
-}
-
-static int accept_msg(struct rtnl_ctrl_data *ctrl,
-		      struct nlmsghdr *n, void *arg)
-{
-	FILE *fp = (FILE *)arg;
-
-	switch (n->nlmsg_type) {
-	case RTM_NEWROUTE:
[…]

### 2019-11-15 iproute2/ip/ipmroute.c

@@ -1,338 +1,0 @@
-/*
- * ipmroute.c		"ip mroute".
- *
- *		This program is free software; you can redistribute it and/or
- *		modify it under the terms of the GNU General Public License
- *		as published by the Free Software Foundation; either version
- *		2 of the License, or (at your option) any later version.
- *
- * Authors:	Alexey Kuznetsov, <kuznet@ms2.inr.ac.ru>
- *
- */
-
-#include <stdio.h>
-#include <stdlib.h>
-#include <unistd.h>
-#include <fcntl.h>
-#include <inttypes.h>
-#include <sys/ioctl.h>
-#include <sys/socket.h>
-#include <netinet/in.h>
-#include <arpa/inet.h>
-#include <string.h>
-
-#include <linux/netdevice.h>
-#include <linux/if.h>
-#include <linux/if_arp.h>
-#include <linux/sockios.h>
-
-#include <rt_names.h>
-#include "utils.h"
-#include "ip_common.h"
-#include "json_print.h"
-
-static void usage(void) __attribute__((noreturn));
-
-static void usage(void)
-{
-	fprintf(stderr, "Usage: ip mroute show [ [ to ] PREFIX ] [ from PREFIX ] [ iif DEVICE ]\n");
-	fprintf(stderr, "                      [ table TABLE_ID ]\n");
-	fprintf(stderr, "TABLE_ID := [ local | main | default | all | NUMBER ]\n");
-#if 0
-	fprintf(stderr, "Usage: ip mroute [ add | del ] DESTINATION from SOURCE [ iif DEVICE ] [ oif DEVICE ]\n");
-#endif
-	exit(-1);
-}
-
-struct rtfilter {
-	int tb;
-	int af;
-	int iif;
-	inet_prefix mdst;
-	inet_prefix msrc;
-} filter;
-
-int print_mroute(struct nlmsghdr *n, void *arg)
-{
-	struct rtmsg *r = NLMSG_DATA(n);
-	int len = n->nlmsg_len;
-	struct rtattr *tb[RTA_MAX+1];
-	FILE *fp = arg;
-	const char *src, *dst;
-	SPRINT_BUF(b1);
-	SPRINT_BUF(b2);
-	__u32 table;
-	int iif = 0;
-	int family;
-
[…]

### 2019-11-15 iproute2/ip/ipneigh.c

@@ -1,628 +1,0 @@
-/*
- * ipneigh.c		"ip neigh".
- *
- *		This program is free software; you can redistribute it and/or
- *		modify it under the terms of the GNU General Public License
- *		as published by the Free Software Foundation; either version
- *		2 of the License, or (at your option) any later version.
- *
- * Authors:	Alexey Kuznetsov, <kuznet@ms2.inr.ac.ru>
- *
- */
-
-#include <stdio.h>
-#include <stdlib.h>
-#include <unistd.h>
-#include <fcntl.h>
-#include <string.h>
-#include <sys/time.h>
-#include <sys/socket.h>
-#include <netinet/in.h>
-#include <netinet/ip.h>
-
-#include "rt_names.h"
-#include "utils.h"
-#include "ip_common.h"
-#include "json_print.h"
-
-#define NUD_VALID	(NUD_PERMANENT|NUD_NOARP|NUD_REACHABLE|NUD_PROBE|NUD_STALE|NUD_DELAY)
-#define MAX_ROUNDS	10
-
-static struct
-{
-	int family;
-	int index;
-	int state;
-	int unused_only;
-	inet_prefix pfx;
-	int flushed;
-	char *flushb;
-	int flushp;
-	int flushe;
-	int master;
-	int protocol;
-	__u8 ndm_flags;
-} filter;
-
-static void usage(void) __attribute__((noreturn));
-
-static void usage(void)
-{
-	fprintf(stderr, "Usage: ip neigh { add | del | change | replace }\n"
-			"                { ADDR [ lladdr LLADDR ] [ nud STATE ] | proxy ADDR } [ dev DEV ]\n");
-	fprintf(stderr, "                [ router ] [ extern_learn ] [ protocol PROTO ]\n\n");
-	fprintf(stderr, "       ip neigh { show | flush } [ proxy ] [ to PREFIX ] [ dev DEV ] [ nud STATE ]\n");
-	fprintf(stderr, "                                 [ vrf NAME ]\n\n");
-	fprintf(stderr, "STATE := { permanent | noarp | stale | reachable | none |\n"
[…]

### 2019-11-15 iproute2/ip/ipnetns.c

@@ -1,902 +1,0 @@
-/* SPDX-License-Identifier: GPL-2.0 */
-#define _ATFILE_SOURCE
-#include <sys/types.h>
-#include <sys/stat.h>
-#include <sys/wait.h>
-#include <sys/inotify.h>
-#include <sys/mount.h>
-#include <sys/syscall.h>
-#include <stdio.h>
-#include <string.h>
-#include <sched.h>
-#include <fcntl.h>
-#include <dirent.h>
-#include <errno.h>
-#include <unistd.h>
-#include <ctype.h>
-#include <linux/limits.h>
-
-#include <linux/net_namespace.h>
-
-#include "utils.h"
-#include "list.h"
-#include "ip_common.h"
-#include "namespace.h"
-#include "json_print.h"
-
-static int usage(void)
-{
-	fprintf(stderr, "Usage: ip netns list\n");
-	fprintf(stderr, "       ip netns add NAME\n");
-	fprintf(stderr, "       ip netns attach NAME PID\n");
-	fprintf(stderr, "       ip netns set NAME NETNSID\n");
-	fprintf(stderr, "       ip [-all] netns delete [NAME]\n");
-	fprintf(stderr, "       ip netns identify [PID]\n");
-	fprintf(stderr, "       ip netns pids NAME\n");
-	fprintf(stderr, "       ip [-all] netns exec [NAME] cmd ...\n");
-	fprintf(stderr, "       ip netns monitor\n");
-	fprintf(stderr, "       ip netns list-id\n");
-	fprintf(stderr, "NETNSID := auto | POSITIVE-INT\n");
-	exit(-1);
-}
-
-/* This socket is used to get nsid */
-static struct rtnl_handle rtnsh = { .fd = -1 };
-
-static int have_rtnl_getnsid = -1;
-
-static int ipnetns_accept_msg(struct rtnl_ctrl_data *ctrl,
-			      struct nlmsghdr *n, void *arg)
-{
-	struct nlmsgerr *err = (struct nlmsgerr *)NLMSG_DATA(n);
-
-	if (n->nlmsg_type == NLMSG_ERROR &&
-	    (err->error == -EOPNOTSUPP || err->error == -EINVAL))
-		have_rtnl_getnsid = 0;
-	else
-		have_rtnl_getnsid = 1;
-	return -1;
-}
-
[…]

### 2019-11-15 iproute2/ip/ipntable.c

@@ -1,696 +1,0 @@
-/*
- * Copyright (C)2006 USAGI/WIDE Project
- *
- * This program is free software; you can redistribute it and/or modify
- * it under the terms of the GNU General Public License as published by
- * the Free Software Foundation; either version 2 of the License, or
- * (at your option) any later version.
- *
- * This program is distributed in the hope that it will be useful,
- * but WITHOUT ANY WARRANTY; without even the implied warranty of
- * MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
- * GNU General Public License for more details.
- *
- * You should have received a copy of the GNU General Public License
- * along with this program; if not, see <http://www.gnu.org/licenses>.
- */
-/*
- * based on ipneigh.c
- */
-/*
- * Authors:
- *	Masahide NAKAMURA @USAGI
- */
-
-#include <stdio.h>
-#include <stdlib.h>
-#include <string.h>
-#include <sys/time.h>
-#include <sys/socket.h>
-#include <time.h>
-
-#include "utils.h"
-#include "ip_common.h"
-#include "json_print.h"
-
-static struct
-{
-	int family;
-	int index;
-#define NONE_DEV	(-1)
-	const char *name;
-} filter;
-
-static void usage(void) __attribute__((noreturn));
-
-static void usage(void)
-{
-	fprintf(stderr,
-		"Usage: ip ntable change name NAME [ dev DEV ]\n"
-		"          [ thresh1 VAL ] [ thresh2 VAL ] [ thresh3 VAL ] [ gc_int MSEC ]\n"
-		"          [ PARMS ]\n"
-		"Usage: ip ntable show [ dev DEV ] [ name NAME ]\n"
-
-		"PARMS := [ base_reachable MSEC ] [ retrans MSEC ] [ gc_stale MSEC ]\n"
-		"         [ delay_probe MSEC ] [ queue LEN ]\n"
-		"         [ app_probes VAL ] [ ucast_probes VAL ] [ mcast_probes VAL ]\n"
[…]

### 2019-11-15 iproute2/ip/iproute.c

@@ -1,2310 +1,0 @@
-/*
- * iproute.c		"ip route".
- *
- *		This program is free software; you can redistribute it and/or
- *		modify it under the terms of the GNU General Public License
- *		as published by the Free Software Foundation; either version
- *		2 of the License, or (at your option) any later version.
- *
- * Authors:	Alexey Kuznetsov, <kuznet@ms2.inr.ac.ru>
- *
- */
-
-#include <stdio.h>
-#include <stdlib.h>
-#include <unistd.h>
-#include <fcntl.h>
-#include <string.h>
-#include <time.h>
-#include <sys/time.h>
-#include <sys/socket.h>
-#include <netinet/in.h>
-#include <netinet/ip.h>
-#include <arpa/inet.h>
-#include <linux/in_route.h>
-#include <linux/icmpv6.h>
-#include <errno.h>
-
-#include "rt_names.h"
-#include "utils.h"
-#include "ip_common.h"
-
-#ifndef RTAX_RTTVAR
-#define RTAX_RTTVAR RTAX_HOPS
-#endif
-
-enum list_action {
-	IPROUTE_LIST,
-	IPROUTE_FLUSH,
-	IPROUTE_SAVE,
-};
-static const char *mx_names[RTAX_MAX+1] = {
-	[RTAX_MTU]			= "mtu",
-	[RTAX_WINDOW]			= "window",
-	[RTAX_RTT]			= "rtt",
-	[RTAX_RTTVAR]			= "rttvar",
-	[RTAX_SSTHRESH]			= "ssthresh",
-	[RTAX_CWND]			= "cwnd",
-	[RTAX_ADVMSS]			= "advmss",
-	[RTAX_REORDERING]		= "reordering",
-	[RTAX_HOPLIMIT]			= "hoplimit",
-	[RTAX_INITCWND]			= "initcwnd",
-	[RTAX_FEATURES]			= "features",
-	[RTAX_RTO_MIN]			= "rto_min",
-	[RTAX_INITRWND]			= "initrwnd",
-	[RTAX_QUICKACK]			= "quickack",
-	[RTAX_CC_ALGO]			= "congctl",
-	[RTAX_FASTOPEN_NO_COOKIE]	= "fastopen_no_cookie"
-};
-static void usage(void) __attribute__((noreturn));
-
-static void usage(void)
-{
-	fprintf(stderr,
-		"Usage: ip route { list | flush } SELECTOR\n"
[…]

### 2019-11-15 iproute2/ip/iproute_lwtunnel.c

@@ -1,1173 +1,0 @@
-/*
- * iproute_lwtunnel.c
- *
- *		This program is free software; you can redistribute it and/or
- *		modify it under the terms of the GNU General Public License
- *		as published by the Free Software Foundation; either version
- *		2 of the License, or (at your option) any later version.
- *
- * Authors:	Roopa Prabhu, <roopa@cumulusnetworks.com>
- *		Thomas Graf <tgraf@suug.ch>
- *
- */
-
-#include <stdio.h>
-#include <stdlib.h>
-#include <unistd.h>
-#include <fcntl.h>
-#include <string.h>
-#include <linux/ila.h>
-#include <linux/lwtunnel.h>
-#include <linux/mpls_iptunnel.h>
-#include <errno.h>
-
-#include "rt_names.h"
-#include "bpf_util.h"
-#include "utils.h"
-#include "ip_common.h"
-#include "ila_common.h"
-
-#include <linux/seg6.h>
-#include <linux/seg6_iptunnel.h>
-#include <linux/seg6_hmac.h>
-#include <linux/seg6_local.h>
-#include <linux/if_tunnel.h>
-
-static const char *format_encap_type(int type)
-{
-	switch (type) {
-	case LWTUNNEL_ENCAP_MPLS:
-		return "mpls";
-	case LWTUNNEL_ENCAP_IP:
-		return "ip";
-	case LWTUNNEL_ENCAP_IP6:
-		return "ip6";
-	case LWTUNNEL_ENCAP_ILA:
-		return "ila";
-	case LWTUNNEL_ENCAP_BPF:
-		return "bpf";
-	case LWTUNNEL_ENCAP_SEG6:
-		return "seg6";
-	case LWTUNNEL_ENCAP_SEG6_LOCAL:
-		return "seg6local";
-	default:
-		return "unknown";
-	}
-}
-
-static void encap_type_usage(void)
-{
-	int i;
-
-	fprintf(stderr, "Usage: ip route ... encap TYPE [ OPTIONS ] [...]\n");
-
[…]

### 2019-11-15 iproute2/ip/iptoken.c

@@ -1,208 +1,0 @@
-/*
- * iptoken.c    "ip token"
- *
- *              This program is free software; you can redistribute it and/or
- *              modify it under the terms of the GNU General Public License
- *              as published by the Free Software Foundation; either version
- *              2 of the License, or (at your option) any later version.
- *
- * Authors:     Daniel Borkmann, <borkmann@redhat.com>
- */
-
-#include <stdio.h>
-#include <stdlib.h>
-#include <stdbool.h>
-#include <unistd.h>
-#include <fcntl.h>
-#include <string.h>
-#include <sys/socket.h>
-#include <netinet/in.h>
-#include <netinet/ip.h>
-#include <arpa/inet.h>
-#include <linux/types.h>
-#include <linux/if.h>
-
-#include "rt_names.h"
-#include "utils.h"
-#include "ip_common.h"
-#include "json_print.h"
-
-extern struct rtnl_handle rth;
-
-struct rtnl_dump_args {
-	FILE *fp;
-	int ifindex;
-};
-
-static void usage(void) __attribute__((noreturn));
-
-static void usage(void)
-{
-	fprintf(stderr, "Usage: ip token [ list | set | del | get ] [ TOKEN ] [ dev DEV ]\n");
-	exit(-1);
-}
-
-static int print_token(struct nlmsghdr *n, void *arg)
-{
-	struct rtnl_dump_args *args = arg;
-	FILE *fp = args->fp;
-	int ifindex = args->ifindex;
-	struct ifinfomsg *ifi = NLMSG_DATA(n);
-	int len = n->nlmsg_len;
-	struct rtattr *tb[IFLA_MAX + 1];
-	struct rtattr *ltb[IFLA_INET6_MAX + 1];
-
-	if (n->nlmsg_type != RTM_NEWLINK)
-		return -1;
-
-	len -= NLMSG_LENGTH(sizeof(*ifi));
-	if (len < 0)
-		return -1;
-
-	if (ifi->ifi_family != AF_INET6)
-		return -1;
-	if (ifi->ifi_index == 0)
-		return -1;
-	if (ifindex > 0 && ifi->ifi_index != ifindex)
-		return 0;
-	if (ifi->ifi_flags & (IFF_LOOPBACK | IFF_NOARP))
-		return 0;
-
[…]

### 2019-11-15 iproute2/ip/iptuntap.c

@@ -1,598 +1,0 @@
-/*
- * iptunnel.c	       "ip tuntap"
- *
- *		This program is free software; you can redistribute it and/or
- *		modify it under the terms of the GNU General Public License
- *		as published by the Free Software Foundation; either version
- *		2 of the License, or (at your option) any later version.
- *
- * Authors:	David Woodhouse <David.Woodhouse@intel.com>
- *
- */
-
-#include <stdio.h>
-#include <stdlib.h>
-#include <string.h>
-#include <unistd.h>
-#include <sys/types.h>
-#include <sys/socket.h>
-#include <arpa/inet.h>
-#include <sys/ioctl.h>
-#include <linux/if.h>
-#include <linux/if_tun.h>
-#include <linux/if_arp.h>
-#include <pwd.h>
-#include <grp.h>
-#include <fcntl.h>
-#include <dirent.h>
-#include <errno.h>
-#include <glob.h>
-
-#include "rt_names.h"
-#include "utils.h"
-#include "ip_common.h"
-
-static const char drv_name[] = "tun";
-
-#define TUNDEV "/dev/net/tun"
-
-static void usage(void) __attribute__((noreturn));
-
-static void usage(void)
-{
-	fprintf(stderr, "Usage: ip tuntap { add | del | show | list | lst | help } [ dev PHYS_DEV ]\n");
-	fprintf(stderr, "          [ mode { tun | tap } ] [ user USER ] [ group GROUP ]\n");
-	fprintf(stderr, "          [ one_queue ] [ pi ] [ vnet_hdr ] [ multi_queue ] [ name NAME ]\n");
-	fprintf(stderr, "\n");
-	fprintf(stderr, "Where: USER  := { STRING | NUMBER }\n");
-	fprintf(stderr, "       GROUP := { STRING | NUMBER }\n");
-	exit(-1);
-}
-
-static int tap_add_ioctl(struct ifreq *ifr, uid_t uid, gid_t gid)
-{
-	int fd;
-	int ret = -1;
-
-#ifdef IFF_TUN_EXCL
-	ifr->ifr_flags |= IFF_TUN_EXCL;
-#endif
-
-	fd = open(TUNDEV, O_RDWR);
[…]

### 2019-11-15 iproute2/ip/ipvrf.c

@@ -1,638 +1,0 @@
-/*
- * ipvrf.c	"ip vrf"
- *
- *		This program is free software; you can redistribute it and/or
- *		modify it under the terms of the GNU General Public License
- *		as published by the Free Software Foundation; either version
- *		2 of the License, or (at your option) any later version.
- *
- * Authors:	David Ahern <dsa@cumulusnetworks.com>
- *
- */
-
-#include <sys/types.h>
-#include <sys/stat.h>
-#include <sys/socket.h>
-#include <sys/mount.h>
-#include <linux/bpf.h>
-#include <linux/if.h>
-#include <fcntl.h>
-#include <stdio.h>
-#include <stdlib.h>
-#include <unistd.h>
-#include <string.h>
-#include <dirent.h>
-#include <errno.h>
-#include <limits.h>
-
-#include "rt_names.h"
-#include "utils.h"
-#include "ip_common.h"
-#include "bpf_util.h"
-
-#define CGRP_PROC_FILE  "/cgroup.procs"
-
-static struct link_filter vrf_filter;
-
-static void usage(void)
-{
-	fprintf(stderr, "Usage: ip vrf show [NAME] ...\n");
-	fprintf(stderr, "       ip vrf exec [NAME] cmd ...\n");
-	fprintf(stderr, "       ip vrf identify [PID]\n");
-	fprintf(stderr, "       ip vrf pids [NAME]\n");
-
-	exit(-1);
-}
-
-/*
- * parse process based cgroup file looking for PATH/vrf/NAME where
- * NAME is the name of the vrf the process is associated with
- */
-static int vrf_identify(pid_t pid, char *name, size_t len)
-{
-	char path[PATH_MAX];
-	char buf[4096];
-	char *vrf, *end;
-	FILE *fp;
-
-	snprintf(path, sizeof(path), "/proc/%d/cgroup", pid);
-	fp = fopen(path, "r");
-	if (!fp)
-		return -1;
-
-	memset(name, 0, len);
-
-	while (fgets(buf, sizeof(buf), fp)) {
-		/* want the controller-less cgroup */
-		if (strstr(buf, "::/") == NULL)
-			continue;
-
[…]

### 2019-11-15 iproute2/ip/ipxfrm.c

@@ -1,1552 +1,0 @@
-/* $USAGI: $ */
-
-/*
- * Copyright (C)2004 USAGI/WIDE Project
- *
- * This program is free software; you can redistribute it and/or modify
- * it under the terms of the GNU General Public License as published by
- * the Free Software Foundation; either version 2 of the License, or
- * (at your option) any later version.
- *
- * This program is distributed in the hope that it will be useful,
- * but WITHOUT ANY WARRANTY; without even the implied warranty of
- * MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
- * GNU General Public License for more details.
- *
- * You should have received a copy of the GNU General Public License
- * along with this program; if not, see <http://www.gnu.org/licenses>.
- */
-/*
- * based on ip.c, iproute.c
- */
-/*
- * Authors:
- *	Masahide NAKAMURA @USAGI
- */
-
-#include <alloca.h>
-#include <stdio.h>
-#include <stdlib.h>
-#include <string.h>
-#include <sys/types.h>
-#include <sys/socket.h>
-#include <time.h>
-#include <netdb.h>
-#include <linux/netlink.h>
-#include <linux/rtnetlink.h>
-
-#include "utils.h"
-#include "xfrm.h"
-#include "ip_common.h"
-
-#define STRBUF_SIZE	(128)
-
-struct xfrm_filter filter;
-
-static void usage(void) __attribute__((noreturn));
-
-static void usage(void)
-{
-	fprintf(stderr,
-		"Usage: ip xfrm XFRM-OBJECT { COMMAND | help }\n"
-		"where  XFRM-OBJECT := state | policy | monitor\n");
-	exit(-1);
-}
-
-/* This is based on utils.c(inet_addr_match) */
-int xfrm_addr_match(xfrm_address_t *x1, xfrm_address_t *x2, int bits)
-{
-	__u32 *a1 = (__u32 *)x1;
-	__u32 *a2 = (__u32 *)x2;
-	int words = bits >> 0x05;
-
-	bits &= 0x1f;
-
-	if (words)
[…]

### 2019-11-15 iproute2/ip/link_gre.c

@@ -1,576 +1,0 @@
-/*
- * link_gre.c	gre driver module
- *
- *		This program is free software; you can redistribute it and/or
- *		modify it under the terms of the GNU General Public License
- *		as published by the Free Software Foundation; either version
- *		2 of the License, or (at your option) any later version.
- *
- * Authors:	Herbert Xu <herbert@gondor.apana.org.au>
- *
- */
-
-#include <string.h>
-#include <net/if.h>
-#include <sys/types.h>
-#include <sys/socket.h>
-#include <arpa/inet.h>
-
-#include <linux/ip.h>
-#include <linux/if_tunnel.h>
-#include "rt_names.h"
-#include "utils.h"
-#include "ip_common.h"
-#include "tunnel.h"
-
-static void gre_print_help(struct link_util *lu, int argc, char **argv, FILE *f)
-{
-	fprintf(f,
-		"Usage: ... %-9s [ remote ADDR ]\n",
-		lu->id
-	);
-	fprintf(f,
-		"                     [ local ADDR ]\n"
-		"                     [ [no][i|o]seq ]\n"
-		"                     [ [i|o]key KEY | no[i|o]key ]\n"
-		"                     [ [no][i|o]csum ]\n"
-		"                     [ ttl TTL ]\n"
-		"                     [ tos TOS ]\n"
-		"                     [ [no]pmtudisc ]\n"
-		"                     [ [no]ignore-df ]\n"
-		"                     [ dev PHYS_DEV ]\n"
-		"                     [ fwmark MARK ]\n"
-		"                     [ external ]\n"
-		"                     [ noencap ]\n"
-		"                     [ encap { fou | gue | none } ]\n"
-		"                     [ encap-sport PORT ]\n"
-		"                     [ encap-dport PORT ]\n"
-		"                     [ [no]encap-csum ]\n"
[…]

### 2019-11-15 iproute2/ip/link_gre6.c

@@ -1,644 +1,0 @@
-/*
- * link_gre6.c	gre driver module
- *
- *		This program is free software; you can redistribute it and/or
- *		modify it under the terms of the GNU General Public License
- *		as published by the Free Software Foundation; either version
- *		2 of the License, or (at your option) any later version.
- *
- * Authors:	Dmitry Kozlov <xeb@mail.ru>
- *
- */
-
-#include <string.h>
-#include <net/if.h>
-#include <sys/types.h>
-#include <sys/socket.h>
-#include <arpa/inet.h>
-
-#include <linux/ip.h>
-#include <linux/if_tunnel.h>
-#include <linux/ip6_tunnel.h>
-
-#include "rt_names.h"
-#include "utils.h"
-#include "ip_common.h"
-#include "tunnel.h"
-
-#define IP6_FLOWINFO_TCLASS	htonl(0x0FF00000)
-#define IP6_FLOWINFO_FLOWLABEL	htonl(0x000FFFFF)
-
-#define DEFAULT_TNL_HOP_LIMIT	(64)
-
-static void gre_print_help(struct link_util *lu, int argc, char **argv, FILE *f)
-{
-	fprintf(f,
-		"Usage: ... %-9s [ remote ADDR ]\n",
-		lu->id
-	);
-	fprintf(f,
-		"                     [ local ADDR ]\n"
-		"                     [ [no][i|o]seq ]\n"
-		"                     [ [i|o]key KEY | no[i|o]key ]\n"
-		"                     [ [no][i|o]csum ]\n"
-		"                     [ hoplimit TTL ]\n"
-		"                     [ encaplimit ELIM ]\n"
-		"                     [ tclass TCLASS ]\n"
-		"                     [ flowlabel FLOWLABEL ]\n"
-		"                     [ dscp inherit ]\n"
-		"                     [ dev PHYS_DEV ]\n"
-		"                     [ fwmark MARK ]\n"
-		"                     [ [no]allow-localremote ]\n"
[…]

### 2019-11-15 iproute2/ip/link_ip6tnl.c

@@ -1,477 +1,0 @@
-/*
- * link_ip6tnl.c	ip6tnl driver module
- *
- *		This program is free software; you can redistribute it and/or
- *		modify it under the terms of the GNU General Public License
- *		as published by the Free Software Foundation; either version
- *		2 of the License, or (at your option) any later version.
- *
- * Authors:	Nicolas Dichtel <nicolas.dichtel@6wind.com>
- *
- */
-
-#include <string.h>
-#include <net/if.h>
-#include <sys/types.h>
-#include <sys/socket.h>
-#include <arpa/inet.h>
-
-#include <linux/ip.h>
-#include <linux/if_tunnel.h>
-#include <linux/ip6_tunnel.h>
-#include "rt_names.h"
-#include "utils.h"
-#include "ip_common.h"
-#include "tunnel.h"
-
-#define IP6_FLOWINFO_TCLASS	htonl(0x0FF00000)
-#define IP6_FLOWINFO_FLOWLABEL	htonl(0x000FFFFF)
-
-#define DEFAULT_TNL_HOP_LIMIT	(64)
-
-static void ip6tunnel_print_help(struct link_util *lu, int argc, char **argv,
-				 FILE *f)
-{
-	const char *mode;
-
-	fprintf(f,
-		"Usage: ... %-6s [ remote ADDR ]\n",
-		lu->id
-	);
-	fprintf(f,
-		"                  [ local ADDR ]\n"
-		"                  [ encaplimit ELIM ]\n"
-		"                  [ hoplimit HLIM ]\n"
-		"                  [ tclass TCLASS ]\n"
-		"                  [ flowlabel FLOWLABEL ]\n"
-		"                  [ dscp inherit ]\n"
-		"                  [ [no]allow-localremote ]\n"
-		"                  [ dev PHYS_DEV ]\n"
-		"                  [ fwmark MARK ]\n"
-		"                  [ external ]\n"
-		"                  [ noencap ]\n"
[…]

### 2019-11-15 iproute2/ip/link_iptnl.c

@@ -1,504 +1,0 @@
-/*
- * link_iptnl.c	ipip and sit driver module
- *
- *		This program is free software; you can redistribute it and/or
- *		modify it under the terms of the GNU General Public License
- *		as published by the Free Software Foundation; either version
- *		2 of the License, or (at your option) any later version.
- *
- * Authors:	Nicolas Dichtel <nicolas.dichtel@6wind.com>
- *
- */
-
-#include <string.h>
-#include <net/if.h>
-#include <sys/types.h>
-#include <sys/socket.h>
-#include <arpa/inet.h>
-
-#include <linux/in.h>
-#include <linux/ip.h>
-#include <linux/if_tunnel.h>
-#include "rt_names.h"
-#include "utils.h"
-#include "ip_common.h"
-#include "tunnel.h"
-
-static void iptunnel_print_help(struct link_util *lu, int argc, char **argv,
-				FILE *f)
-{
-	const char *mode;
-
-	fprintf(f,
-		"Usage: ... %-6s [ remote ADDR ]\n",
-		lu->id
-	);
-	fprintf(f,
-		"                  [ local ADDR ]\n"
-		"                  [ ttl TTL ]\n"
-		"                  [ tos TOS ]\n"
-		"                  [ [no]pmtudisc ]\n"
-		"                  [ 6rd-prefix ADDR ]\n"
-		"                  [ 6rd-relay_prefix ADDR ]\n"
-		"                  [ 6rd-reset ]\n"
-		"                  [ dev PHYS_DEV ]\n"
-		"                  [ fwmark MARK ]\n"
-		"                  [ external ]\n"
-		"                  [ noencap ]\n"
-		"                  [ encap { fou | gue | none } ]\n"
-		"                  [ encap-sport PORT ]\n"
-		"                  [ encap-dport PORT ]\n"
-		"                  [ [no]encap-csum ]\n"
[…]

### 2019-11-15 iproute2/ip/link_vti.c

@@ -1,222 +1,0 @@
-/*
- * link_vti.c	VTI driver module
- *
- *		This program is free software; you can redistribute it and/or
- *		modify it under the terms of the GNU General Public License
- *		as published by the Free Software Foundation; either version
- *		2 of the License, or (at your option) any later version.
- *
- * Authors:	Herbert Xu <herbert@gondor.apana.org.au>
- *          Saurabh Mohan <saurabh.mohan@vyatta.com> Modified link_gre.c for VTI
- */
-
-#include <string.h>
-#include <net/if.h>
-#include <sys/types.h>
-#include <sys/socket.h>
-#include <arpa/inet.h>
-
-#include <linux/ip.h>
-#include <linux/if_tunnel.h>
-#include "rt_names.h"
-#include "utils.h"
-#include "ip_common.h"
-#include "tunnel.h"
-
-static void vti_print_help(struct link_util *lu, int argc, char **argv, FILE *f)
-{
-	fprintf(f,
-		"Usage: ... %-4s [ remote ADDR ]\n",
-		lu->id
-	);
-	fprintf(f,
-		"                [ local ADDR ]\n"
-		"                [ [i|o]key KEY ]\n"
-		"                [ dev PHYS_DEV ]\n"
-		"                [ fwmark MARK ]\n"
-		"\n"
-	);
-	fprintf(f,
-		"Where: ADDR := { IP%s_ADDRESS }\n"
-		"       KEY  := { DOTTED_QUAD | NUMBER }\n"
-		"       MARK := { 0x0..0xffffffff }\n",
-		""
-	);
-}
-
-static int vti_parse_opt(struct link_util *lu, int argc, char **argv,
-			 struct nlmsghdr *n)
-{
-	struct ifinfomsg *ifi = NLMSG_DATA(n);
-	struct {
-		struct nlmsghdr n;
-		struct ifinfomsg i;
-	} req = {
-		.n.nlmsg_len = NLMSG_LENGTH(sizeof(*ifi)),
-		.n.nlmsg_flags = NLM_F_REQUEST,
-		.n.nlmsg_type = RTM_GETLINK,
[…]

### 2019-11-15 iproute2/ip/link_vti6.c

@@ -1,224 +1,0 @@
-/*
- * link_vti6.c	VTI driver module
- *
- *		This program is free software; you can redistribute it and/or
- *		modify it under the terms of the GNU General Public License
- *		as published by the Free Software Foundation; either version
- *		2 of the License, or (at your option) any later version.
- *
- * Authors:	Herbert Xu <herbert@gondor.apana.org.au>
- *		Saurabh Mohan <saurabh.mohan@vyatta.com> Modified link_gre.c for VTI
- *		Steffen Klassert <steffen.klassert@secunet.com> Modified link_vti.c for IPv6
- */
-
-#include <string.h>
-#include <net/if.h>
-#include <sys/types.h>
-#include <sys/socket.h>
-#include <arpa/inet.h>
-
-#include <linux/ip.h>
-#include <linux/if_tunnel.h>
-#include "rt_names.h"
-#include "utils.h"
-#include "ip_common.h"
-#include "tunnel.h"
-
-static void vti6_print_help(struct link_util *lu, int argc, char **argv,
-			    FILE *f)
-{
-	fprintf(f,
-		"Usage: ... %-4s [ remote ADDR ]\n",
-		lu->id
-	);
-	fprintf(f,
-		"                [ local ADDR ]\n"
-		"                [ [i|o]key KEY ]\n"
-		"                [ dev PHYS_DEV ]\n"
-		"                [ fwmark MARK ]\n"
-		"\n"
-	);
-	fprintf(f,
-		"Where: ADDR := { IP%s_ADDRESS }\n"
-		"       KEY  := { DOTTED_QUAD | NUMBER }\n"
-		"       MARK := { 0x0..0xffffffff }\n",
-		"V6"
-	);
-}
-
-static int vti6_parse_opt(struct link_util *lu, int argc, char **argv,
-			  struct nlmsghdr *n)
-{
-	struct ifinfomsg *ifi = NLMSG_DATA(n);
-	struct {
-		struct nlmsghdr n;
-		struct ifinfomsg i;
-	} req = {
-		.n.nlmsg_len = NLMSG_LENGTH(sizeof(*ifi)),
[…]

### 2019-11-15 iproute2/ip/link_xfrm.c

@@ -1,77 +1,0 @@
-// SPDX-License-Identifier: GPL-2.0
-/*
- * link_xfrm.c	Virtual XFRM Interface driver module
- *
- * Authors:	Matt Ellison <matt@arroyo.io>
- */
-
-#include <string.h>
-#include <linux/if_link.h>
-
-#include "rt_names.h"
-#include "utils.h"
-#include "ip_common.h"
-#include "tunnel.h"
-
-static void xfrm_print_help(struct link_util *lu, int argc, char **argv,
-			    FILE *f)
-{
-	fprintf(f, "Usage: ... %-4s dev PHYS_DEV [ if_id IF-ID ]\n", lu->id);
-	fprintf(f, "\nWhere: IF-ID := { 0x0..0xffffffff }\n");
-}
-
-static int xfrm_parse_opt(struct link_util *lu, int argc, char **argv,
-			  struct nlmsghdr *n)
-{
-	unsigned int link = 0;
-	__u32 if_id = 0;
-
-	while (argc > 0) {
-		if (!matches(*argv, "dev")) {
-			NEXT_ARG();
-			link = ll_name_to_index(*argv);
-			if (!link)
-				exit(nodev(*argv));
-		} else if (!matches(*argv, "if_id")) {
-			NEXT_ARG();
-			if (!get_u32(&if_id, *argv, 0))
-				addattr32(n, 1024, IFLA_XFRM_IF_ID, if_id);
-		} else {
-			xfrm_print_help(lu, argc, argv, stderr);
-			return -1;
-		}
-		argc--; argv++;
-	}
-
-	if (link) {
-		addattr32(n, 1024, IFLA_XFRM_LINK, link);
-	} else {
-		fprintf(stderr, "must specify physical device\n");
-		return -1;
-	}
-
-	return 0;
-}
-
-static void xfrm_print_opt(struct link_util *lu, FILE *f, struct rtattr *tb[])
-{
-
-	if (!tb)
-		return;
-
-	if (tb[IFLA_XFRM_IF_ID]) {
-		__u32 id = rta_getattr_u32(tb[IFLA_XFRM_IF_ID]);
-
-		print_0xhex(PRINT_ANY, "if_id", "if_id %#llx ", id);
-
-	}
-
-}
-
-struct link_util xfrm_link_util = {
-	.id = "xfrm",
-	.maxattr = IFLA_XFRM_MAX,
-	.parse_opt = xfrm_parse_opt,
[…]

### 2019-11-15 iproute2/ip/rtmon.c

@@ -1,183 +1,0 @@
-/*
- * rtmon.c		RTnetlink listener.
- *
- *		This program is free software; you can redistribute it and/or
- *		modify it under the terms of the GNU General Public License
- *		as published by the Free Software Foundation; either version
- *		2 of the License, or (at your option) any later version.
- *
- * Authors:	Alexey Kuznetsov, <kuznet@ms2.inr.ac.ru>
- *
- */
-
-#include <stdio.h>
-#include <stdlib.h>
-#include <unistd.h>
-#include <fcntl.h>
-#include <sys/socket.h>
-#include <sys/time.h>
-#include <netinet/in.h>
-#include <string.h>
-
-#include "SNAPSHOT.h"
-
-#include "utils.h"
-#include "libnetlink.h"
-
-static int init_phase = 1;
-
-static void write_stamp(FILE *fp)
-{
-	char buf[128];
-	struct nlmsghdr *n1 = (void *)buf;
-	struct timeval tv;
-
-	n1->nlmsg_type = NLMSG_TSTAMP;
-	n1->nlmsg_flags = 0;
-	n1->nlmsg_seq = 0;
-	n1->nlmsg_pid = 0;
-	n1->nlmsg_len = NLMSG_LENGTH(4*2);
-	gettimeofday(&tv, NULL);
-	((__u32 *)NLMSG_DATA(n1))[0] = tv.tv_sec;
-	((__u32 *)NLMSG_DATA(n1))[1] = tv.tv_usec;
-	fwrite((void *)n1, 1, NLMSG_ALIGN(n1->nlmsg_len), fp);
-}
-
-static int dump_msg(struct rtnl_ctrl_data *ctrl,
-		    struct nlmsghdr *n, void *arg)
-{
-	FILE *fp = (FILE *)arg;
-
-	if (!init_phase)
-		write_stamp(fp);
-	fwrite((void *)n, 1, NLMSG_ALIGN(n->nlmsg_len), fp);
-	fflush(fp);
-	return 0;
-}
-
-static int dump_msg2(struct nlmsghdr *n, void *arg)
-{
-	return dump_msg(NULL, n, arg);
-}
-
-static void usage(void)
-{
-	fprintf(stderr, "Usage: rtmon [ OPTIONS ] file FILE [ all | LISTofOBJECTS ]\n");
-	fprintf(stderr, "OPTIONS := { -f[amily] { inet | inet6 | link | help } |\n"
[…]

### 2019-11-15 iproute2/ip/tcp_metrics.c

@@ -1,541 +1,0 @@
-/*
- * tcp_metrics.c	"ip tcp_metrics/tcpmetrics"
- *
- *		This program is free software; you can redistribute it and/or
- *		modify it under the terms of the GNU General Public License
- *		version 2 as published by the Free Software Foundation;
- *
- * Authors:	Julian Anastasov <ja@ssi.bg>, August 2012
- */
-
-#include <stdio.h>
-#include <stdlib.h>
-#include <string.h>
-#include <unistd.h>
-#include <errno.h>
-#include <sys/types.h>
-#include <sys/socket.h>
-#include <arpa/inet.h>
-#include <sys/ioctl.h>
-#include <linux/if.h>
-
-#include <linux/genetlink.h>
-#include <linux/tcp_metrics.h>
-
-#include "utils.h"
-#include "ip_common.h"
-#include "libgenl.h"
-
-static void usage(void)
-{
-	fprintf(stderr, "Usage: ip tcp_metrics/tcpmetrics { COMMAND | help }\n");
-	fprintf(stderr, "       ip tcp_metrics { show | flush } SELECTOR\n");
-	fprintf(stderr, "       ip tcp_metrics delete [ address ] ADDRESS\n");
-	fprintf(stderr, "SELECTOR := [ [ address ] PREFIX ]\n");
-	exit(-1);
-}
-
-/* netlink socket */
-static struct rtnl_handle grth = { .fd = -1 };
-static int genl_family = -1;
-static const double usec_per_sec = 1000000.;
-
-#define TCPM_REQUEST(_req, _bufsiz, _cmd, _flags) \
-	GENL_REQUEST(_req, _bufsiz, genl_family, 0, \
-		     TCP_METRICS_GENL_VERSION, _cmd, _flags)
-
-#define CMD_LIST	0x0001	/* list, lst, show		*/
-#define CMD_DEL		0x0002	/* delete, remove		*/
-#define CMD_FLUSH	0x0004	/* flush			*/
-
-static const struct {
-	const char *name;
-	int	code;
-} cmds[] = {
-	{	"list",		CMD_LIST	},
-	{	"lst",		CMD_LIST	},
-	{	"show",		CMD_LIST	},
[…]

### 2019-11-15 iproute2/ip/tunnel.c

@@ -1,405 +1,0 @@
-/*
- * Copyright (C)2006 USAGI/WIDE Project
- *
- * This program is free software; you can redistribute it and/or modify
- * it under the terms of the GNU General Public License as published by
- * the Free Software Foundation; either version 2 of the License, or
- * (at your option) any later version.
- *
- * This program is distributed in the hope that it will be useful,
- * but WITHOUT ANY WARRANTY; without even the implied warranty of
- * MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
- * GNU General Public License for more details.
- *
- * You should have received a copy of the GNU General Public License
- * along with this program; if not, see <http://www.gnu.org/licenses>.
- */
-/*
- * split from ip_tunnel.c
- */
-/*
- * Author:
- *	Masahide NAKAMURA @USAGI
- */
-
-#include <stdio.h>
-#include <string.h>
-#include <unistd.h>
-#include <errno.h>
-#include <sys/types.h>
-#include <sys/socket.h>
-#include <sys/ioctl.h>
-#include <netinet/in.h>
-#include <linux/if.h>
-#include <linux/ip.h>
-#include <linux/if_tunnel.h>
-#include <linux/if_arp.h>
-
-#include "utils.h"
-#include "tunnel.h"
-#include "json_print.h"
-
-const char *tnl_strproto(__u8 proto)
-{
-	switch (proto) {
-	case IPPROTO_IPIP:
-		return "ip";
-	case IPPROTO_GRE:
-		return "gre";
-	case IPPROTO_IPV6:
-		return "ipv6";
-	case IPPROTO_ESP:
-		return "esp";
-	case IPPROTO_MPLS:
-		return "mpls";
-	case 0:
-		return "any";
-	default:
-		return "unknown";
-	}
-}
-
-int tnl_get_ioctl(const char *basedev, void *p)
-{
-	struct ifreq ifr;
-	int fd;
-	int err;
-
-	strlcpy(ifr.ifr_name, basedev, IFNAMSIZ);
-	ifr.ifr_ifru.ifru_data = (void *)p;
-
[…]

### 2019-11-15 iproute2/ip/xfrm_monitor.c

@@ -1,425 +1,0 @@
-/* $USAGI: $ */
-
-/*
- * Copyright (C)2005 USAGI/WIDE Project
- *
- * This program is free software; you can redistribute it and/or modify
- * it under the terms of the GNU General Public License as published by
- * the Free Software Foundation; either version 2 of the License, or
- * (at your option) any later version.
- *
- * This program is distributed in the hope that it will be useful,
- * but WITHOUT ANY WARRANTY; without even the implied warranty of
- * MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
- * GNU General Public License for more details.
- *
- * You should have received a copy of the GNU General Public License
- * along with this program; if not, see <http://www.gnu.org/licenses>.
- */
-/*
- * based on ipmonitor.c
- */
-/*
- * Authors:
- *	Masahide NAKAMURA @USAGI
- */
-
-#include <stdio.h>
-#include <stdlib.h>
-#include <string.h>
-#include <netinet/in.h>
-
-#include "utils.h"
-#include "xfrm.h"
-#include "ip_common.h"
-
-static void usage(void) __attribute__((noreturn));
-static int listen_all_nsid;
-static bool nokeys;
-
-static void usage(void)
-{
-	fprintf(stderr, "Usage: ip xfrm monitor [ nokeys ] [ all-nsid ] [ all | OBJECTS | help ]\n");
-	fprintf(stderr, "OBJECTS := { acquire | expire | SA | aevent | policy | report }\n");
-	exit(-1);
-}
-
-static int xfrm_acquire_print(struct nlmsghdr *n, void *arg)
-{
-	FILE *fp = (FILE *)arg;
-	struct xfrm_user_acquire *xacq = NLMSG_DATA(n);
-	int len = n->nlmsg_len;
-	struct rtattr *tb[XFRMA_MAX+1];
-	__u16 family;
-
-	len -= NLMSG_LENGTH(sizeof(*xacq));
[…]

### 2019-11-15 iproute2/ip/xfrm_policy.c

@@ -1,1195 +1,0 @@
-/* $USAGI: $ */
-
-/*
- * Copyright (C)2004 USAGI/WIDE Project
- *
- * This program is free software; you can redistribute it and/or modify
- * it under the terms of the GNU General Public License as published by
- * the Free Software Foundation; either version 2 of the License, or
- * (at your option) any later version.
- *
- * This program is distributed in the hope that it will be useful,
- * but WITHOUT ANY WARRANTY; without even the implied warranty of
- * MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
- * GNU General Public License for more details.
- *
- * You should have received a copy of the GNU General Public License
- * along with this program; if not, see <http://www.gnu.org/licenses>.
- */
-/*
- * based on iproute.c
- */
-/*
- * Authors:
- *	Masahide NAKAMURA @USAGI
- */
-
-#include <stdio.h>
-#include <stdlib.h>
-#include <string.h>
-#include <netdb.h>
-#include <linux/netlink.h>
-#include "utils.h"
-#include "xfrm.h"
-#include "ip_common.h"
-
-/* #define NLMSG_DELETEALL_BUF_SIZE (4096-512) */
-#define NLMSG_DELETEALL_BUF_SIZE 8192
-
-/*
- * Receiving buffer defines:
- * nlmsg
- *   data = struct xfrm_userpolicy_info
- *   rtattr
- *     data = struct xfrm_user_tmpl[]
- */
-#define NLMSG_BUF_SIZE 4096
-#define RTA_BUF_SIZE 2048
-#define XFRM_TMPLS_BUF_SIZE 1024
-#define CTX_BUF_SIZE 256
-
-static void usage(void) __attribute__((noreturn));
-
-static void usage(void)
-{
-	fprintf(stderr, "Usage: ip xfrm policy { add | update } SELECTOR dir DIR [ ctx CTX ]\n");
[…]

### 2019-11-15 iproute2/ip/xfrm_state.c

@@ -1,1470 +1,0 @@
-/* $USAGI: $ */
-
-/*
- * Copyright (C)2004 USAGI/WIDE Project
- *
- * This program is free software; you can redistribute it and/or modify
- * it under the terms of the GNU General Public License as published by
- * the Free Software Foundation; either version 2 of the License, or
- * (at your option) any later version.
- *
- * This program is distributed in the hope that it will be useful,
- * but WITHOUT ANY WARRANTY; without even the implied warranty of
- * MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
- * GNU General Public License for more details.
- *
- * You should have received a copy of the GNU General Public License
- * along with this program; if not, see <http://www.gnu.org/licenses>.
- */
-/*
- * based on iproute.c
- */
-/*
- * Authors:
- *	Masahide NAKAMURA @USAGI
- */
-
-#include <stdio.h>
-#include <stdlib.h>
-#include <string.h>
-#include <netdb.h>
-#include "utils.h"
-#include "xfrm.h"
-#include "ip_common.h"
-
-/* #define NLMSG_DELETEALL_BUF_SIZE (4096-512) */
-#define NLMSG_DELETEALL_BUF_SIZE 8192
-
-/*
- * Receiving buffer defines:
- * nlmsg
- *   data = struct xfrm_usersa_info
- *   rtattr
- *   rtattr
- *   ... (max count of rtattr is XFRM_MAX+1
- *
- *  each rtattr data = struct xfrm_algo(dynamic size) or xfrm_address_t
- */
-#define NLMSG_BUF_SIZE 4096
-#define RTA_BUF_SIZE 2048
-#define XFRM_ALGO_KEY_BUF_SIZE 512
-#define CTX_BUF_SIZE 256
-
-static void usage(void) __attribute__((noreturn));
-
-static void usage(void)
-{
-	fprintf(stderr, "Usage: ip xfrm state { add | update } ID [ ALGO-LIST ] [ mode MODE ]\n");
[…]

### 2019-11-15 iproute2/lib/bpf.c

@@ -1,3157 +1,0 @@
-/*
- * bpf.c	BPF common code
- *
- *		This program is free software; you can distribute it and/or
- *		modify it under the terms of the GNU General Public License
- *		as published by the Free Software Foundation; either version
- *		2 of the License, or (at your option) any later version.
- *
- * Authors:	Daniel Borkmann <daniel@iogearbox.net>
- *		Jiri Pirko <jiri@resnulli.us>
- *		Alexei Starovoitov <ast@kernel.org>
- */
-
-#include <stdio.h>
-#include <stdlib.h>
-#include <unistd.h>
-#include <string.h>
-#include <stdbool.h>
-#include <stdint.h>
-#include <errno.h>
-#include <fcntl.h>
-#include <stdarg.h>
-#include <limits.h>
-#include <assert.h>
-
-#ifdef HAVE_ELF
-#include <libelf.h>
-#include <gelf.h>
-#endif
-
-#include <sys/types.h>
-#include <sys/stat.h>
-#include <sys/un.h>
-#include <sys/vfs.h>
-#include <sys/mount.h>
-#include <sys/syscall.h>
-#include <sys/sendfile.h>
-#include <sys/resource.h>
-
-#include <arpa/inet.h>
-
-#include "utils.h"
-#include "json_print.h"
-
-#include "bpf_util.h"
-#include "bpf_elf.h"
-#include "bpf_scm.h"
-
-struct bpf_prog_meta {
-	const char *type;
-	const char *subdir;
-	const char *section;
-	bool may_uds_export;
-};
-
-static const enum bpf_prog_type __bpf_types[] = {
-	BPF_PROG_TYPE_SCHED_CLS,
-	BPF_PROG_TYPE_SCHED_ACT,
-	BPF_PROG_TYPE_XDP,
-	BPF_PROG_TYPE_LWT_IN,
-	BPF_PROG_TYPE_LWT_OUT,
-	BPF_PROG_TYPE_LWT_XMIT,
-};
-
-static const struct bpf_prog_meta __bpf_prog_meta[] = {
-	[BPF_PROG_TYPE_SCHED_CLS] = {
-		.type		= "cls",
-		.subdir		= "tc",
-		.section	= ELF_SECTION_CLASSIFIER,
-		.may_uds_export	= true,
-	},
-	[BPF_PROG_TYPE_SCHED_ACT] = {
-		.type		= "act",
-		.subdir		= "tc",
[…]

### 2019-11-15 iproute2/lib/json_writer.c

@@ -1,389 +1,0 @@
-/* SPDX-License-Identifier: (GPL-2.0 OR BSD-2-Clause) */
-/*
- * Simple streaming JSON writer
- *
- * This takes care of the annoying bits of JSON syntax like the commas
- * after elements
- *
- * Authors:	Stephen Hemminger <stephen@networkplumber.org>
- */
-
-#include <stdio.h>
-#include <stdbool.h>
-#include <stdarg.h>
-#include <assert.h>
-#include <malloc.h>
-#include <inttypes.h>
-#include <stdint.h>
-
-#include "json_writer.h"
-
-struct json_writer {
-	FILE		*out;	/* output file */
-	unsigned	depth;  /* nesting */
-	bool		pretty; /* optional whitepace */
-	char		sep;	/* either nul or comma */
-};
-
-/* indentation for pretty print */
-static void jsonw_indent(json_writer_t *self)
-{
-	unsigned i;
-	for (i = 0; i < self->depth; ++i)
-		fputs("    ", self->out);
-}
-
-/* end current line and indent if pretty printing */
-static void jsonw_eol(json_writer_t *self)
-{
-	if (!self->pretty)
-		return;
-
-	putc('\n', self->out);
-	jsonw_indent(self);
-}
-
-/* If current object is not empty print a comma */
-static void jsonw_eor(json_writer_t *self)
-{
-	if (self->sep != '\0')
-		putc(self->sep, self->out);
-	self->sep = ',';
-}
-
-
-/* Output JSON encoded string */
-/* Handles C escapes, does not do Unicode */
-static void jsonw_puts(json_writer_t *self, const char *str)
-{
-	putc('"', self->out);
-	for (; *str; ++str)
-		switch (*str) {
-		case '\t':
-			fputs("\\t", self->out);
-			break;
-		case '\n':
-			fputs("\\n", self->out);
-			break;
-		case '\r':
-			fputs("\\r", self->out);
-			break;
-		case '\f':
-			fputs("\\f", self->out);
-			break;
-		case '\b':
-			fputs("\\b", self->out);
-			break;
-		case '\\':
-			fputs("\\n", self->out);
-			break;
-		case '"':
-			fputs("\\\"", self->out);
-			break;
-		case '\'':
[…]

### 2019-11-15 iproute2/lib/libnetlink.c

@@ -1,1399 +1,0 @@
-/*
- * libnetlink.c	RTnetlink service routines.
- *
- *		This program is free software; you can redistribute it and/or
- *		modify it under the terms of the GNU General Public License
- *		as published by the Free Software Foundation; either version
- *		2 of the License, or (at your option) any later version.
- *
- * Authors:	Alexey Kuznetsov, <kuznet@ms2.inr.ac.ru>
- *
- */
-
-#include <stdio.h>
-#include <stdlib.h>
-#include <stdbool.h>
-#include <unistd.h>
-#include <fcntl.h>
-#include <net/if_arp.h>
-#include <sys/socket.h>
-#include <netinet/in.h>
-#include <string.h>
-#include <errno.h>
-#include <time.h>
-#include <sys/uio.h>
-#include <linux/fib_rules.h>
-#include <linux/if_addrlabel.h>
-#include <linux/if_bridge.h>
-
-#include "libnetlink.h"
-
-#define __aligned(x)		__attribute__((aligned(x)))
-
-#ifndef SOL_NETLINK
-#define SOL_NETLINK 270
-#endif
-
-#ifndef MIN
-#define MIN(a, b) ((a) < (b) ? (a) : (b))
-#endif
-
-int rcvbuf = 1024 * 1024;
-
-#ifdef HAVE_LIBMNL
-#include <libmnl/libmnl.h>
-
-static const enum mnl_attr_data_type extack_policy[NLMSGERR_ATTR_MAX + 1] = {
-	[NLMSGERR_ATTR_MSG]	= MNL_TYPE_NUL_STRING,
-	[NLMSGERR_ATTR_OFFS]	= MNL_TYPE_U32,
-};
-
-static int err_attr_cb(const struct nlattr *attr, void *data)
-{
-	const struct nlattr **tb = data;
-	uint16_t type;
-
-	if (mnl_attr_type_valid(attr, NLMSGERR_ATTR_MAX) < 0) {
-		fprintf(stderr, "Invalid extack attribute\n");
-		return MNL_CB_ERROR;
-	}
-
-	type = mnl_attr_get_type(attr);
-	if (mnl_attr_validate(attr, extack_policy[type]) < 0) {
[…]

### 2019-11-15 iproute2/lib/ll_map.c

@@ -1,295 +1,0 @@
-/*
- * ll_map.c
- *
- *		This program is free software; you can redistribute it and/or
- *		modify it under the terms of the GNU General Public License
- *		as published by the Free Software Foundation; either version
- *		2 of the License, or (at your option) any later version.
- *
- * Authors:	Alexey Kuznetsov, <kuznet@ms2.inr.ac.ru>
- *
- */
-
-#include <stdio.h>
-#include <stdlib.h>
-#include <unistd.h>
-#include <fcntl.h>
-#include <sys/socket.h>
-#include <netinet/in.h>
-#include <string.h>
-#include <net/if.h>
-
-#include "libnetlink.h"
-#include "ll_map.h"
-#include "list.h"
-
-struct ll_cache {
-	struct hlist_node idx_hash;
-	struct hlist_node name_hash;
-	unsigned	flags;
-	unsigned 	index;
-	unsigned short	type;
-	char		name[];
-};
-
-#define IDXMAP_SIZE	1024
-static struct hlist_head idx_head[IDXMAP_SIZE];
-static struct hlist_head name_head[IDXMAP_SIZE];
-
-static struct ll_cache *ll_get_by_index(unsigned index)
-{
-	struct hlist_node *n;
-	unsigned h = index & (IDXMAP_SIZE - 1);
-
-	hlist_for_each(n, &idx_head[h]) {
-		struct ll_cache *im
-			= container_of(n, struct ll_cache, idx_hash);
-		if (im->index == index)
-			return im;
-	}
-
-	return NULL;
-}
-
-unsigned namehash(const char *str)
-{
-	unsigned hash = 5381;
-
-	while (*str)
-		hash = ((hash << 5) + hash) + *str++; /* hash * 33 + c */
-
-	return hash;
-}
-
-static struct ll_cache *ll_get_by_name(const char *name)
-{
-	struct hlist_node *n;
-	unsigned h = namehash(name) & (IDXMAP_SIZE - 1);
-
-	hlist_for_each(n, &name_head[h]) {
-		struct ll_cache *im
-			= container_of(n, struct ll_cache, name_hash);
-
[…]

### 2019-11-15 iproute2/lib/ll_types.c

@@ -1,121 +1,0 @@
-/*
- * ll_types.c
- *
- *		This program is free software; you can redistribute it and/or
- *		modify it under the terms of the GNU General Public License
- *		as published by the Free Software Foundation; either version
- *		2 of the License, or (at your option) any later version.
- *
- * Authors:	Alexey Kuznetsov, <kuznet@ms2.inr.ac.ru>
- */
-
-#include <stdio.h>
-#include <stdlib.h>
-#include <unistd.h>
-#include <fcntl.h>
-#include <sys/ioctl.h>
-#include <sys/socket.h>
-#include <netinet/in.h>
-#include <arpa/inet.h>
-#include <string.h>
-
-#include <linux/netdevice.h>
-#include <linux/if_arp.h>
-#include <linux/sockios.h>
-
-#include "rt_names.h"
-
-const char * ll_type_n2a(int type, char *buf, int len)
-{
-#define __PF(f,n) { ARPHRD_##f, #n },
-static const struct {
-	int type;
-	const char *name;
-} arphrd_names[] = {
-__PF(NETROM,netrom)
-__PF(ETHER,ether)
-__PF(EETHER,eether)
-__PF(AX25,ax25)
-__PF(PRONET,pronet)
-__PF(CHAOS,chaos)
-__PF(IEEE802,ieee802)
-__PF(ARCNET,arcnet)
-__PF(APPLETLK,atalk)
-__PF(DLCI,dlci)
-__PF(ATM,atm)
-__PF(METRICOM,metricom)
-__PF(IEEE1394,ieee1394)
-__PF(INFINIBAND,infiniband)
-__PF(SLIP,slip)
-__PF(CSLIP,cslip)
-__PF(SLIP6,slip6)
-__PF(CSLIP6,cslip6)
-__PF(RSRVD,rsrvd)
-__PF(ADAPT,adapt)
-__PF(ROSE,rose)
-__PF(X25,x25)
-__PF(HWX25,hwx25)
-__PF(CAN,can)
-__PF(PPP,ppp)
-__PF(HDLC,hdlc)
-__PF(LAPB,lapb)
-__PF(DDCMP,ddcmp)
[…]

### 2019-11-15 iproute2/misc/nstat.c

@@ -1,756 +1,0 @@
-/*
- * nstat.c	handy utility to read counters /proc/net/netstat and snmp
- *
- *		This program is free software; you can redistribute it and/or
- *		modify it under the terms of the GNU General Public License
- *		as published by the Free Software Foundation; either version
- *		2 of the License, or (at your option) any later version.
- *
- * Authors:	Alexey Kuznetsov, <kuznet@ms2.inr.ac.ru>
- */
-
-#include <stdio.h>
-#include <stdlib.h>
-#include <unistd.h>
-#include <fcntl.h>
-#include <string.h>
-#include <errno.h>
-#include <time.h>
-#include <sys/time.h>
-#include <fnmatch.h>
-#include <sys/file.h>
-#include <sys/socket.h>
-#include <sys/un.h>
-#include <sys/poll.h>
-#include <sys/wait.h>
-#include <sys/stat.h>
-#include <signal.h>
-#include <math.h>
-#include <getopt.h>
-
-#include <json_writer.h>
-#include <SNAPSHOT.h>
-#include "utils.h"
-
-int dump_zeros;
-int reset_history;
-int ignore_history;
-int no_output;
-int json_output;
-int no_update;
-int scan_interval;
-int time_constant;
-double W;
-char **patterns;
-int npatterns;
-
-char info_source[128];
-int source_mismatch;
-
-static int generic_proc_open(const char *env, char *name)
-{
-	char store[128];
-	char *p = getenv(env);
-
-	if (!p) {
-		p = getenv("PROC_ROOT") ? : "/proc";
-		snprintf(store, sizeof(store)-1, "%s/%s", p, name);
-		p = store;
-	}
-	return open(p, O_RDONLY);
-}
-
-static int net_netstat_open(void)
-{
-	return generic_proc_open("PROC_NET_NETSTAT", "net/netstat");
-}
-
-static int net_snmp_open(void)
-{
-	return generic_proc_open("PROC_NET_SNMP", "net/snmp");
-}
-
-static int net_snmp6_open(void)
-{
-	return generic_proc_open("PROC_NET_SNMP6", "net/snmp6");
-}
-
-static int net_sctp_snmp_open(void)
-{
[…]

### 2019-11-15 iproute2/misc/ss.c

@@ -1,5353 +1,0 @@
-/*
- * ss.c		"sockstat", socket statistics
- *
- *		This program is free software; you can redistribute it and/or
- *		modify it under the terms of the GNU General Public License
- *		as published by the Free Software Foundation; either version
- *		2 of the License, or (at your option) any later version.
- *
- * Authors:	Alexey Kuznetsov, <kuznet@ms2.inr.ac.ru>
- */
-
-#include <stdio.h>
-#include <stdlib.h>
-#include <unistd.h>
-#include <fcntl.h>
-#include <sys/ioctl.h>
-#include <sys/socket.h>
-#include <sys/uio.h>
-#include <sys/sysmacros.h>
-#include <netinet/in.h>
-#include <string.h>
-#include <errno.h>
-#include <netdb.h>
-#include <arpa/inet.h>
-#include <dirent.h>
-#include <fnmatch.h>
-#include <getopt.h>
-#include <stdbool.h>
-#include <limits.h>
-#include <stdarg.h>
-
-#include "utils.h"
-#include "rt_names.h"
-#include "ll_map.h"
-#include "libnetlink.h"
-#include "namespace.h"
-#include "SNAPSHOT.h"
-
-#include <linux/tcp.h>
-#include <linux/sock_diag.h>
-#include <linux/inet_diag.h>
-#include <linux/unix_diag.h>
-#include <linux/netdevice.h>	/* for MAX_ADDR_LEN */
-#include <linux/filter.h>
-#include <linux/xdp_diag.h>
-#include <linux/packet_diag.h>
-#include <linux/netlink_diag.h>
-#include <linux/sctp.h>
-#include <linux/vm_sockets_diag.h>
-#include <linux/net.h>
-#include <linux/tipc.h>
-#include <linux/tipc_netlink.h>
-#include <linux/tipc_sockets_diag.h>
-
-/* AF_VSOCK/PF_VSOCK is only provided since glibc 2.18 */
-#ifndef PF_VSOCK
-#define PF_VSOCK 40
-#endif
-#ifndef AF_VSOCK
[…]

### 2019-11-15 iproute2/rdma/dev.c

@@ -1,307 +1,0 @@
-// SPDX-License-Identifier: GPL-2.0 OR Linux-OpenIB
-/*
- * dev.c	RDMA tool
- * Authors:     Leon Romanovsky <leonro@mellanox.com>
- */
-
-#include "rdma.h"
-
-static int dev_help(struct rd *rd)
-{
-	pr_out("Usage: %s dev show [DEV]\n", rd->filename);
-	pr_out("       %s dev set [DEV] name DEVNAME\n", rd->filename);
-	return 0;
-}
-
-static const char *dev_caps_to_str(uint32_t idx)
-{
-#define RDMA_DEV_FLAGS_LOW(x) \
-	x(RESIZE_MAX_WR, 0) \
-	x(BAD_PKEY_CNTR, 1) \
-	x(BAD_QKEY_CNTR, 2) \
-	x(RAW_MULTI, 3) \
-	x(AUTO_PATH_MIG, 4) \
-	x(CHANGE_PHY_PORT, 5) \
-	x(UD_AV_PORT_ENFORCE_PORT_ENFORCE, 6) \
-	x(CURR_QP_STATE_MOD, 7) \
-	x(SHUTDOWN_PORT, 8) \
-	x(INIT_TYPE, 9) \
-	x(PORT_ACTIVE_EVENT, 10) \
-	x(SYS_IMAGE_GUID, 11) \
-	x(RC_RNR_NAK_GEN, 12) \
-	x(SRQ_RESIZE, 13) \
-	x(N_NOTIFY_CQ, 14) \
-	x(LOCAL_DMA_LKEY, 15) \
-	x(MEM_WINDOW, 17) \
-	x(UD_IP_CSUM, 18) \
-	x(UD_TSO, 19) \
-	x(XRC, 20) \
-	x(MEM_MGT_EXTENSIONS, 21) \
-	x(BLOCK_MULTICAST_LOOPBACK, 22) \
-	x(MEM_WINDOW_TYPE_2A, 23) \
-	x(MEM_WINDOW_TYPE_2B, 24) \
-	x(RC_IP_CSUM, 25) \
-	x(RAW_IP_CSUM, 26) \
-	x(CROSS_CHANNEL, 27) \
-	x(MANAGED_FLOW_STEERING, 29) \
-	x(SIGNATURE_HANDOVER, 30) \
-	x(ON_DEMAND_PAGING, 31)
-
-#define RDMA_DEV_FLAGS_HIGH(x) \
[…]

### 2019-11-15 iproute2/rdma/include/uapi/rdma/rdma_netlink.h

@@ -1,480 +1,0 @@
-/* SPDX-License-Identifier: GPL-2.0 WITH Linux-syscall-note */
-#ifndef _RDMA_NETLINK_H
-#define _RDMA_NETLINK_H
-
-#include <linux/types.h>
-
-enum {
-	RDMA_NL_IWCM = 2,
-	RDMA_NL_RSVD,
-	RDMA_NL_LS,	/* RDMA Local Services */
-	RDMA_NL_NLDEV,	/* RDMA device interface */
-	RDMA_NL_NUM_CLIENTS
-};
-
-enum {
-	RDMA_NL_GROUP_IWPM = 2,
-	RDMA_NL_GROUP_LS,
-	RDMA_NL_NUM_GROUPS
-};
-
-#define RDMA_NL_GET_CLIENT(type) ((type & (((1 << 6) - 1) << 10)) >> 10)
-#define RDMA_NL_GET_OP(type) (type & ((1 << 10) - 1))
-#define RDMA_NL_GET_TYPE(client, op) ((client << 10) + op)
-
-/* The minimum version that the iwpm kernel supports */
-#define IWPM_UABI_VERSION_MIN	3
-
-/* The latest version that the iwpm kernel supports */
-#define IWPM_UABI_VERSION	4
-
-/* iwarp port mapper message flags */
-enum {
-
-	/* Do not map the port for this IWPM request */
-	IWPM_FLAGS_NO_PORT_MAP = (1 << 0),
-};
-
-/* iwarp port mapper op-codes */
-enum {
-	RDMA_NL_IWPM_REG_PID = 0,
-	RDMA_NL_IWPM_ADD_MAPPING,
-	RDMA_NL_IWPM_QUERY_MAPPING,
-	RDMA_NL_IWPM_REMOVE_MAPPING,
-	RDMA_NL_IWPM_REMOTE_INFO,
-	RDMA_NL_IWPM_HANDLE_ERR,
-	RDMA_NL_IWPM_MAPINFO,
-	RDMA_NL_IWPM_MAPINFO_NUM,
-	RDMA_NL_IWPM_HELLO,
-	RDMA_NL_IWPM_NUM_OPS
-};
-
-struct rdma_cm_id_stats {
-	__u32	qp_num;
-	__u32	bound_dev_if;
-	__u32	port_space;
-	__s32	pid;
-	__u8	cm_state;
-	__u8	node_type;
-	__u8	port_num;
-	__u8	qp_type;
-};
-
-enum {
[…]

### 2019-11-15 iproute2/rdma/rdma.c

@@ -1,198 +1,0 @@
-// SPDX-License-Identifier: GPL-2.0 OR Linux-OpenIB
-/*
- * rdma.c	RDMA tool
- * Authors:     Leon Romanovsky <leonro@mellanox.com>
- */
-
-#include "rdma.h"
-#include "SNAPSHOT.h"
-
-static void help(char *name)
-{
-	pr_out("Usage: %s [ OPTIONS ] OBJECT { COMMAND | help }\n"
-	       "       %s [ -f[orce] ] -b[atch] filename\n"
-	       "where  OBJECT := { dev | link | resource | help }\n"
-	       "       OPTIONS := { -V[ersion] | -d[etails] | -j[son] | -p[retty]}\n", name, name);
-}
-
-static int cmd_help(struct rd *rd)
-{
-	help(rd->filename);
-	return 0;
-}
-
-static int rd_cmd(struct rd *rd, int argc, char **argv)
-{
-	const struct rd_cmd cmds[] = {
-		{ NULL,		cmd_help },
-		{ "help",	cmd_help },
-		{ "dev",	cmd_dev },
-		{ "link",	cmd_link },
-		{ "resource",	cmd_res },
-		{ 0 }
-	};
-
-	rd->argc = argc;
-	rd->argv = argv;
-
-	return rd_exec_cmd(rd, cmds, "object");
-}
-
-static int rd_batch(struct rd *rd, const char *name, bool force)
-{
-	char *line = NULL;
-	size_t len = 0;
-	int ret = 0;
-
-	if (name && strcmp(name, "-") != 0) {
-		if (!freopen(name, "r", stdin)) {
-			pr_err("Cannot open file \"%s\" for reading: %s\n",
-			       name, strerror(errno));
-			return errno;
-		}
-	}
-
-	cmdlineno = 0;
-	while (getcmdline(&line, &len, stdin) != -1) {
-		char *largv[512];
-		int largc;
-
-		largc = makeargs(line, largv, ARRAY_SIZE(largv));
-		if (!largc)
-			continue;	/* blank line */
-
-		ret = rd_cmd(rd, largc, largv);
-		if (ret) {
-			pr_err("Command failed %s:%d\n", name, cmdlineno);
-			if (!force)
-				break;
-		}
-	}
-
-	free(line);
-
-	return ret;
-}
-
-static int rd_init(struct rd *rd, char *filename)
-{
-	uint32_t seq;
-	int ret;
-
-	rd->filename = filename;
[…]

### 2019-11-15 iproute2/rdma/rdma.h

@@ -1,139 +1,0 @@
-/* SPDX-License-Identifier: GPL-2.0 OR Linux-OpenIB */
-/*
- * rdma.c	RDMA tool
- * Authors:     Leon Romanovsky <leonro@mellanox.com>
- */
-#ifndef _RDMA_TOOL_H_
-#define _RDMA_TOOL_H_
-
-#include <stdlib.h>
-#include <string.h>
-#include <errno.h>
-#include <getopt.h>
-#include <netinet/in.h>
-#include <libmnl/libmnl.h>
-#include <rdma/rdma_netlink.h>
-#include <rdma/rdma_user_cm.h>
-#include <time.h>
-#include <net/if_arp.h>
-
-#include "list.h"
-#include "utils.h"
-#include "json_writer.h"
-
-#define pr_err(args...) fprintf(stderr, ##args)
-#define pr_out(args...) fprintf(stdout, ##args)
-
-#define RDMA_BITMAP_ENUM(name, bit_no) RDMA_BITMAP_##name = BIT(bit_no),
-#define RDMA_BITMAP_NAMES(name, bit_no) [bit_no] = #name,
-
-#define MAX_NUMBER_OF_FILTERS 64
-struct filters {
-	const char *name;
-	uint8_t is_number:1;
-	uint8_t is_doit:1;
-};
-
-struct filter_entry {
-	struct list_head list;
-	char *key;
-	char *value;
-	/*
-	 * This field menas that we can try to issue .doit calback
-	 * on value above. This value can be converted to integer
-	 * with simple atoi(). Otherwise "is_doit" will be false.
-	 */
-	uint8_t is_doit:1;
-};
-
-struct dev_map {
-	struct list_head list;
-	char *dev_name;
-	uint32_t num_ports;
-	uint32_t idx;
-};
-
-struct rd {
-	int argc;
-	char **argv;
-	char *filename;
-	bool show_details;
-	bool show_driver_details;
-	struct list_head dev_map_list;
-	uint32_t dev_idx;
-	uint32_t port_idx;
-	struct mnl_socket *nl;
-	struct nlmsghdr *nlh;
-	char *buff;
-	json_writer_t *jw;
-	bool json_output;
-	bool pretty_output;
-	bool suppress_errors;
-	struct list_head filter_list;
-	char *link_name;
-	char *link_type;
-};
-
-struct rd_cmd {
-	const char *cmd;
-	int (*func)(struct rd *rd);
-};
-
-/*
- * Parser interface
- */
[…]

### 2019-11-15 iproute2/rdma/res-cmid.c

@@ -1,275 +1,0 @@
-// SPDX-License-Identifier: GPL-2.0 OR Linux-OpenIB
-/*
- * res-cmid.c	RDMA tool
- * Authors:     Leon Romanovsky <leonro@mellanox.com>
- */
-
-#include "res.h"
-#include <inttypes.h>
-
-static void print_qp_type(struct rd *rd, uint32_t val)
-{
-	if (rd->json_output)
-		jsonw_string_field(rd->jw, "qp-type", qp_types_to_str(val));
-	else
-		pr_out("qp-type %s ", qp_types_to_str(val));
-}
-
-static const char *cm_id_state_to_str(uint8_t idx)
-{
-	static const char *const cm_id_states_str[] = {
-		"IDLE",		  "ADDR_QUERY",     "ADDR_RESOLVED",
-		"ROUTE_QUERY",    "ROUTE_RESOLVED", "CONNECT",
-		"DISCONNECT",     "ADDR_BOUND",     "LISTEN",
-		"DEVICE_REMOVAL", "DESTROYING"
-	};
-
-	if (idx < ARRAY_SIZE(cm_id_states_str))
-		return cm_id_states_str[idx];
-	return "UNKNOWN";
-}
-
-static const char *cm_id_ps_to_str(uint32_t ps)
-{
-	switch (ps) {
-	case RDMA_PS_IPOIB:
-		return "IPoIB";
-	case RDMA_PS_IB:
-		return "IPoIB";
-	case RDMA_PS_TCP:
-		return "TCP";
-	case RDMA_PS_UDP:
-		return "UDP";
-	default:
-		return "---";
-	}
-}
-
-static void print_cm_id_state(struct rd *rd, uint8_t state)
-{
-	if (rd->json_output) {
-		jsonw_string_field(rd->jw, "state", cm_id_state_to_str(state));
-		return;
-	}
-	pr_out("state %s ", cm_id_state_to_str(state));
-}
-
-static void print_ps(struct rd *rd, uint32_t ps)
-{
-	if (rd->json_output) {
-		jsonw_string_field(rd->jw, "ps", cm_id_ps_to_str(ps));
-		return;
-	}
-	pr_out("ps %s ", cm_id_ps_to_str(ps));
-}
-
[…]

### 2019-11-15 iproute2/rdma/res-cq.c

@@ -1,160 +1,0 @@
-// SPDX-License-Identifier: GPL-2.0 OR Linux-OpenIB
-/*
- * res-cq.c	RDMA tool
- * Authors:     Leon Romanovsky <leonro@mellanox.com>
- */
-
-#include "res.h"
-#include <inttypes.h>
-
-static const char *poll_ctx_to_str(uint8_t idx)
-{
-	static const char * const cm_id_states_str[] = {
-		"DIRECT", "SOFTIRQ", "WORKQUEUE", "UNBOUND_WORKQUEUE"};
-
-	if (idx < ARRAY_SIZE(cm_id_states_str))
-		return cm_id_states_str[idx];
-	return "UNKNOWN";
-}
-
-static void print_poll_ctx(struct rd *rd, uint8_t poll_ctx, struct nlattr *attr)
-{
-	if (!attr)
-		return;
-
-	if (rd->json_output) {
-		jsonw_string_field(rd->jw, "poll-ctx",
-				   poll_ctx_to_str(poll_ctx));
-		return;
-	}
-	pr_out("poll-ctx %s ", poll_ctx_to_str(poll_ctx));
-}
-
-static int res_cq_line(struct rd *rd, const char *name, int idx,
-		       struct nlattr **nla_line)
-{
-	char *comm = NULL;
-	uint32_t pid = 0;
-	uint8_t poll_ctx = 0;
-	uint32_t ctxn = 0;
-	uint32_t cqn = 0;
-	uint64_t users;
-	uint32_t cqe;
-
-	if (!nla_line[RDMA_NLDEV_ATTR_RES_CQE] ||
-	    !nla_line[RDMA_NLDEV_ATTR_RES_USECNT] ||
-	    (!nla_line[RDMA_NLDEV_ATTR_RES_PID] &&
-	     !nla_line[RDMA_NLDEV_ATTR_RES_KERN_NAME])) {
-		return MNL_CB_ERROR;
-	}
-
-	cqe = mnl_attr_get_u32(nla_line[RDMA_NLDEV_ATTR_RES_CQE]);
-
-	users = mnl_attr_get_u64(nla_line[RDMA_NLDEV_ATTR_RES_USECNT]);
-	if (rd_is_filtered_attr(rd, "users", users,
[…]

### 2019-11-15 iproute2/rdma/res-mr.c

@@ -1,133 +1,0 @@
-// SPDX-License-Identifier: GPL-2.0 OR Linux-OpenIB
-/*
- * res-mr.c	RDMA tool
- * Authors:     Leon Romanovsky <leonro@mellanox.com>
- */
-
-#include "res.h"
-#include <inttypes.h>
-
-static int res_mr_line(struct rd *rd, const char *name, int idx,
-		       struct nlattr **nla_line)
-{
-	uint32_t rkey = 0, lkey = 0;
-	uint64_t iova = 0, mrlen;
-	char *comm = NULL;
-	uint32_t pdn = 0;
-	uint32_t mrn = 0;
-	uint32_t pid = 0;
-
-	if (!nla_line[RDMA_NLDEV_ATTR_RES_MRLEN] ||
-	    (!nla_line[RDMA_NLDEV_ATTR_RES_PID] &&
-	     !nla_line[RDMA_NLDEV_ATTR_RES_KERN_NAME])) {
-		return MNL_CB_ERROR;
-	}
-
-	if (nla_line[RDMA_NLDEV_ATTR_RES_RKEY])
-		rkey = mnl_attr_get_u32(nla_line[RDMA_NLDEV_ATTR_RES_RKEY]);
-	if (nla_line[RDMA_NLDEV_ATTR_RES_LKEY])
-		lkey = mnl_attr_get_u32(nla_line[RDMA_NLDEV_ATTR_RES_LKEY]);
-	if (nla_line[RDMA_NLDEV_ATTR_RES_IOVA])
-		iova = mnl_attr_get_u64(nla_line[RDMA_NLDEV_ATTR_RES_IOVA]);
-
-	mrlen = mnl_attr_get_u64(nla_line[RDMA_NLDEV_ATTR_RES_MRLEN]);
-	if (rd_is_filtered_attr(rd, "mrlen", mrlen,
-				nla_line[RDMA_NLDEV_ATTR_RES_MRLEN]))
-		goto out;
-
-	if (nla_line[RDMA_NLDEV_ATTR_RES_PID]) {
-		pid = mnl_attr_get_u32(nla_line[RDMA_NLDEV_ATTR_RES_PID]);
-		comm = get_task_name(pid);
-	}
-
-	if (rd_is_filtered_attr(rd, "pid", pid,
-				nla_line[RDMA_NLDEV_ATTR_RES_PID]))
-		goto out;
-
[…]

### 2019-11-15 iproute2/rdma/res-pd.c

@@ -1,136 +1,0 @@
-// SPDX-License-Identifier: GPL-2.0 OR Linux-OpenIB
-/*
- * res-pd.c	RDMA tool
- * Authors:     Leon Romanovsky <leonro@mellanox.com>
- */
-
-#include "res.h"
-#include <inttypes.h>
-
-static int res_pd_line(struct rd *rd, const char *name, int idx,
-		       struct nlattr **nla_line)
-{
-	uint32_t local_dma_lkey = 0, unsafe_global_rkey = 0;
-	char *comm = NULL;
-	uint32_t ctxn = 0;
-	uint32_t pid = 0;
-	uint32_t pdn = 0;
-	uint64_t users;
-
-	if (!nla_line[RDMA_NLDEV_ATTR_RES_USECNT] ||
-	    (!nla_line[RDMA_NLDEV_ATTR_RES_PID] &&
-	     !nla_line[RDMA_NLDEV_ATTR_RES_KERN_NAME])) {
-		return MNL_CB_ERROR;
-	}
-
-	if (nla_line[RDMA_NLDEV_ATTR_RES_LOCAL_DMA_LKEY])
-		local_dma_lkey = mnl_attr_get_u32(
-			nla_line[RDMA_NLDEV_ATTR_RES_LOCAL_DMA_LKEY]);
-
-	users = mnl_attr_get_u64(nla_line[RDMA_NLDEV_ATTR_RES_USECNT]);
-	if (rd_is_filtered_attr(rd, "users", users,
-				nla_line[RDMA_NLDEV_ATTR_RES_USECNT]))
-		goto out;
-
-	if (nla_line[RDMA_NLDEV_ATTR_RES_UNSAFE_GLOBAL_RKEY])
-		unsafe_global_rkey = mnl_attr_get_u32(
-			nla_line[RDMA_NLDEV_ATTR_RES_UNSAFE_GLOBAL_RKEY]);
-
-	if (nla_line[RDMA_NLDEV_ATTR_RES_PID]) {
-		pid = mnl_attr_get_u32(nla_line[RDMA_NLDEV_ATTR_RES_PID]);
-		comm = get_task_name(pid);
-	}
-
-	if (rd_is_filtered_attr(rd, "pid", pid,
-				nla_line[RDMA_NLDEV_ATTR_RES_PID]))
-		goto out;
-
-	if (nla_line[RDMA_NLDEV_ATTR_RES_CTXN])
[…]

### 2019-11-15 iproute2/rdma/res-qp.c

@@ -1,240 +1,0 @@
-// SPDX-License-Identifier: GPL-2.0 OR Linux-OpenIB
-/*
- * res-qp.c	RDMA tool
- * Authors:     Leon Romanovsky <leonro@mellanox.com>
- */
-
-#include "res.h"
-#include <inttypes.h>
-
-static const char *path_mig_to_str(uint8_t idx)
-{
-	static const char *const path_mig_str[] = { "MIGRATED", "REARM",
-						    "ARMED" };
-
-	if (idx < ARRAY_SIZE(path_mig_str))
-		return path_mig_str[idx];
-	return "UNKNOWN";
-}
-
-static const char *qp_states_to_str(uint8_t idx)
-{
-	static const char *const qp_states_str[] = { "RESET", "INIT", "RTR",
-						     "RTS",   "SQD",  "SQE",
-						     "ERR" };
-
-	if (idx < ARRAY_SIZE(qp_states_str))
-		return qp_states_str[idx];
-	return "UNKNOWN";
-}
-
-static void print_rqpn(struct rd *rd, uint32_t val, struct nlattr **nla_line)
-{
-	if (!nla_line[RDMA_NLDEV_ATTR_RES_RQPN])
-		return;
-
-	if (rd->json_output)
-		jsonw_uint_field(rd->jw, "rqpn", val);
-	else
-		pr_out("rqpn %u ", val);
-}
-
-static void print_type(struct rd *rd, uint32_t val)
-{
-	if (rd->json_output)
-		jsonw_string_field(rd->jw, "type", qp_types_to_str(val));
-	else
-		pr_out("type %s ", qp_types_to_str(val));
-}
-
-static void print_state(struct rd *rd, uint32_t val)
-{
-	if (rd->json_output)
-		jsonw_string_field(rd->jw, "state", qp_states_to_str(val));
-	else
-		pr_out("state %s ", qp_states_to_str(val));
-}
-
-static void print_rqpsn(struct rd *rd, uint32_t val, struct nlattr **nla_line)
-{
[…]

### 2019-11-15 iproute2/rdma/res.c

@@ -1,283 +1,0 @@
-// SPDX-License-Identifier: GPL-2.0 OR Linux-OpenIB
-/*
- * res.c	RDMA tool
- * Authors:     Leon Romanovsky <leonro@mellanox.com>
- */
-
-#include "res.h"
-#include <inttypes.h>
-
-static int res_help(struct rd *rd)
-{
-	pr_out("Usage: %s resource\n", rd->filename);
-	pr_out("          resource show [DEV]\n");
-	pr_out("          resource show [qp|cm_id|pd|mr|cq]\n");
-	pr_out("          resource show qp link [DEV/PORT]\n");
-	pr_out("          resource show qp link [DEV/PORT] [FILTER-NAME FILTER-VALUE]\n");
-	pr_out("          resource show cm_id link [DEV/PORT]\n");
-	pr_out("          resource show cm_id link [DEV/PORT] [FILTER-NAME FILTER-VALUE]\n");
-	pr_out("          resource show cq link [DEV/PORT]\n");
-	pr_out("          resource show cq link [DEV/PORT] [FILTER-NAME FILTER-VALUE]\n");
-	pr_out("          resource show pd dev [DEV]\n");
-	pr_out("          resource show pd dev [DEV] [FILTER-NAME FILTER-VALUE]\n");
-	pr_out("          resource show mr dev [DEV]\n");
-	pr_out("          resource show mr dev [DEV] [FILTER-NAME FILTER-VALUE]\n");
-	return 0;
-}
-
-static int res_print_summary(struct rd *rd, struct nlattr **tb)
-{
-	struct nlattr *nla_table = tb[RDMA_NLDEV_ATTR_RES_SUMMARY];
-	struct nlattr *nla_entry;
-	const char *name;
-	uint64_t curr;
-	int err;
-
-	mnl_attr_for_each_nested(nla_entry, nla_table) {
-		struct nlattr *nla_line[RDMA_NLDEV_ATTR_MAX] = {};
-
-		err = mnl_attr_parse_nested(nla_entry, rd_attr_cb, nla_line);
-		if (err != MNL_CB_OK)
-			return -EINVAL;
-
[…]

### 2019-11-15 iproute2/rdma/utils.c

@@ -1,929 +1,0 @@
-// SPDX-License-Identifier: GPL-2.0 OR Linux-OpenIB
-/*
- * utils.c	RDMA tool
- * Authors:     Leon Romanovsky <leonro@mellanox.com>
- */
-
-#include "rdma.h"
-#include <ctype.h>
-#include <inttypes.h>
-
-int rd_argc(struct rd *rd)
-{
-	return rd->argc;
-}
-
-char *rd_argv(struct rd *rd)
-{
-	if (!rd_argc(rd))
-		return NULL;
-	return *rd->argv;
-}
-
-int strcmpx(const char *str1, const char *str2)
-{
-	if (strlen(str1) > strlen(str2))
-		return -1;
-	return strncmp(str1, str2, strlen(str1));
-}
-
-static bool rd_argv_match(struct rd *rd, const char *pattern)
-{
-	if (!rd_argc(rd))
-		return false;
-	return strcmpx(rd_argv(rd), pattern) == 0;
-}
-
-void rd_arg_inc(struct rd *rd)
-{
-	if (!rd_argc(rd))
-		return;
-	rd->argc--;
-	rd->argv++;
-}
-
-bool rd_no_arg(struct rd *rd)
-{
-	return rd_argc(rd) == 0;
-}
-
-/*
- * Possible input:output
- * dev/port    | first port | is_dump_all
- * mlx5_1      | 0          | true
- * mlx5_1/     | 0          | true
- * mlx5_1/0    | 0          | false
- * mlx5_1/1    | 1          | false
- * mlx5_1/-    | 0          | false
- *
- * In strict mode, /- will return error.
- */
-static int get_port_from_argv(struct rd *rd, uint32_t *port,
-			      bool *is_dump_all, bool strict_port)
-{
-	char *slash;
-
-	*port = 0;
-	*is_dump_all = true;
-
-	slash = strchr(rd_argv(rd), '/');
-	/* if no port found, return 0 */
-	if (slash++) {
-		if (*slash == '-') {
-			if (strict_port)
-				return -EINVAL;
-			*is_dump_all = false;
-			return 0;
-		}
-
-		if (isdigit(*slash)) {
[…]

### 2019-11-15 iproute2/tc/emp_ematch.yacc.h

@@ -1,106 +1,0 @@
-/* A Bison parser, made by GNU Bison 3.3.2.  */
-
-/* Bison interface for Yacc-like parsers in C
-
-   Copyright (C) 1984, 1989-1990, 2000-2015, 2018-2019 Free Software Foundation,
-   Inc.
-
-   This program is free software: you can redistribute it and/or modify
-   it under the terms of the GNU General Public License as published by
-   the Free Software Foundation, either version 3 of the License, or
-   (at your option) any later version.
-
-   This program is distributed in the hope that it will be useful,
-   but WITHOUT ANY WARRANTY; without even the implied warranty of
-   MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
-   GNU General Public License for more details.
-
-   You should have received a copy of the GNU General Public License
-   along with this program.  If not, see <http://www.gnu.org/licenses/>.  */
-
-/* As a special exception, you may create a larger work that contains
-   part or all of the Bison parser skeleton and distribute that work
-   under terms of your choice, so long as that work isn't itself a
-   parser generator using the skeleton or a modified version thereof
-   as a parser skeleton.  Alternatively, if you modify or redistribute
-   the parser skeleton itself, you may (at your option) remove this
-   special exception, which will cause the skeleton and the resulting
-   Bison output files to be licensed under the GNU General Public
-   License without this special exception.
-
-   This special exception was added by the Free Software Foundation in
-   version 2.2 of Bison.  */
-
-/* Undocumented macros, especially those whose name start with YY_,
-   are private implementation details.  Do not rely on them.  */
-
-#ifndef YY_EMATCH_EMP_EMATCH_YACC_H_INCLUDED
-# define YY_EMATCH_EMP_EMATCH_YACC_H_INCLUDED
-/* Debug traces.  */
-#ifndef EMATCH_DEBUG
-# if defined YYDEBUG
-#if YYDEBUG
-#   define EMATCH_DEBUG 1
-#  else
[…]

### 2019-11-15 iproute2/tc/f_bpf.c

@@ -1,267 +1,0 @@
-/*
- * f_bpf.c	BPF-based Classifier
- *
- *		This program is free software; you can distribute it and/or
- *		modify it under the terms of the GNU General Public License
- *		as published by the Free Software Foundation; either version
- *		2 of the License, or (at your option) any later version.
- *
- * Authors:	Daniel Borkmann <daniel@iogearbox.net>
- */
-
-#include <stdio.h>
-#include <stdlib.h>
-
-#include <linux/bpf.h>
-
-#include "utils.h"
-
-#include "tc_util.h"
-#include "bpf_util.h"
-
-static const enum bpf_prog_type bpf_type = BPF_PROG_TYPE_SCHED_CLS;
-
-static void explain(void)
-{
-	fprintf(stderr, "Usage: ... bpf ...\n");
-	fprintf(stderr, "\n");
-	fprintf(stderr, "BPF use case:\n");
-	fprintf(stderr, " bytecode BPF_BYTECODE\n");
-	fprintf(stderr, " bytecode-file FILE\n");
-	fprintf(stderr, "\n");
-	fprintf(stderr, "eBPF use case:\n");
-	fprintf(stderr, " object-file FILE [ section CLS_NAME ] [ export UDS_FILE ]");
-	fprintf(stderr, " [ verbose ] [ direct-action ] [ skip_hw | skip_sw ]\n");
-	fprintf(stderr, " object-pinned FILE [ direct-action ] [ skip_hw | skip_sw ]\n");
-	fprintf(stderr, "\n");
-	fprintf(stderr, "Common remaining options:\n");
-	fprintf(stderr, " [ action ACTION_SPEC ]\n");
-	fprintf(stderr, " [ classid CLASSID ]\n");
-	fprintf(stderr, "\n");
-	fprintf(stderr, "Where BPF_BYTECODE := \'s,c t f k,c t f k,c t f k,...\'\n");
-	fprintf(stderr, "c,t,f,k and s are decimals; s denotes number of 4-tuples\n");
-	fprintf(stderr, "\n");
-	fprintf(stderr, "Where FILE points to a file containing the BPF_BYTECODE string,\n");
[…]

### 2019-11-15 iproute2/tc/f_flower.c

@@ -1,1962 +1,0 @@
-/*
- * f_flower.c		Flower Classifier
- *
- *		This program is free software; you can distribute it and/or
- *		modify it under the terms of the GNU General Public License
- *		as published by the Free Software Foundation; either version
- *		2 of the License, or (at your option) any later version.
- *
- * Authors:     Jiri Pirko <jiri@resnulli.us>
- */
-
-#include <stdio.h>
-#include <stdlib.h>
-#include <unistd.h>
-#include <string.h>
-#include <net/if.h>
-#include <linux/limits.h>
-#include <linux/if_arp.h>
-#include <linux/if_ether.h>
-#include <linux/ip.h>
-#include <linux/tc_act/tc_vlan.h>
-#include <linux/mpls.h>
-
-#include "utils.h"
-#include "tc_util.h"
-#include "rt_names.h"
-
-enum flower_matching_flags {
-	FLOWER_IP_FLAGS,
-};
-
-enum flower_endpoint {
-	FLOWER_ENDPOINT_SRC,
-	FLOWER_ENDPOINT_DST
-};
-
-enum flower_icmp_field {
-	FLOWER_ICMP_FIELD_TYPE,
-	FLOWER_ICMP_FIELD_CODE
-};
-
-static void explain(void)
-{
-	fprintf(stderr,
-		"Usage: ... flower [ MATCH-LIST ] [ verbose ]\n"
-		"                  [ skip_sw | skip_hw ]\n"
-		"                  [ action ACTION-SPEC ] [ classid CLASSID ]\n"
-		"\n"
-		"Where: MATCH-LIST := [ MATCH-LIST ] MATCH\n"
-		"       MATCH      := { indev DEV-NAME |\n"
-		"                       vlan_id VID |\n"
-		"                       vlan_prio PRIORITY |\n"
-		"                       vlan_ethtype [ ipv4 | ipv6 | ETH-TYPE ] |\n"
-		"                       cvlan_id VID |\n"
-		"                       cvlan_prio PRIORITY |\n"
-		"                       cvlan_ethtype [ ipv4 | ipv6 | ETH-TYPE ] |\n"
-		"                       dst_mac MASKED-LLADDR |\n"
-		"                       src_mac MASKED-LLADDR |\n"
[…]

### 2019-11-15 iproute2/tc/f_matchall.c

@@ -1,170 +1,0 @@
-/*
- * f_matchall.c		Match-all Classifier
- *
- *		This program is free software; you can distribute it and/or
- *		modify it under the terms of the GNU General Public License
- *		as published by the Free Software Foundation; either version
- *		2 of the License, or (at your option) any later version.
- *
- * Authors:	Jiri Pirko <jiri@mellanox.com>, Yotam Gigi <yotamg@mellanox.com>
- *
- */
-
-#include <stdio.h>
-#include <stdlib.h>
-#include <unistd.h>
-#include <fcntl.h>
-#include <sys/socket.h>
-#include <netinet/in.h>
-#include <arpa/inet.h>
-#include <string.h>
-#include <linux/if.h>
-
-#include "utils.h"
-#include "tc_util.h"
-
-static void explain(void)
-{
-	fprintf(stderr, "Usage: ... matchall [skip_sw | skip_hw]\n");
-	fprintf(stderr, "                 [ action ACTION_SPEC ] [ classid CLASSID ]\n");
-	fprintf(stderr, "\n");
-	fprintf(stderr, "Where: SELECTOR := SAMPLE SAMPLE ...\n");
-	fprintf(stderr, "       FILTERID := X:Y:Z\n");
-	fprintf(stderr, "       ACTION_SPEC := ... look at individual actions\n");
-	fprintf(stderr, "\nNOTE: CLASSID is parsed as hexadecimal input.\n");
-}
-
-static int matchall_parse_opt(struct filter_util *qu, char *handle,
-			   int argc, char **argv, struct nlmsghdr *n)
-{
-	struct tcmsg *t = NLMSG_DATA(n);
-	struct rtattr *tail;
-	__u32 flags = 0;
-	long h = 0;
-
-	if (handle) {
-		h = strtol(handle, NULL, 0);
-		if (h == LONG_MIN || h == LONG_MAX) {
-			fprintf(stderr, "Illegal handle \"%s\", must be numeric.\n",
-			    handle);
-			return -1;
-		}
-	}
-	t->tcm_handle = h;
-
-	if (argc == 0)
-		return 0;
-
[…]

### 2019-11-15 iproute2/tc/m_action.c

@@ -1,775 +1,0 @@
-/*
- * m_action.c		Action Management
- *
- *		This program is free software; you can distribute it and/or
- *		modify it under the terms of the GNU General Public License
- *		as published by the Free Software Foundation; either version
- *		2 of the License, or (at your option) any later version.
- *
- * Authors:  J Hadi Salim (hadi@cyberus.ca)
- *
- * TODO:
- * - parse to be passed a filedescriptor for logging purposes
- *
- */
-
-#include <stdio.h>
-#include <stdlib.h>
-#include <stdbool.h>
-#include <unistd.h>
-#include <fcntl.h>
-#include <sys/socket.h>
-#include <netinet/in.h>
-#include <arpa/inet.h>
-#include <string.h>
-#include <dlfcn.h>
-
-#include "utils.h"
-#include "tc_common.h"
-#include "tc_util.h"
-
-static struct action_util *action_list;
-#ifdef CONFIG_GACT
-static int gact_ld; /* f*ckin backward compatibility */
-#endif
-static int tab_flush;
-
-static void act_usage(void)
-{
-	/*XXX: In the near future add a action->print_help to improve
-	 * usability
-	 * This would mean new tc will not be backward compatible
-	 * with any action .so from the old days. But if someone really
-	 * does that, they would know how to fix this ..
-	 *
-	 */
-	fprintf(stderr, "usage: tc actions <ACTSPECOP>*\n");
-	fprintf(stderr,
-		"Where: \tACTSPECOP := ACR | GD | FL\n"
-			"\tACR := add | change | replace <ACTSPEC>*\n"
-			"\tGD := get | delete | <ACTISPEC>*\n"
-			"\tFL := ls | list | flush | <ACTNAMESPEC>\n"
-			"\tACTNAMESPEC :=  action <ACTNAME>\n"
-			"\tACTISPEC := <ACTNAMESPEC> <INDEXSPEC>\n"
-			"\tACTSPEC := action <ACTDETAIL> [INDEXSPEC]\n"
-			"\tINDEXSPEC := index <32 bit indexvalue>\n"
[…]

### 2019-11-15 iproute2/tc/m_connmark.c

@@ -1,148 +1,0 @@
-/*
- * m_connmark.c		Connection tracking marking import
- *
- * Copyright (c) 2011 Felix Fietkau <nbd@openwrt.org>
- *
- * This program is free software; you can redistribute it and/or modify it
- * under the terms and conditions of the GNU General Public License,
- * version 2, as published by the Free Software Foundation.
- *
- * This program is distributed in the hope it will be useful, but WITHOUT
- * ANY WARRANTY; without even the implied warranty of MERCHANTABILITY or
- * FITNESS FOR A PARTICULAR PURPOSE.  See the GNU General Public License for
- * more details.
- *
- * You should have received a copy of the GNU General Public License along with
- * this program; if not, see <http://www.gnu.org/licenses>.
- */
-
-#include <stdio.h>
-#include <stdlib.h>
-#include <unistd.h>
-#include <string.h>
-#include "utils.h"
-#include "tc_util.h"
-#include <linux/tc_act/tc_connmark.h>
-
-static void
-explain(void)
-{
-	fprintf(stderr, "Usage: ... connmark [zone ZONE] [CONTROL] [index <INDEX>]\n");
-	fprintf(stderr, "where :\n"
-		"\tZONE is the conntrack zone\n"
-		"\tCONTROL := reclassify | pipe | drop | continue | ok |\n"
-		"\t           goto chain <CHAIN_INDEX>\n");
-}
-
-static void
-usage(void)
-{
-	explain();
-	exit(-1);
-}
-
-static int
-parse_connmark(struct action_util *a, int *argc_p, char ***argv_p, int tca_id,
-	      struct nlmsghdr *n)
-{
-	struct tc_connmark sel = {};
-	char **argv = *argv_p;
-	int argc = *argc_p;
-	int ok = 0;
-	struct rtattr *tail;
-
-	while (argc > 0) {
-		if (matches(*argv, "connmark") == 0) {
-			ok = 1;
-			argc--;
-			argv++;
-		} else if (matches(*argv, "help") == 0) {
-			usage();
-		} else {
-			break;
-		}
-
-	}
-
-	if (!ok) {
-		explain();
-		return -1;
-	}
-
-	if (argc) {
[…]

### 2019-11-15 iproute2/tc/m_ife.c

@@ -1,337 +1,0 @@
-/*
- * m_ife.c	IFE actions module
- *
- *		This program is free software; you can distribute it and/or
- *		modify it under the terms of the GNU General Public License
- *		as published by the Free Software Foundation; either version
- *		2 of the License, or (at your option) any later version.
- *
- * Authors:  J Hadi Salim (jhs@mojatatu.com)
- *
- */
-
-#include <stdio.h>
-#include <stdlib.h>
-#include <unistd.h>
-#include <fcntl.h>
-#include <sys/socket.h>
-#include <netinet/in.h>
-#include <arpa/inet.h>
-#include <string.h>
-#include <linux/netdevice.h>
-
-#include "rt_names.h"
-#include "utils.h"
-#include "tc_util.h"
-#include <linux/tc_act/tc_ife.h>
-
-static void ife_explain(void)
-{
-	fprintf(stderr,
-		"Usage:... ife {decode|encode} [{ALLOW|USE} ATTR] [dst DMAC] [src SMAC] [type TYPE] [CONTROL] [index INDEX]\n");
-	fprintf(stderr,
-		"\tALLOW := Encode direction. Allows encoding specified metadata\n"
-		"\t\t e.g \"allow mark\"\n"
-		"\tUSE := Encode direction. Enforce Static encoding of specified metadata\n"
-		"\t\t e.g \"use mark 0x12\"\n"
-		"\tATTR := mark (32-bit), prio (32-bit), tcindex (16-bit)\n"
-		"\tDMAC := 6 byte Destination MAC address to encode\n"
-		"\tSMAC := optional 6 byte Source MAC address to encode\n"
-		"\tTYPE := optional 16 bit ethertype to encode\n"
-		"\tCONTROL := reclassify|pipe|drop|continue|ok\n"
-		"\tINDEX := optional IFE table index value used\n");
-	fprintf(stderr, "encode is used for sending IFE packets\n");
-	fprintf(stderr, "decode is used for receiving IFE packets\n");
-}
-
-static void ife_usage(void)
-{
-	ife_explain();
[…]

### 2019-11-15 iproute2/tc/m_ipt.c

@@ -1,516 +1,0 @@
-/*
- * m_ipt.c	iptables based targets
- *		utilities mostly ripped from iptables <duh, its the linux way>
- *
- *		This program is free software; you can distribute it and/or
- *		modify it under the terms of the GNU General Public License
- *		as published by the Free Software Foundation; either version
- *		2 of the License, or (at your option) any later version.
- *
- * Authors:  J Hadi Salim (hadi@cyberus.ca)
- */
-
-#include <sys/socket.h>
-#include <netinet/in.h>
-#include <arpa/inet.h>
-#include <iptables.h>
-#include <linux/netfilter.h>
-#include <linux/netfilter_ipv4/ip_tables.h>
-#include "utils.h"
-#include "tc_util.h"
-#include <linux/tc_act/tc_ipt.h>
-#include <stdio.h>
-#include <dlfcn.h>
-#include <getopt.h>
-#include <errno.h>
-#include <string.h>
-#include <netdb.h>
-#include <stdlib.h>
-#include <ctype.h>
-#include <stdarg.h>
-#include <unistd.h>
-#include <fcntl.h>
-#include <sys/wait.h>
-
-static const char *pname = "tc-ipt";
-static const char *tname = "mangle";
-static const char *pversion = "0.1";
-
-static const char *ipthooks[] = {
-	"NF_IP_PRE_ROUTING",
-	"NF_IP_LOCAL_IN",
-	"NF_IP_FORWARD",
-	"NF_IP_LOCAL_OUT",
-	"NF_IP_POST_ROUTING",
-};
-
-static struct option original_opts[] = {
-	{"jump", 1, 0, 'j'},
-	{0, 0, 0, 0}
-};
-
-static struct xtables_target *t_list;
-static struct option *opts = original_opts;
-static unsigned int global_option_offset;
-#define OPTION_OFFSET 256
-
-char *lib_dir;
-
-void
-xtables_register_target(struct xtables_target *me)
-{
-	me->next = t_list;
-	t_list = me;
-
-}
-
-static void exit_tryhelp(int status)
-{
-	fprintf(stderr, "Try `%s -h' or '%s --help' for more information.\n",
-		pname, pname);
-	exit(status);
-}
-
[…]

### 2019-11-15 iproute2/tc/m_police.c

@@ -1,335 +1,0 @@
-/*
- * m_police.c		Parse/print policing module options.
- *
- *		This program is free software; you can u32istribute it and/or
- *		modify it under the terms of the GNU General Public License
- *		as published by the Free Software Foundation; either version
- *		2 of the License, or (at your option) any later version.
- *
- * Authors:	Alexey Kuznetsov, <kuznet@ms2.inr.ac.ru>
- * FIXES:       19990619 - J Hadi Salim (hadi@cyberus.ca)
- *		simple addattr packaging fix.
- *		2002: J Hadi Salim - Add tc action extensions syntax
- *
- */
-
-#include <stdio.h>
-#include <stdlib.h>
-#include <unistd.h>
-#include <fcntl.h>
-#include <sys/socket.h>
-#include <netinet/in.h>
-#include <arpa/inet.h>
-#include <string.h>
-
-#include "utils.h"
-#include "tc_util.h"
-
-static int act_parse_police(struct action_util *a, int *argc_p,
-			    char ***argv_p, int tca_id, struct nlmsghdr *n);
-static int print_police(struct action_util *a, FILE *f, struct rtattr *tb);
-
-struct action_util police_action_util = {
-	.id = "police",
-	.parse_aopt = act_parse_police,
-	.print_aopt = print_police,
-};
-
-static void usage(void)
-{
-	fprintf(stderr, "Usage: ... police rate BPS burst BYTES[/BYTES] [ mtu BYTES[/BYTES] ]\n");
-	fprintf(stderr, "                [ peakrate BPS ] [ avrate BPS ] [ overhead BYTES ]\n");
-	fprintf(stderr, "                [ linklayer TYPE ] [ CONTROL ]\n");
-
-	fprintf(stderr, "Where: CONTROL := conform-exceed <EXCEEDACT>[/NOTEXCEEDACT]\n");
-	fprintf(stderr, "                  Define how to handle packets which exceed (<EXCEEDACT>)\n");
-	fprintf(stderr, "                  or conform (<NOTEXCEEDACT>) the configured bandwidth limit.\n");
[…]

### 2019-11-15 iproute2/tc/m_tunnel_key.c

@@ -1,556 +1,0 @@
-/*
- * m_tunnel_key.c	ip tunnel manipulation module
- *
- *              This program is free software; you can redistribute it and/or
- *              modify it under the terms of the GNU General Public License
- *              as published by the Free Software Foundation; either version
- *              2 of the License, or (at your option) any later version.
- *
- * Authors:     Amir Vadai <amir@vadai.me>
- */
-
-#include <stdio.h>
-#include <stdlib.h>
-#include <unistd.h>
-#include <string.h>
-#include <linux/if_ether.h>
-#include "utils.h"
-#include "rt_names.h"
-#include "tc_util.h"
-#include <linux/tc_act/tc_tunnel_key.h>
-
-static void explain(void)
-{
-	fprintf(stderr, "Usage: tunnel_key unset\n");
-	fprintf(stderr, "       tunnel_key set <TUNNEL_KEY>\n");
-	fprintf(stderr,
-		"Where TUNNEL_KEY is a combination of:\n"
-		"id <TUNNELID>\n"
-		"src_ip <IP> (mandatory)\n"
-		"dst_ip <IP> (mandatory)\n"
-		"dst_port <UDP_PORT>\n"
-		"geneve_opts <OPTIONS>\n"
-		"csum | nocsum (default is \"csum\")\n");
-}
-
-static void usage(void)
-{
-	explain();
-	exit(-1);
-}
-
-static int tunnel_key_parse_ip_addr(const char *str, int addr4_type,
-				    int addr6_type, struct nlmsghdr *n)
-{
-	inet_prefix addr;
-	int ret;
-
-	ret = get_addr(&addr, str, AF_UNSPEC);
-	if (ret)
-		return ret;
-
-	addattr_l(n, MAX_MSG, addr.family == AF_INET ? addr4_type : addr6_type,
-		  addr.data, addr.bytelen);
-
-	return 0;
-}
-
-static int tunnel_key_parse_key_id(const char *str, int type,
-				   struct nlmsghdr *n)
-{
-	__be32 key_id;
-	int ret;
-
-	ret = get_be32(&key_id, str, 10);
-	if (!ret)
[…]

### 2019-11-15 iproute2/tc/m_xt_old.c

@@ -1,439 +1,0 @@
-/*
- * m_xt.c	xtables based targets
- *		utilities mostly ripped from iptables <duh, its the linux way>
- *
- *		This program is free software; you can distribute it and/or
- *		modify it under the terms of the GNU General Public License
- *		as published by the Free Software Foundation; either version
- *		2 of the License, or (at your option) any later version.
- *
- * Authors:  J Hadi Salim (hadi@cyberus.ca)
- */
-
-/*XXX: in the future (xtables 1.4.3?) get rid of everything tagged
- * as TC_CONFIG_XT_H */
-
-#include <sys/socket.h>
-#include <netinet/in.h>
-#include <arpa/inet.h>
-#include <net/if.h>
-#include <linux/netfilter.h>
-#include <linux/netfilter_ipv4/ip_tables.h>
-#include <xtables.h>
-#include "utils.h"
-#include "tc_util.h"
-#include <linux/tc_act/tc_ipt.h>
-#include <stdio.h>
-#include <getopt.h>
-#include <errno.h>
-#include <string.h>
-#include <netdb.h>
-#include <stdlib.h>
-#include <ctype.h>
-#include <stdarg.h>
-#include <limits.h>
-#include <unistd.h>
-#include <fcntl.h>
-#include <sys/wait.h>
-#ifdef TC_CONFIG_XT_H
-#include "xt-internal.h"
-#endif
-
-#ifndef ALIGN
-#define ALIGN(x, a)		__ALIGN_MASK(x, (typeof(x))(a)-1)
-#define __ALIGN_MASK(x, mask)	(((x)+(mask))&~(mask))
-#endif
-
-static const char *pname = "tc-ipt";
-static const char *tname = "mangle";
-static const char *pversion = "0.2";
-
-static const char *ipthooks[] = {
-	"NF_IP_PRE_ROUTING",
-	"NF_IP_LOCAL_IN",
-	"NF_IP_FORWARD",
-	"NF_IP_LOCAL_OUT",
-	"NF_IP_POST_ROUTING",
-};
-
-static struct option original_opts[] = {
-	{"jump", 1, 0, 'j'},
-	{0, 0, 0, 0}
-};
-
-static struct option *opts = original_opts;
-static unsigned int global_option_offset;
-char *lib_dir;
[…]

### 2019-11-15 iproute2/tc/p_ip6.c

@@ -1,89 +1,0 @@
-/*
- * p_ip6.c		packet editor: IPV6 header
- *
- *		This program is free software; you can distribute it and/or
- *		modify it under the terms of the GNU General Public License
- *		as published by the Free Software Foundation; either version
- *		2 of the License, or (at your option) any later version.
- *
- * Authors:  Amir Vadai <amir@vadai.me>
- *
- */
-
-#include <stdio.h>
-#include <stdlib.h>
-#include <unistd.h>
-#include <fcntl.h>
-#include <sys/socket.h>
-#include <netinet/in.h>
-#include <arpa/inet.h>
-#include <string.h>
-#include "utils.h"
-#include "tc_util.h"
-#include "m_pedit.h"
-
-static int
-parse_ip6(int *argc_p, char ***argv_p,
-	  struct m_pedit_sel *sel, struct m_pedit_key *tkey)
-{
-	int res = -1;
-	int argc = *argc_p;
-	char **argv = *argv_p;
-
-	if (argc < 2)
-		return -1;
-
-	if (!sel->extended)
-		return -1;
-
-	tkey->htype = TCA_PEDIT_KEY_EX_HDR_TYPE_IP6;
-
-	if (strcmp(*argv, "src") == 0) {
-		NEXT_ARG();
-		tkey->off = 8;
-		res = parse_cmd(&argc, &argv, 16, TIPV6, RU32, sel, tkey);
-		goto done;
-	}
-	if (strcmp(*argv, "dst") == 0) {
-		NEXT_ARG();
-		tkey->off = 24;
-		res = parse_cmd(&argc, &argv, 16, TIPV6, RU32, sel, tkey);
-		goto done;
-	}
-	if (strcmp(*argv, "flow_lbl") == 0) {
-		NEXT_ARG();
-		tkey->off = 0;
-		res = parse_cmd(&argc, &argv, 4, TU32, 0x0007ffff, sel, tkey);
-		goto done;
-	}
-	if (strcmp(*argv, "payload_len") == 0) {
-		NEXT_ARG();
-		tkey->off = 4;
-		res = parse_cmd(&argc, &argv, 2, TU32, RU16, sel, tkey);
-		goto done;
-	}
-	if (strcmp(*argv, "nexthdr") == 0) {
-		NEXT_ARG();
-		tkey->off = 6;
[…]

### 2019-11-15 iproute2/tc/q_cake.c

@@ -1,829 +1,0 @@
-// SPDX-License-Identifier: GPL-2.0 OR BSD-3-Clause
-
-/*
- * Common Applications Kept Enhanced  --  CAKE
- *
- *  Copyright (C) 2014-2018 Jonathan Morton <chromatix99@gmail.com>
- *  Copyright (C) 2017-2018 Toke Høiland-Jørgensen <toke@toke.dk>
- */
-
-#include <stddef.h>
-#include <stdio.h>
-#include <stdlib.h>
-#include <unistd.h>
-#include <syslog.h>
-#include <fcntl.h>
-#include <sys/socket.h>
-#include <netinet/in.h>
-#include <arpa/inet.h>
-#include <string.h>
-#include <inttypes.h>
-
-#include "utils.h"
-#include "tc_util.h"
-
-struct cake_preset {
-	char *name;
-	unsigned int target;
-	unsigned int interval;
-};
-
-static struct cake_preset presets[] = {
-	{"datacentre",		5,		100},
-	{"lan",			50,		1000},
-	{"metro",		500,		10000},
-	{"regional",		1500,		30000},
-	{"internet",		5000,		100000},
-	{"oceanic",		15000,		300000},
-	{"satellite",		50000,		1000000},
-	{"interplanetary",	50000000,	1000000000},
-};
-
-static const char * diffserv_names[CAKE_DIFFSERV_MAX] = {
-	[CAKE_DIFFSERV_DIFFSERV3] = "diffserv3",
-	[CAKE_DIFFSERV_DIFFSERV4] = "diffserv4",
-	[CAKE_DIFFSERV_DIFFSERV8] = "diffserv8",
-	[CAKE_DIFFSERV_BESTEFFORT] = "besteffort",
-	[CAKE_DIFFSERV_PRECEDENCE] = "precedence",
-};
-
-static const char * flowmode_names[CAKE_FLOW_MAX] = {
-	[CAKE_FLOW_NONE] = "flowblind",
-	[CAKE_FLOW_SRC_IP] = "srchost",
-	[CAKE_FLOW_DST_IP] = "dsthost",
-	[CAKE_FLOW_HOSTS] = "hosts",
-	[CAKE_FLOW_FLOWS] = "flows",
-	[CAKE_FLOW_DUAL_SRC] = "dual-srchost",
-	[CAKE_FLOW_DUAL_DST] = "dual-dsthost",
[…]

### 2019-11-15 iproute2/tc/q_choke.c

@@ -1,227 +1,0 @@
-/*
- * q_choke.c		CHOKE.
- *
- *		This program is free software; you can redistribute it and/or
- *		modify it under the terms of the GNU General Public License
- *		as published by the Free Software Foundation; either version
- *		2 of the License, or (at your option) any later version.
- *
- * Authors:	Stephen Hemminger <shemminger@vyatta.com>
- */
-
-#include <stdio.h>
-#include <stdlib.h>
-#include <unistd.h>
-#include <fcntl.h>
-#include <sys/socket.h>
-#include <netinet/in.h>
-#include <arpa/inet.h>
-#include <string.h>
-#include <math.h>
-
-#include "utils.h"
-#include "tc_util.h"
-
-#include "tc_red.h"
-
-static void explain(void)
-{
-	fprintf(stderr, "Usage: ... choke limit PACKETS bandwidth KBPS [ecn]\n");
-	fprintf(stderr, "                 [ min PACKETS ] [ max PACKETS ] [ burst PACKETS ]\n");
-}
-
-static int choke_parse_opt(struct qdisc_util *qu, int argc, char **argv,
-			   struct nlmsghdr *n, const char *dev)
-{
-	struct tc_red_qopt opt = {};
-	unsigned int burst = 0;
-	unsigned int avpkt = 1000;
-	double probability = 0.02;
-	unsigned int rate = 0;
-	int ecn_ok = 0;
-	int wlog;
-	__u8 sbuf[256];
-	__u32 max_P;
-	struct rtattr *tail;
-
-	while (argc > 0) {
-		if (strcmp(*argv, "limit") == 0) {
-			NEXT_ARG();
-			if (get_unsigned(&opt.limit, *argv, 0)) {
-				fprintf(stderr, "Illegal \"limit\"\n");
-				return -1;
-			}
-		} else if (strcmp(*argv, "bandwidth") == 0) {
-			NEXT_ARG();
-			if (strchr(*argv, '%')) {
-				if (get_percent_rate(&rate, *argv, dev)) {
-					fprintf(stderr, "Illegal \"bandwidth\"\n");
-					return -1;
-				}
-			} else if (get_rate(&rate, *argv)) {
[…]

### 2019-11-15 iproute2/tc/q_fq.c

@@ -1,385 +1,0 @@
-/*
- * Fair Queue
- *
- *  Copyright (C) 2013-2015 Eric Dumazet <edumazet@google.com>
- *
- * Redistribution and use in source and binary forms, with or without
- * modification, are permitted provided that the following conditions
- * are met:
- * 1. Redistributions of source code must retain the above copyright
- *    notice, this list of conditions, and the following disclaimer,
- *    without modification.
- * 2. Redistributions in binary form must reproduce the above copyright
- *    notice, this list of conditions and the following disclaimer in the
- *    documentation and/or other materials provided with the distribution.
- * 3. The names of the authors may not be used to endorse or promote products
- *    derived from this software without specific prior written permission.
- *
- * Alternatively, provided that this notice is retained in full, this
- * software may be distributed under the terms of the GNU General
- * Public License ("GPL") version 2, in which case the provisions of the
- * GPL apply INSTEAD OF those given above.
- *
- * THIS SOFTWARE IS PROVIDED BY THE COPYRIGHT HOLDERS AND CONTRIBUTORS
- * "AS IS" AND ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT
- * LIMITED TO, THE IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS FOR
- * A PARTICULAR PURPOSE ARE DISCLAIMED. IN NO EVENT SHALL THE COPYRIGHT
- * OWNER OR CONTRIBUTORS BE LIABLE FOR ANY DIRECT, INDIRECT, INCIDENTAL,
- * SPECIAL, EXEMPLARY, OR CONSEQUENTIAL DAMAGES (INCLUDING, BUT NOT
- * LIMITED TO, PROCUREMENT OF SUBSTITUTE GOODS OR SERVICES; LOSS OF USE,
[…]

### 2019-11-15 iproute2/tc/q_gred.c

@@ -1,529 +1,0 @@
-/*
- * q_gred.c		GRED.
- *
- *		This program is free software; you can redistribute it and/or
- *		modify it under the terms of the GNU General Public License
- *		as published by the Free Software Foundation; either version
- *		2 of the License, or (at your option) any later version.
- *
- * Authors:    J Hadi Salim(hadi@nortelnetworks.com)
- *             code ruthlessly ripped from
- *	       Alexey Kuznetsov, <kuznet@ms2.inr.ac.ru>
- *
- */
-
-#include <stdio.h>
-#include <stdlib.h>
-#include <unistd.h>
-#include <fcntl.h>
-#include <sys/socket.h>
-#include <netinet/in.h>
-#include <arpa/inet.h>
-#include <string.h>
-#include <math.h>
-
-#include "utils.h"
-#include "tc_util.h"
-
-#include "tc_red.h"
-
-
-#if 0
-#define DPRINTF(format, args...) fprintf(stderr, format, ##args)
-#else
-#define DPRINTF(format, args...)
-#endif
-
-static void explain(void)
-{
-	fprintf(stderr, "Usage: tc qdisc { add | replace | change } ... gred setup vqs NUMBER\n");
-	fprintf(stderr, "           default DEFAULT_VQ [ grio ] [ limit BYTES ] [ecn] [harddrop]\n");
-	fprintf(stderr, "       tc qdisc change ... gred vq VQ [ prio VALUE ] limit BYTES\n");
-	fprintf(stderr, "           min BYTES max BYTES avpkt BYTES [ burst PACKETS ]\n");
-	fprintf(stderr, "           [ probability PROBABILITY ] [ bandwidth KBPS ] [ecn] [harddrop]\n");
-}
-
-static int init_gred(struct qdisc_util *qu, int argc, char **argv,
-		     struct nlmsghdr *n)
-{
-
-	struct rtattr *tail;
-	struct tc_gred_sopt opt = { 0 };
-	__u32 limit = 0;
-
-	opt.def_DP = MAX_DPs;
-
-	while (argc > 0) {
[…]

### 2019-11-15 iproute2/tc/q_netem.c

@@ -1,789 +1,0 @@
-/*
- * q_netem.c		NETEM.
- *
- *		This program is free software; you can redistribute it and/or
- *		modify it under the terms of the GNU General Public License
- *		as published by the Free Software Foundation; either version
- *		2 of the License, or (at your option) any later version.
- *
- * Authors:	Stephen Hemminger <shemminger@linux-foundation.org>
- *
- */
-
-#include <stdio.h>
-#include <stdlib.h>
-#include <math.h>
-#include <ctype.h>
-#include <unistd.h>
-#include <fcntl.h>
-#include <stdint.h>
-#include <sys/socket.h>
-#include <netinet/in.h>
-#include <arpa/inet.h>
-#include <string.h>
-#include <errno.h>
-
-#include "utils.h"
-#include "tc_util.h"
-#include "tc_common.h"
-
-static void explain(void)
-{
-	fprintf(stderr,
-"Usage: ... netem [ limit PACKETS ]\n" \
-"                 [ delay TIME [ JITTER [CORRELATION]]]\n" \
-"                 [ distribution {uniform|normal|pareto|paretonormal} ]\n" \
-"                 [ corrupt PERCENT [CORRELATION]]\n" \
-"                 [ duplicate PERCENT [CORRELATION]]\n" \
-"                 [ loss random PERCENT [CORRELATION]]\n" \
-"                 [ loss state P13 [P31 [P32 [P23 P14]]]\n" \
-"                 [ loss gemodel PERCENT [R [1-H [1-K]]]\n" \
-"                 [ ecn ]\n" \
-"                 [ reorder PRECENT [CORRELATION] [ gap DISTANCE ]]\n" \
-"                 [ rate RATE [PACKETOVERHEAD] [CELLSIZE] [CELLOVERHEAD]]\n" \
-"                 [ slot MIN_DELAY [MAX_DELAY] [packets MAX_PACKETS]" \
-" [bytes MAX_BYTES]]\n" \
-"                 [ slot distribution" \
-" {uniform|normal|pareto|paretonormal|custom} DELAY JITTER" \
-" [packets MAX_PACKETS] [bytes MAX_BYTES]]\n");
-}
-
-static void explain1(const char *arg)
-{
-	fprintf(stderr, "Illegal \"%s\"\n", arg);
-}
-
[…]

### 2019-11-15 iproute2/tc/q_pie.c

@@ -1,215 +1,0 @@
-/* Copyright (C) 2013 Cisco Systems, Inc, 2013.
- *
- * This program is free software; you can redistribute it and/or
- * modify it under the terms of the GNU General Public License
- * as published by the Free Software Foundation; either version 2
- * of the License.
- *
- * This program is distributed in the hope that it will be useful,
- * but WITHOUT ANY WARRANTY; without even the implied warranty of
- * MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
- * GNU General Public License for more details.
- *
- * Author: Vijay Subramanian <vijaynsu@cisco.com>
- * Author: Mythili Prabhu <mysuryan@cisco.com>
- *
- */
-
-#include <stdio.h>
-#include <stdlib.h>
-#include <unistd.h>
-#include <fcntl.h>
-#include <sys/socket.h>
-#include <netinet/in.h>
-#include <arpa/inet.h>
-#include <string.h>
-#include <math.h>
-
-#include "utils.h"
-#include "tc_util.h"
-
-static void explain(void)
-{
-	fprintf(stderr, "Usage: ... pie [ limit PACKETS ][ target TIME us]\n");
-	fprintf(stderr, "              [ tupdate TIME us][ alpha ALPHA ]");
-	fprintf(stderr, "[beta BETA ][bytemode | nobytemode][ecn | noecn ]\n");
-}
-
-#define ALPHA_MAX 32
-#define BETA_MAX 32
-
-static int pie_parse_opt(struct qdisc_util *qu, int argc, char **argv,
-			 struct nlmsghdr *n, const char *dev)
-{
-	unsigned int limit   = 0;
-	unsigned int target  = 0;
-	unsigned int tupdate = 0;
-	unsigned int alpha   = 0;
-	unsigned int beta    = 0;
-	int ecn = -1;
-	int bytemode = -1;
-	struct rtattr *tail;
-
-	while (argc > 0) {
-		if (strcmp(*argv, "limit") == 0) {
-			NEXT_ARG();
[…]

### 2019-11-15 iproute2/tc/q_red.c

@@ -1,233 +1,0 @@
-/*
- * q_red.c		RED.
- *
- *		This program is free software; you can redistribute it and/or
- *		modify it under the terms of the GNU General Public License
- *		as published by the Free Software Foundation; either version
- *		2 of the License, or (at your option) any later version.
- *
- * Authors:	Alexey Kuznetsov, <kuznet@ms2.inr.ac.ru>
- *
- */
-
-#include <stdio.h>
-#include <stdlib.h>
-#include <unistd.h>
-#include <fcntl.h>
-#include <sys/socket.h>
-#include <netinet/in.h>
-#include <arpa/inet.h>
-#include <string.h>
-#include <math.h>
-
-#include "utils.h"
-#include "tc_util.h"
-
-#include "tc_red.h"
-
-static void explain(void)
-{
-	fprintf(stderr, "Usage: ... red limit BYTES [min BYTES] [max BYTES] avpkt BYTES [burst PACKETS]\n");
-	fprintf(stderr, "               [adaptive] [probability PROBABILITY] [bandwidth KBPS]\n");
-	fprintf(stderr, "               [ecn] [harddrop]\n");
-}
-
-static int red_parse_opt(struct qdisc_util *qu, int argc, char **argv,
-			 struct nlmsghdr *n, const char *dev)
-{
-	struct tc_red_qopt opt = {};
-	unsigned int burst = 0;
-	unsigned int avpkt = 0;
-	double probability = 0.02;
-	unsigned int rate = 0;
-	int parm;
-	__u8 sbuf[256];
-	__u32 max_P;
-	struct rtattr *tail;
-
-	while (argc > 0) {
-		if (strcmp(*argv, "limit") == 0) {
-			NEXT_ARG();
-			if (get_size(&opt.limit, *argv)) {
-				fprintf(stderr, "Illegal \"limit\"\n");
-				return -1;
-			}
-		} else if (strcmp(*argv, "min") == 0) {
-			NEXT_ARG();
-			if (get_size(&opt.qth_min, *argv)) {
-				fprintf(stderr, "Illegal \"min\"\n");
-				return -1;
-			}
-		} else if (strcmp(*argv, "max") == 0) {
-			NEXT_ARG();
[…]

### 2019-11-15 iproute2/tc/q_sfq.c

@@ -1,274 +1,0 @@
-/*
- * q_sfq.c		SFQ.
- *
- *		This program is free software; you can redistribute it and/or
- *		modify it under the terms of the GNU General Public License
- *		as published by the Free Software Foundation; either version
- *		2 of the License, or (at your option) any later version.
- *
- * Authors:	Alexey Kuznetsov, <kuznet@ms2.inr.ac.ru>
- *
- */
-
-#include <stdio.h>
-#include <stdlib.h>
-#include <unistd.h>
-#include <fcntl.h>
-#include <sys/socket.h>
-#include <netinet/in.h>
-#include <arpa/inet.h>
-#include <string.h>
-#include <math.h>
-
-#include "utils.h"
-#include "tc_util.h"
-#include "tc_red.h"
-
-static void explain(void)
-{
-	fprintf(stderr, "Usage: ... sfq [ limit NUMBER ] [ perturb SECS ] [ quantum BYTES ]\n");
-	fprintf(stderr, "               [ divisor NUMBER ] [ flows NUMBER] [ depth NUMBER ]\n");
-	fprintf(stderr, "               [ headdrop ]\n");
-	fprintf(stderr, "               [ redflowlimit BYTES ] [ min BYTES ] [ max BYTES ]\n");
-	fprintf(stderr, "               [ avpkt BYTES ] [ burst PACKETS ] [ probability P ]\n");
-	fprintf(stderr, "               [ ecn ] [ harddrop ]\n");
-}
-
-static int sfq_parse_opt(struct qdisc_util *qu, int argc, char **argv, struct nlmsghdr *n, const char *dev)
-{
-	int ok = 0, red = 0;
-	struct tc_sfq_qopt_v1 opt = {};
-	unsigned int burst = 0;
-	int wlog;
-	unsigned int avpkt = 1000;
-	double probability = 0.02;
-
-	while (argc > 0) {
-		if (strcmp(*argv, "quantum") == 0) {
-			NEXT_ARG();
-			if (get_size(&opt.v0.quantum, *argv)) {
-				fprintf(stderr, "Illegal \"limit\"\n");
-				return -1;
-			}
-			ok++;
-		} else if (strcmp(*argv, "perturb") == 0) {
-			NEXT_ARG();
[…]

### 2019-11-15 iproute2/tc/q_taprio.c

@@ -1,466 +1,0 @@
-/*
- * q_taprio.c	Time Aware Priority Scheduler
- *
- *		This program is free software; you can redistribute it and/or
- *		modify it under the terms of the GNU General Public License
- *		as published by the Free Software Foundation; either version
- *		2 of the License, or (at your option) any later version.
- *
- * Authors:	Vinicius Costa Gomes <vinicius.gomes@intel.com>
- * 		Jesus Sanchez-Palencia <jesus.sanchez-palencia@intel.com>
- */
-
-#include <stdio.h>
-#include <stdlib.h>
-#include <unistd.h>
-#include <syslog.h>
-#include <fcntl.h>
-#include <inttypes.h>
-#include <sys/socket.h>
-#include <netinet/in.h>
-#include <arpa/inet.h>
-#include <string.h>
-
-#include "utils.h"
-#include "tc_util.h"
-#include "list.h"
-
-struct sched_entry {
-	struct list_head list;
-	uint32_t index;
-	uint32_t interval;
-	uint32_t gatemask;
-	uint8_t cmd;
-};
-
-#define CLOCKID_INVALID (-1)
-static const struct static_clockid {
-	const char *name;
-	clockid_t clockid;
-} clockids_sysv[] = {
-	{ "REALTIME", CLOCK_REALTIME },
-	{ "TAI", CLOCK_TAI },
-	{ "BOOTTIME", CLOCK_BOOTTIME },
-	{ "MONOTONIC", CLOCK_MONOTONIC },
-	{ NULL }
-};
-
-static void explain(void)
-{
-	fprintf(stderr, "Usage: ... taprio clockid CLOCKID\n");
-	fprintf(stderr, "                  [num_tc NUMBER] [map P0 P1 ...] ");
-	fprintf(stderr, "                  [queues COUNT@OFFSET COUNT@OFFSET COUNT@OFFSET ...] ");
-	fprintf(stderr, "                  [ [sched-entry index cmd gate-mask interval] ... ] ");
-	fprintf(stderr, "                  [base-time time] ");
[…]

### 2019-11-15 iproute2/tc/tc_cbq.c

@@ -1,58 +1,0 @@
-/*
- * tc_cbq.c		CBQ maintanance routines.
- *
- *		This program is free software; you can redistribute it and/or
- *		modify it under the terms of the GNU General Public License
- *		as published by the Free Software Foundation; either version
- *		2 of the License, or (at your option) any later version.
- *
- * Authors:	Alexey Kuznetsov, <kuznet@ms2.inr.ac.ru>
- *
- */
-
-#include <stdio.h>
-#include <stdlib.h>
-#include <unistd.h>
-#include <fcntl.h>
-#include <math.h>
-#include <sys/socket.h>
-#include <netinet/in.h>
-#include <arpa/inet.h>
-#include <string.h>
-
-#include "utils.h"
-#include "tc_core.h"
-#include "tc_cbq.h"
-
-unsigned int tc_cbq_calc_maxidle(unsigned int bndw, unsigned int rate, unsigned int avpkt,
-			     int ewma_log, unsigned int maxburst)
-{
-	double maxidle;
-	double g = 1.0 - 1.0/(1<<ewma_log);
-	double xmt = (double)avpkt/bndw;
-
-	maxidle = xmt*(1-g);
-	if (bndw != rate && maxburst) {
-		double vxmt = (double)avpkt/rate - xmt;
-
-		vxmt *= (pow(g, -(double)maxburst) - 1);
-		if (vxmt > maxidle)
-			maxidle = vxmt;
-	}
-	return tc_core_time2tick(maxidle*(1<<ewma_log)*TIME_UNITS_PER_SEC);
-}
-
-unsigned int tc_cbq_calc_offtime(unsigned int bndw, unsigned int rate, unsigned int avpkt,
-			     int ewma_log, unsigned int minburst)
-{
-	double g = 1.0 - 1.0/(1<<ewma_log);
-	double offtime = (double)avpkt/rate - (double)avpkt/bndw;
-
-	if (minburst == 0)
-		return 0;
-	if (minburst == 1)
-		offtime *= pow(g, -(double)minburst) - 1;
-	else
-		offtime *= 1 + (pow(g, -(double)(minburst-1)) - 1)/(1-g);
-	return tc_core_time2tick(offtime*TIME_UNITS_PER_SEC);
-}

### 2019-11-15 iproute2/tc/tc_common.h

@@ -1,30 +1,0 @@
-/* SPDX-License-Identifier: GPL-2.0 */
-
-#define TCA_BUF_MAX	(64*1024)
-#define MSG_IOV_MAX	128
-
-extern struct rtnl_handle rth;
-
-int do_qdisc(int argc, char **argv);
-int do_class(int argc, char **argv);
-int do_filter(int argc, char **argv, void *buf, size_t buflen);
-int do_chain(int argc, char **argv, void *buf, size_t buflen);
-int do_action(int argc, char **argv, void *buf, size_t buflen);
-int do_tcmonitor(int argc, char **argv);
-int do_exec(int argc, char **argv);
-
-int print_action(struct nlmsghdr *n, void *arg);
-int print_filter(struct nlmsghdr *n, void *arg);
-int print_qdisc(struct nlmsghdr *n, void *arg);
-int print_class(struct nlmsghdr *n, void *arg);
-void print_size_table(FILE *fp, const char *prefix, struct rtattr *rta);
-
-struct tc_estimator;
-int parse_estimator(int *p_argc, char ***p_argv, struct tc_estimator *est);
-
-struct tc_sizespec;
-int parse_size_table(int *p_argc, char ***p_argv, struct tc_sizespec *s);
-int check_size_table_opts(struct tc_sizespec *s);
-
-extern int show_graph;
-extern bool use_names;

### 2019-11-15 iproute2/tc/tc_core.c

@@ -1,229 +1,0 @@
-/*
- * tc_core.c		TC core library.
- *
- *		This program is free software; you can redistribute it and/or
- *		modify it under the terms of the GNU General Public License
- *		as published by the Free Software Foundation; either version
- *		2 of the License, or (at your option) any later version.
- *
- * Authors:	Alexey Kuznetsov, <kuznet@ms2.inr.ac.ru>
- *
- */
-
-#include <stdio.h>
-#include <stdlib.h>
-#include <stdint.h>
-#include <unistd.h>
-#include <fcntl.h>
-#include <math.h>
-#include <sys/socket.h>
-#include <netinet/in.h>
-#include <arpa/inet.h>
-#include <string.h>
-
-#include "utils.h"
-#include "tc_core.h"
-#include <linux/atm.h>
-
-static double tick_in_usec = 1;
-static double clock_factor = 1;
-
-int tc_core_time2big(unsigned int time)
-{
-	__u64 t = time;
-
-	t *= tick_in_usec;
-	return (t >> 32) != 0;
-}
-
-
-unsigned int tc_core_time2tick(unsigned int time)
-{
-	return time*tick_in_usec;
-}
-
-unsigned int tc_core_tick2time(unsigned int tick)
-{
-	return tick/tick_in_usec;
-}
-
-unsigned int tc_core_time2ktime(unsigned int time)
-{
-	return time * clock_factor;
-}
-
-unsigned int tc_core_ktime2time(unsigned int ktime)
-{
-	return ktime / clock_factor;
-}
-
-unsigned int tc_calc_xmittime(__u64 rate, unsigned int size)
-{
-	return tc_core_time2tick(TIME_UNITS_PER_SEC*((double)size/(double)rate));
-}
-
-unsigned int tc_calc_xmitsize(__u64 rate, unsigned int ticks)
-{
-	return ((double)rate*tc_core_tick2time(ticks))/TIME_UNITS_PER_SEC;
-}
-
-/*
- * The align to ATM cells is used for determining the (ATM) SAR
- * alignment overhead at the ATM layer. (SAR = Segmentation And
- * Reassembly).  This is for example needed when scheduling packet on
- * an ADSL connection.  Note that the extra ATM-AAL overhead is _not_
[…]

### 2019-11-15 iproute2/tc/tc_filter.c

@@ -1,808 +1,0 @@
-/*
- * tc_filter.c		"tc filter".
- *
- *		This program is free software; you can redistribute it and/or
- *		modify it under the terms of the GNU General Public License
- *		as published by the Free Software Foundation; either version
- *		2 of the License, or (at your option) any later version.
- *
- * Authors:	Alexey Kuznetsov, <kuznet@ms2.inr.ac.ru>
- *
- */
-
-#include <stdio.h>
-#include <stdlib.h>
-#include <unistd.h>
-#include <fcntl.h>
-#include <sys/socket.h>
-#include <netinet/in.h>
-#include <arpa/inet.h>
-#include <string.h>
-#include <linux/if_ether.h>
-
-#include "rt_names.h"
-#include "utils.h"
-#include "tc_util.h"
-#include "tc_common.h"
-
-static void usage(void)
-{
-	fprintf(stderr,
-		"Usage: tc filter [ add | del | change | replace | show ] [ dev STRING ]\n"
-		"       tc filter [ add | del | change | replace | show ] [ block BLOCK_INDEX ]\n"
-		"       tc filter get dev STRING parent CLASSID protocol PROTO handle FILTERID pref PRIO FILTER_TYPE\n"
-		"       tc filter get block BLOCK_INDEX protocol PROTO handle FILTERID pref PRIO FILTER_TYPE\n"
-		"       [ pref PRIO ] protocol PROTO [ chain CHAIN_INDEX ]\n"
-		"       [ estimator INTERVAL TIME_CONSTANT ]\n"
-		"       [ root | ingress | egress | parent CLASSID ]\n"
-		"       [ handle FILTERID ] [ [ FILTER_TYPE ] [ help | OPTIONS ] ]\n"
-		"\n"
-		"       tc filter show [ dev STRING ] [ root | ingress | egress | parent CLASSID ]\n"
-		"       tc filter show [ block BLOCK_INDEX ]\n"
-		"Where:\n"
-		"FILTER_TYPE := { rsvp | u32 | bpf | fw | route | etc. }\n"
-		"FILTERID := ... format depends on classifier, see there\n"
[…]

### 2019-11-15 iproute2/tc/tc_qdisc.c

@@ -1,503 +1,0 @@
-/*
- * tc_qdisc.c		"tc qdisc".
- *
- *		This program is free software; you can redistribute it and/or
- *		modify it under the terms of the GNU General Public License
- *		as published by the Free Software Foundation; either version
- *		2 of the License, or (at your option) any later version.
- *
- * Authors:	Alexey Kuznetsov, <kuznet@ms2.inr.ac.ru>
- *		J Hadi Salim: Extension to ingress
- */
-
-#include <stdio.h>
-#include <stdlib.h>
-#include <unistd.h>
-#include <fcntl.h>
-#include <sys/socket.h>
-#include <netinet/in.h>
-#include <arpa/inet.h>
-#include <string.h>
-#include <math.h>
-#include <malloc.h>
-
-#include "utils.h"
-#include "tc_util.h"
-#include "tc_common.h"
-
-static int usage(void)
-{
-	fprintf(stderr, "Usage: tc qdisc [ add | del | replace | change | show ] dev STRING\n");
-	fprintf(stderr, "       [ handle QHANDLE ] [ root | ingress | clsact | parent CLASSID ]\n");
-	fprintf(stderr, "       [ estimator INTERVAL TIME_CONSTANT ]\n");
-	fprintf(stderr, "       [ stab [ help | STAB_OPTIONS] ]\n");
-	fprintf(stderr, "       [ ingress_block BLOCK_INDEX ] [ egress_block BLOCK_INDEX ]\n");
-	fprintf(stderr, "       [ [ QDISC_KIND ] [ help | OPTIONS ] ]\n");
-	fprintf(stderr, "\n");
-	fprintf(stderr, "       tc qdisc show [ dev STRING ] [ ingress | clsact ] [ invisible ]\n");
-	fprintf(stderr, "Where:\n");
-	fprintf(stderr, "QDISC_KIND := { [p|b]fifo | tbf | prio | cbq | red | etc. }\n");
-	fprintf(stderr, "OPTIONS := ... try tc qdisc add <desired QDISC_KIND> help\n");
-	fprintf(stderr, "STAB_OPTIONS := ... try tc qdisc add stab help\n");
-	return -1;
-}
-
[…]

### 2019-11-15 iproute2/tc/tc_util.c

@@ -1,899 +1,0 @@
-/*
- * tc_util.c		Misc TC utility functions.
- *
- *		This program is free software; you can redistribute it and/or
- *		modify it under the terms of the GNU General Public License
- *		as published by the Free Software Foundation; either version
- *		2 of the License, or (at your option) any later version.
- *
- * Authors:	Alexey Kuznetsov, <kuznet@ms2.inr.ac.ru>
- *
- */
-
-#include <stdio.h>
-#include <stdlib.h>
-#include <unistd.h>
-#include <fcntl.h>
-#include <sys/socket.h>
-#include <sys/param.h>
-#include <netinet/in.h>
-#include <arpa/inet.h>
-#include <string.h>
-#include <math.h>
-#include <errno.h>
-
-#include "utils.h"
-#include "names.h"
-#include "tc_util.h"
-#include "tc_common.h"
-
-#ifndef LIBDIR
-#define LIBDIR "/usr/lib"
-#endif
-
-static struct db_names *cls_names;
-
-#define NAMES_DB "/etc/iproute2/tc_cls"
-
-int cls_names_init(char *path)
-{
-	int ret;
-
-	cls_names = db_names_alloc();
-	if (!cls_names)
-		return -1;
-
-	ret = db_names_load(cls_names, path ?: NAMES_DB);
-	if (ret == -ENOENT && path) {
-		fprintf(stderr, "Can't open class names file: %s\n", path);
-		return -1;
-	}
-	if (ret) {
-		db_names_free(cls_names);
-		cls_names = NULL;
-	}
-
-	return 0;
-}
-
-void cls_names_uninit(void)
-{
-	db_names_free(cls_names);
-}
-
-const char *get_tc_lib(void)
-{
-	const char *lib_dir;
-
-	lib_dir = getenv("TC_LIB_DIR");
-	if (!lib_dir)
-		lib_dir = LIBDIR "/tc/";
-
-	return lib_dir;
-}
-
-int get_qdisc_handle(__u32 *h, const char *str)
-{
-	__u32 maj;
-	char *p;
-
-	maj = TC_H_UNSPEC;
-	if (strcmp(str, "none") == 0)
-		goto ok;
-	maj = strtoul(str, &p, 16);
-	if (p == str || maj >= (1 << 16))
-		return -1;
-	maj <<= 16;
[…]

### 2019-11-28 tex/paper.tex
Update presentation

@@
 % Modify this to remove all ACM and conference information (for arXiv for example)
-\def\removeHeaders{no}
+\def\removeHeaders{yes}
 
@@
 \def\tempNo{no}
-\documentclass[10pt,sigconf,letterpaper,dvipsnames\ifx\removeHeaders\tempYes ,nonacm\fi]{acmart}
+\documentclass[sigconf, 10pt, letterpaper,dvipsnames\ifx\removeHeaders\tempYes ,nonacm\fi]{acmart}
+
+%\documentclass[sigconf]{acmart}
+
+% Copyright
+\renewcommand\footnotetextcopyrightpermission[1]{} % removes footnote with conference info
+\setcopyright{none}
+%\setcopyright{acmcopyright}
+%\setcopyright{acmlicensed}
+%\setcopyright{rightsretained}
+%\setcopyright{usgov}
+%\setcopyright{usgovmixed}
+%\setcopyright{cagov}
+%\setcopyright{cagovmixed}
+
+\settopmatter{printacmref=false, printccs=false, printfolios=true}
+
+% DOI
+\acmDOI{}
+
+% ISBN
+\acmISBN{}
+
+%Conference
+%\acmConference[Submitted for review to SIGCOMM]{}
+%\acmYear{2018}
+%\copyrightyear{}
+
+%% {} with no args suppresses printing of the price
+\acmPrice{}
+
 \acmYear{2019}
@@
 
-\ifx\removeHeaders\tempYes
-\settopmatter{printacmref=false} % Removes citation information below abstract
-\renewcommand\footnotetextcopyrightpermission[1]{} % removes footnote with conference information in first column
-\fi
+%\ifx\removeHeaders\tempYes
+%\settopmatter{printacmref=false} % Removes citation information below abstract
+%\renewcommand\footnotetextcopyrightpermission[1]{} % removes footnote with conference information in first column
+%\fi
 
@@
 We propose the new AQM scheduler \textit{cocoa} based on fair queuing, which adapts the buffer size depending on the needs of each flow without requiring active participation from the endpoint.
-We implement this scheduler for the Linux kernel and show that it interacts well with the most common congestion control algorithms and can significantly increase throughput compared to \codel{} while avoiding overbuffering.
+We implement this scheduler for the Linux kernel and show that it interacts well with the most common congestion control algorithms and can significantly increase throughput compared to fair \codel{} while avoiding overbuffering.
 \end{abstract}
@@
 
-We show that a prototype of our solution \textit{\gls{cocoa}} can achieve these objectives for the most common loss-based \glspl{cca}, Reno \cite {jacobson_congestion_1988} and Cubic \cite{ha_cubic:_2008}.
+We show that a prototype of our solution \textit{\gls{cocoa}} can achieve the aforementioned objectives for the most common loss-based \glspl{cca}, Reno \cite {jacobson_congestion_1988} and Cubic \cite{ha_cubic:_2008}.
 These \glspl{cca} operate by continuously increasing the number of bytes that are allowed to be in the network (congestion window) if no packet loss is experienced and by sharply decreasing this number if a packet is lost (multiplicative decrease).
@@
 From these considerations it follows that for our experiments we generally use a multiplier of $1.25$.
-The behavior is different for BBR~v2 which drastically changed the probing phase and now consists of a random 2-3\,s phase of constant sending rate (cruising) followed by a quick probing phase followed by a decrease phase \cite{cardwell_bbr_2019}.
+The behavior is different for BBR~v2 which drastically changed the probing phase and now consists of a random 2-3\,s phase with a constant sending rate (cruising) followed by a quick probing phase followed by a decrease phase \cite{cardwell_bbr_2019}.
 We thus conjecture that a multiplier of $0.5$ is sufficient for BBR~v2.
@@
 First we evaluate if \gls{cocoa} is able to maintain a small buffer like it is necessary in case of small \glspl{bdp}.
-As can be seen in \autoref{fig:fq} the regular fq qdisc maintains a standing queue and unnecessarily leads to a significant increase in \gls{rtt} (100\,ms minimum, when the real minimum is 10\,ms).
-In contrast, \autoref{fig:cn} maintains the full throughput while at the same time keeping the delay minimal and not having a standing queue.
+As can be seen in \autoref{fig:fq} the regular fq qdisc (with a standard queue size of 100 packets) maintains a standing queue and unnecessarily leads to a significant increase in \gls{rtt} (100\,ms minimum, when the real minimum is 10\,ms).
+In contrast, \autoref{fig:cn} maintains full throughput while at the same time keeping the delay minimal and not having a standing queue.
 Also, when drastically changing the bandwidth by halving it (\autoref{fig:cn}) or doubling it (\autoref{fig:cn2}) \gls{cocoa} rapidly adapts to the new conditions and returns to the optimum state.
@@
 Besides fq\_codel possibly leading to lower throughput as we have shown, another problem is that for links with a small \gls{bdp} fq\_codel can keep a standing queue akin to fq since fq\_codel only wants queuing delay to fall below 5\,ms once every 100\,ms.
-%For example, on a link with a base \gls{rtt} of 5\,ms, fq\_codel would accept a permanent standing queue of 4.9\,ms, leading to overall \gls{rtt} being doubled, which is not required for achieving optimal throughput.
-We also ran experiments for these low \gls{bdp} scenarios: For a link speed of 20\;Mbit/s and a base \gls{rtt} of 1\;ms for a Cubic flow both \gls{cocoa} and fq\_codel achieve 98\% utilization.
+For example, on a link with a base \gls{rtt} of 1\,ms, fq\_codel would accept a permanent standing queue of 4.9\,ms, leading to overall \gls{rtt} being more than doubled, which is not required for achieving optimal throughput.
+%We also ran experiments for these low \gls{bdp} scenarios: For a link speed of 20\;Mbit/s and a base \gls{rtt} of 1\;ms for a Cubic flow both \gls{cocoa} and fq\_codel achieve 98\% utilization.
 The average \gls{rtt} is 6.86\,ms for \gls{cocoa} and 7.31\;ms for fq\_codel.
@@
 
-We see this work as an initial step towards flow-adaptive fair queuing at bottlenecks.
+We see this work as an initial step towards learning and flow-adaptive fair queuing at bottlenecks and think that the most promising application domain of \gls{cocoa} is at the Internet's edge.
 An interesting further improvement could be to not statically initialize the buffer with a constant 100 packets but instead use experience from previous flows.

### 2020-07-18 tex/paper.tex
Fix parameters in test.py

@@
 \def\tempNo{no}
-\documentclass[sigconf, 10pt, letterpaper,dvipsnames\ifx\removeHeaders\tempYes ,nonacm\fi]{acmart}
+\documentclass[sigconf, 10pt, letterpaper, dvipsnames\ifx\removeHeaders\tempYes ,nonacm\fi]{acmart}
 
@@
 % Copyright
-\renewcommand\footnotetextcopyrightpermission[1]{} % removes footnote with conference info
-\setcopyright{none}
+%\renewcommand\footnotetextcopyrightpermission[1]{} % removes footnote with conference info
+%\setcopyright{none}
 %\setcopyright{acmcopyright}
@@
 
-\settopmatter{printacmref=false, printccs=false, printfolios=true}
+%\settopmatter{printacmref=false, printccs=false, printfolios=true}
 
-% DOI
-\acmDOI{}
-
-% ISBN
-\acmISBN{}
+\copyrightyear{2019}
+\acmYear{2019}
+\setcopyright{acmlicensed}
+\acmConference[BS '19]{2019 Workshop on Buffer Sizing}{December 2--3, 2019}{Palo Alto, CA, USA}
+\acmBooktitle{2019 Workshop on Buffer Sizing (BS '19), December 2--3, 2019, Palo Alto, CA, USA}
+\acmPrice{15.00}
+\acmDOI{10.1145/3375235.3375236}
+\acmISBN{978-1-4503-7745-4/19/12}
 
@@
 %% {} with no args suppresses printing of the price
-\acmPrice{}
-
-\acmYear{2019}
-\copyrightyear{2019}
-\acmConference[BS'19]{Buffer Sizing Workshop}{December 2-3}{Stanford, CA}
+%\acmPrice{}
+%
+%\acmYear{2019}
+%\copyrightyear{2019}
+%\acmConference[BS'19]{Buffer Sizing Workshop}{December 2-3}{Stanford, CA}
 

### 2020-07-18 test.py

@@ -35,9 +35,9 @@ parser = argparse.ArgumentParser()
 parser.add_argument('--bytes_to_capture', type=int, default=100)
 parser.add_argument('--cport', type=int, default=60000)
 parser.add_argument('--delay_to_add', type=int, default=10)
-parser.add_argument('--rate', type=int, default=20)
+parser.add_argument('--rate', type=float, default=20)
 parser.add_argument('--mtu', type=int, default=1514)
-parser.add_argument('--time', type=int, default=60)
+parser.add_argument('--time', type=float, default=60)
 parser.add_argument('--change', type=float, default=0.5)
 parser.add_argument('--qdisc', type=str, default="cocoa")
 parser.add_argument('--cc', type=str, default="cubic")

### 2020-10-21 tex/figures/cocoa_common_beginning.tex
Readd cocoa_intervals

--- 2019-10-17 tex/figures/traq_common_beginning.tex
@@
   mark=at position #1 with {\fill[color=red] circle (0.5mm);}},postaction={decorate}}}  
+\tikzset{-dot3-/.style={decoration={
+  markings,
+  mark=at position #1 with {\draw[-,draw=none,fill=white](0.025,-0.05) -- (-0.025,0.05) -- (-0.015,0.05) -- (0.035,-0.05) -- cycle;%
+\draw[-,color=black](0.025,-0.05) -- (-0.025,0.05); \draw[-,color=black](0.035,-0.05) -- (-0.015,0.05);}},postaction={decorate}}}
     \tikzset{
